---
title: "Hw 1 by Kristin Abijaoude"
author: "Kristin Abijaoude"
description: "LungCap and Prison data HW1"
date: "02/27/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Hw1
  - kristin abijaoude
  - desriptive statistics 
  - probability
---

```{r}
library(ggplot2)
library(readxl)
library(dplyr)
library(summarytools)
library(stats)
```

# Lung Capacity

```{r}
LungCapData <- read_excel("~/Documents/GitHub/Github Help/603_Spring_2023/posts/_data/LungCapData.xls")
LungCapData
```

```{r}
print(dfSummary(LungCapData,
                        varnumbers = FALSE,
                        plain.ascii  = FALSE, 
                        style        = "grid", 
                        graph.magnif = 0.70, 
                        valid.col    = FALSE),
      method = 'render',
      table.classes = 'table-condensed')
```

```{r}
colnames(LungCapData)
dim(LungCapData)
```

```{r}
hist(LungCapData$LungCap)
```
1a. The distribution looks pretty normal to me, with capacity between 6 and 9 being the most frequent. 

```{r}
boxplot(LungCap ~ Gender, data=LungCapData)
```
1b. Separating the two genders, it looks like men have a higher lung capacity rate in comparison to women.

```{r}
LungCapData %>%
  group_by(Smoke) %>%
  summarise(mean = mean(LungCap), n = n())
```
1c. The average lung capacity for a non-smoker is around 7.78, while for smokers it's 8.65. In other words, on average, the smokers have a higher lung capacity rate than non-smokers... this doesn't make sense because smoking is supposed to be bad for your lungs.

```{r}
agegroup <- LungCapData %>%
  mutate(agegroup = case_when(Age <= 13 ~ "Less than 13 years old",
                              Age == 14| Age == 15 ~ "14 to 15 years old",
                              Age == 16 | Age == 17 ~ "16 to 17 years old",
                              Age >= 18 ~ "18 years old and older"))
agegroup %>%
  ggplot(aes(x=LungCap, fill=Smoke)) +
  geom_histogram() +
  facet_wrap(~agegroup)
```
1d. Obviously, older teens are more likely to be smokers, as well as have higher lung capacity, than younger teens. The vast majority of teens 13 years and younger are non-smoker (I would be horrified at the sight of a kid smoking).

```{r}
agegroup <- agegroup %>%
  mutate(AgeGroup = factor(agegroup, level= c("Less than 13 years old", 
                                              "14 to 15 years old",
                                              "16 to 17 years old",
                                              "18 years old and older")))

boxplot(LungCap ~ AgeGroup, data=agegroup)
```

1e. There is a correlation between age and lung capacity. The lung capacity rate increases as the person gets older.

# Prior Convictions

Another dataset I created here deals with prison convictions. The sample size is 810 prisoners in a state prison, some of the prisoners are there for the first time, while others have been imprison as many as 4 times, or have 4 prior convictions in other words. `prior` means numbers of prior convictions. `freq` means how many prisoners have a set of convictions (434 prisoners have 1 prior convictions, 160 prisoners have 2 prior convictions etc.). Finally, I created a new variable called `probability`, where I divided the `freq` variable by the total number of prisoners, to denote the probability that a prisoner had a certain number of prior convictions.

```{r}
df <- data.frame(prior = c(0:4), 
                 freq = c(128, 434, 160, 64, 24)
                 )

df <- df %>%
  mutate(probability = freq/810)
df
```

```{r}
# alternatively
(dbinom(x = 1, size = 1, prob = 160/810))*100
```

2a. There is a less than 20% probability that a randomly selected inmate has exactly 2 prior convictions.

```{r}
128 + 434
(dbinom(x = 1, size = 1, prob = 562/810))*100
```

2b. There is a 69% probability that a randomly selected inmate has fewer than 2 prior convictions.

```{r}
128 + 434 + 160
(dbinom(x = 1, size = 1, prob = 722/810))*100
```

2c. There is a 89% probability that a randomly selected inmate has 2 or fewer prior convictions.

```{r}
64 + 24
(dbinom(x = 1, size = 1, prob = 88/810))*100
```

2d. There is a 10% probability that a randomly selected inmate has more than 2 prior convictions.

```{r}
prior <- df$prior
prob <- df$probability
freq <- df$freq

exval <- sum(prior*prob)
exval
```

2e. The expected value `exval`, or long term mean, is 1.28642. I separated the variables into its own set and multiplied `prior` (# of prior convictions) and `prob` (the probability a given prisoner has a certain number of prior convictions).

```{r}
# variance
var(rep(df$prior, df$freq))

# standard deviation
sd(rep(df$prior, df$freq))
```
The variance is 0.86, which mean the data is close to one another.

The standard deviation is 0.93, which means the data is more clustered around the mean.

