---
title: "Homework 2"
author: "Abigail Balint"
desription: "HW2 Responses"
date: "04/03/23"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - homework2
  - abigailbalint
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(readxl)
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Generating a table here to calculate standard deviation
```{r}
heart <- matrix(c(539, 19, 10, 847, 18, 9), ncol=3, byrow=TRUE)
colnames(heart) <- c('Sample Size','Mean Wait Time','Standard Deviation')
rownames(heart) <- c('Bypass','Angiography')
heart <- as.table(heart)
head(heart)
```

Calculating confidence interval for Bypass:
```{r}
stander1 <- 10/sqrt(539)
confidence_level <- 0.90 
tail_area <- (1-confidence_level)/2
t_score <- qt(p = 1-tail_area, df = 539-1)
t_score
CI <- c(19 - t_score * stander1,
        19 + t_score * stander1)
print(CI)




```

Calculating confidence interval for Angiography:
```{r}
stander2 <- 9/sqrt(847)
confidence_level <- 0.90 
tail_area <- (1-confidence_level)/2
t_score <- qt(p = 1-tail_area, df = 847-1)
t_score
CI <- c(18 - t_score * stander2,
        18 + t_score * stander2)
print(CI)
```

```{r}
BypassInt <- 19.70971 - 18.29029
AngiographyInt <- 18.50922-17.49078 
print(BypassInt)  
print(AngiographyInt) 
```
Question: Is the confidence interval narrower for angiography or bypass surgery?
Answer: The interval for angiography surgery is shorter at an interval of 1.01.

## Question 2

Here I am using a prop test function to find the p value - 

```{r}
prop.test(567, 1031, conf.level = 0.95)
```
Answer: To interpret this, I am looking at my confidence interval and seeing that the true population mean of those who believe college is needed for success is somewhere between 52-58% which makes sense because the reported percentage from this random sample who believe college is needed for success is 54%, inside of that range.

## Question 3

My calculations to find the sample size are step by step in the below code block:

```{r}
#Starting by trying to find margin of error:
#Formula is error= z * SD/sqrt of n
#By squaring the whole thing, can transform this formular to to n=z^2*SD^2/error^2
# Z score for 95% confidence is 1.96
1.96^2
#n=3.8416*SD^2/error^2
#Finding the range and dividing it by 4 since we know SD is quarter of the range
(200-30)/4
42.5^2
#Now formula looks like this and just need to find standard error n=3.8416*1806.25/error^2
#We know that margin of error is 5 (within $5)
5^2
#Now we can fill in the final formula n=3.8416*1806.25/25
(3.8416*1806.25)/25
#n=277.56
```
Answer: The sample size should be about 278 (rounded)

## Question 4

A. Test whether the mean income of female employees differs from $500 per week. Include
assumptions, hypotheses, test statistic, and P-value. Interpret the result.

Assumptions - 
-Randomly generated sample
-From the same general population
-Distribution is normal

Null hypothesis - females earn  $500 a week
Alternative hypothesis - females earn more or less than $500 a week

Calculating the t value below using standard formula -

t= (sample mean-population mean)/(standard deviation/sample size^2)
```{r}
tvalue =  (410 - 500) / (90/sqrt(9))
tvalue
```
Answer: t=-3

This tells us that the mean of the female group is three standard deviations away from the mean of the overall group's pay.

B. Report the P-value for alternative hypothesis: μ < 500. Interpret.

Calculating p-value -

Formula pt(q = t, df =standardeviation-1, lower.tail = TRUE)

```{r}
pt(q = tvalue, df =8, lower.tail = TRUE)

```

Answer: .01 (rounded)

A p-value of .01 means we can reject the alternative hypothesis (mean for females < 500 a week)

C. Report and interpret the P-value for alternative hypothesis: μ > 500.

```{r}
pt(q = tvalue, df =8, lower.tail = FALSE)
```
 
 Answer: .99 (rounded)
This is the same but flipped, can reject the alternative hypothesis (mean for females > 500 a week)


## Question 5


Using formula t= sample mean-population mean/standard error

```{r}
t=(519.5-500)/10
print(t)
pt(q = t, df = 999, lower.tail = FALSE)*2
```
Answer: Jones t=1.95, p=.051

```{r}
t=(519.7-500)/10
print(t)
pt(q = t, df = 999, lower.tail = FALSE)*2
```
Answer: Smith t=1.97, p=.049

B. This makes Smith statistically significant because .049 falls below .05 but .051 does not.

C. This shows that results presented this way can be misleading because even though the p-values are extremely close here, one would report rejecting the null hypothesis and one wouldn't even though the differences in results are marginal.

## Question 6


Null hypothesis: Proportion of those who choose healthy snacks is not equal by grade level.

Generating table -
```{r}
snack <- matrix(c(31, 43, 51, 69, 57, 49), ncol=3, byrow=TRUE)
colnames(snack) <- c('6th','7th','8th')
rownames(snack) <- c('healthy','unhealthy')
snack <- as.table(snack)
head(snack)
```

Performing chi squared test - 

```{r}
chisq.test(snack, .05, correct = FALSE)

```
Since the p value is .01 we can assume that there is a difference by grade level in those who choose unhealthy vs healthy snacks, rejecting null hypothesis.

## Question 7


Null hypothesis: There is no difference in per-pupil costs between areas.

Generating data frame -

```{r}
area <- c(rep("Area1", 6), rep("Area2", 6), rep("Area3", 6))
cost <- c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5, 7.5, 8.2, 8.5, 8.2, 7.0, 9.3,
          5.8, 6.4, 5.6, 7.1, 3.0, 3.5)
tuition <- data.frame(area,cost)
head(tuition)
```

```{r}

anova <- aov(cost ~ area, data = tuition)
summary(anova)
```
Interpretation: The p value is very low indicating there is a difference and we can reject the null hypothesis.
: