---
title: "Homework 1"
author: "Abigail Balint"
desription: "Distributions and probability"
date: "02/23/23"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - homework1
  - abigailbalint
  - lungcap
  - prisoner
  - ggplot2
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(readxl)
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 - Lung Capacity 

Reading in LungCapData --

```{r, echo=T}

lung <- read_excel("_data/LungCapData.xls")
head(lung,2)
```
Looking at some basic descriptive stats --

```{r, echo=T}

glimpse(lung)
mean(lung$LungCap, na.rm = T)
var(lung$LungCap, na.rm = T)
sd(lung$LungCap, na.rm = T)
range(lung$LungCap, na.rm = T)
```

a) What does the distribution of LungCap look like? 

```{r}
ggplot(lung, aes(x = LungCap)) +
  geom_histogram()
```
The distribution looks relatively normal. There is a clear concentration of the sample around 7-8 and the outliers are only a very small portion of the sample.


b) Compare the probability distribution of the LungCap with respect to Males and Females

```{r}
ggplot(lung, aes(x=LungCap)) + 
    geom_boxplot(fill="slateblue", alpha=0.2) + 
    xlab("Lung Capacity") +
  facet_wrap("Gender")
```
The probability distribution is pretty similar between male and female, but males skew to a higher lung capacity overall and the median line is at around 8 whereas female is closer to 7.5. 

c) Compare the mean lung capacities for smokers and non-smokers. Does it make sense?

```{r}
lung %>%
  group_by(Smoke) %>%
  summarise(mean = mean(LungCap), n = n())
```
We would expect the lung capacities for non smokers to be higher but the mean for smokers is actually a little bit higher.

d) Examine the relationship between Smoking and Lung Capacity within age groups: “less than or  equal to 13”, “14 to 15”, “16 to 17”, and “greater than or equal to 18”. 

Recoding the age groups --

```{r}
lunggroup <- lung %>%
  mutate(`AgeGroup` = dplyr::case_when(
    `Age` >= 0 & `Age` < 14 ~ "0-13",
    `Age` >= 14 & `Age` < 16 ~ "14-15",
    `Age` >= 16 & `Age` < 18 ~ "16-17",
    `Age` >= 18 ~ "18+" ))
head(lunggroup)
```
Mean lung capacity by age group -- 

```{r}
lunggroup %>%
group_by(Smoke, AgeGroup) %>%
  summarise(mean = mean(LungCap), n = n())
```
For both smokers and non-smokers, the lung capacity goes up as the age increases with 18+ having the highest average capacity. In all age ranges besides 0-13 (the broadest range), the mean is higher for non-smokers than smokers.

e. Compare the lung capacities for smokers and non-smokers within each age group. Is your answer  different from the one in part c. What could possibly be going on here? 

```{r}
ggplot(lunggroup, aes(x=LungCap, fill=Smoke)) + 
    geom_boxplot() + 
    xlab("Lung Capacity") +
  facet_wrap("AgeGroup")
```
I'm seeing that the results by age group are slightly different than in part C. Above I can see that the average for all age ranges is higher for non-smokers, besides age group 0-13. I can see in my results in part D that the sample size for 0-13 non-smokers is extremely high, much higher than any other group of smokers or non-smokers, so with this higher sample size comes more variance. The median lines are actually pretty close but the outliers are probably affecting the mean.


## Question 2 

Creating a data frame -- 

```{r}
priorconviction <- c(0,1,2,3,4)
prisoners <- c(128,434,160,64,24)
q2 <- data.frame(priorconviction, prisoners)
head(q2)
```


a) What is the probability that a randomly selected inmate has exactly 2 prior convictions? 
```{r}
160/810
```
 I found it to be .1975 or 19.75%
 
b) What is the probability that a randomly selected inmate has fewer than 2 prior convictions? 
```{r}
(434+128)/810
```
To get this I added the sample of 0 or 1 prior conviction and it comes out to .69 or 69%.

c) What is the probability that a randomly selected inmate has 2 or fewer prior convictions? 
```{r}
(128+434+160)/810
```
To get this I added the sample of 0 or 1 or 2 prior convictions and it comes out to .89 or 89%.

d) What is the probability that a randomly selected inmate has more than 2 prior convictions? 
```{r}
(64+24)/810
```
To get this I added the sample of 3 or 4 prior convictions and it comes out to .108 or 11%.

e) What is the expected value1 for the number of prior convictions? 
```{r}
sum(q2$priorconviction*prisoners)
1042/810
```

To get this I summed all of the numbers of prior convictions by the amount of prisoners (1042) then divided this by total sample (810) to get a final expected value of 1.28 prior convictions.

f) Calculate the variance and the standard deviation for the Prior Convictions. 

```{r}
var(q2$priorconviction)
var(q2$priorconviction)*(5-1)/5
```

I used the above code to find a sample variance of 2.5 and a population variance of 2.

```{r}
sd(q2$priorconviction)
```

I used the standard deviation function to calculate the above.


:::