---
title: "Homework 2"
author: "Adithya Parupudi"
description: "my homework 2"
date: "02/27/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - Adithya Parupudi
---

```{r, echo=T}
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(readxl)
```


# Question 1

```{r}
procedure <- c("Bypass", "Angiography")
s_size <- c(539, 847)
mean_wait_time <- c(19, 18)
s_sd <- c(10, 9)

surgery <- data.frame(procedure, s_size, mean_wait_time, s_sd)

colnames(surgery) <- c("Surgical Procedure", "Sample Size","Mean Wait Time", "Standard Deviation")
surgery
```

```{r}
standard_error <- s_sd / sqrt(s_size)
standard_error
```

```{r}
confidence_level <- 0.90
tail_area <- (1-confidence_level)/2
tail_area
```

```{r}
t_score <- qt(p = 1-tail_area, df = s_size-1)
t_score
```

```{r}
confidence_interval <- c(mean_wait_time - t_score * standard_error,mean_wait_time + t_score * standard_error)
confidence_interval
```

We can be 90% confident that the population mean wait time for the bypass procedure is between 18.29029 and 19.70971 days. And for the angiography procedure, We can be 90% confident that the population mean wait time is between 17.49078 and 18.50922 days.

From the above results, confidence interval of angiography procedure is narrower.


# Question 2
```{r}
prop.test(567, 1031, conf.level = .95)
```

# Question 3

```{r}
ME <- 5
z <- 1.96
s_sd <- (200-30)/4

sample_size <- ((z*s_sd)/ME)^2
sample_size
```

The necessary sample size is 278.

# Question 4

## 4A

We assume that the sample is random and that the population has a normal distribution.
Null hypothesis: H0: μ = 500
Alternative hypothesis: Ha: μ ≠ 500
We will reject the null hypothesis at a p-value, p <= 0.05

```{r}
# defining variables
s_mean <- 410
μ <- 500
s_sd <- 90
s_size <- 9
```



```{r}
# Calculating test-statistic
t_score <- (s_mean-μ)/(s_sd/sqrt(s_size))
t_score
```

```{r}
# Calculating p-value

p <- 2*pt(t_score, s_size-1)
p
```

The test-statistic is -3 and p-value is 0.01707168. As p-value is less than the 0.05, we reject the null hypothesis. Therefore, the mean income of female employees is not equal to $500.

## 4B

We assume that the sample is random and that the population has a normal distribution.
Null hypothesis: H0: μ = 500
Alternative hypothesis: Ha: μ < 500
We will reject the null hypothesis at a p-value less than 0.05

```{r}
p_val <- pt(t_score, s_size-1, lower.tail = TRUE)
p_val
```

The p-value is 0.008535841. As p-value is less than the 0.05, we reject the null hypothesis. Therefore, the mean income of female employees is less than $500.

## 4C

We assume that the sample is random and that the population has a normal distribution.
Null hypothesis: H0: μ = 500
Alternative hypothesis: Ha: μ > 500
We will reject the null hypothesis at a p-value less than 0.05

```{r}
p_val_c <- pt(t_score, s_size-1, lower.tail = FALSE)
p_val_c
```

The p-value is 0.9914642. As p-value is less than the 0.05, we reject the null hypothesis. Therefore, the mean income of female employees is greater than $500.


# Question 5

## 5A

We assume that the sample is random and that the population has a normal distribution.
Null hypothesis: H0: μ = 500
Alternative hypothesis: Ha: μ ≠ 500
We will reject the null hypothesis at a p-value less than 0.05


```{r}
# Calculating t-statistic and p-value for Jones
s_mean <- 519.5
μ <- 500
se <- 10
s_size <- 1000

james_t <- (s_mean-μ)/se
james_t

p <- 2*pt(james_t, s_size-1, lower.tail = FALSE)
p
```



```{r}
# Calculating t-statistic and p-value for Smith
s_mean <- 519.7
μ <- 500
se <- 10
s_size <- 1000

smith_t <- (s_mean-μ)/se
smith_t

p <- 2*pt(smith_t, s_size-1, lower.tail = FALSE)
p
```

The test-statistic is 1.95, p-value is 0.05145555 for Jones and the test-statistic is 1.97, p-value is 0.05145555 for Smith.

## 5B

The p-value is 0.05145555 for Jones. As p-value is greater than the 0.05, we fail to reject the null hypothesis. The p-value is 0.04911426 for Jones. As p-value is less than the 0.05, we reject the null hypothesis. Therefore, the result is statistically significant for Smith, but not Jones.

## 5C

If we fail to report the P-value and simply state whether the P-value is less than/equal to or greater than the defined significance level of the test, one cannot determine the strength of the conclusion. In the Jones/Smith example, reporting the results only as *P ≤ 0.05* versus *P > 0.05* will lead to different conclusions about very similar results.

# Question 6
To test the claim that the proportion of students who choose a healthy snack differs by grade level, we will use the chi-squared test of independence. 
The null hypothesis is that the proportion of students who choose a healthy snack is the same across all grade levels. 
The alternative hypothesis is that the proportion of students who choose a healthy snack differs by grade level.

```{r}
# Data in a matrix
table_6 <- matrix(c(31, 43, 51, 69, 57, 49), nrow = 2, byrow = TRUE)

# Row and column names
rownames(table_6) <- c("Healthy snack", "Unhealthy snack")
colnames(table_6) <- c("6th grade", "7th grade", "8th grade")

# Chi-squared test of independence
test <- chisq.test(table_6)
test


```
The p-value (0.01547) is less than the significance level (α = 0.05). Therefore, we reject the null hypothesis and conclude that the proportion of students who choose a healthy snack differs by grade level.)




# Question 7
To test the claim that there is a difference in means for the three areas, we will use a one-way ANOVA test. 
The null hypothesis is that the means for the three areas are the same. 
The alternative hypothesis is that at least one area's mean is different from the others.

```{r}

# Data in vectors
area1 <- c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5)
area2 <- c(7.5, 8.2, 8.5, 8.2, 7.0, 9.3)
area3 <- c(5.8, 6.4, 5.6, 7.1, 3.0, 3.5)

# One-way ANOVA test
test <- aov(c(area1, area2, area3) ~ factor(c(rep("Area 1", length(area1)), rep("Area 2", length(area2)), rep("Area 3", length(area3)))))
summary(test)

```
The p-value (0.00397) is less than the significance level (α = 0.05). Therefore, we reject the null hypothesis and conclude that there is a difference in means for the three areas.
