---
title: "Homework1 - EDA of LungCap Data"
author: "Adithya Parpudi"
description: "HW1 submission"
date: "02/23/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - Adithya Parupudi
---

# Libraries

```{r}
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(readxl)
```


# Read data

```{r, echo=T}

df <- read_excel("_data/LungCapData.xls")
```


# Answering Questions
## Question 1

### 1a) The distribution of LungCap looks as follows:

```{r, echo=T}
hist(df$LungCap)
```

The histogram suggests that the distribution is close to a normal distribution. Most of the observations are close to the mean. Very few observations are close to the margins (0 and 15).

### 1b) Compare the probability distribution of the LungCap with respect to Males and Females? (Hint:make boxplots separated by gender using the boxplot() function)

```{r}


#boxplot code

df %>%
  ggplot( aes(x=Gender, y=LungCap, fill=Gender)) +
    geom_boxplot() +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=12)
    ) +
    ggtitle("Lungcap vs Gender") +
    xlab("Gender") +
  ylab("Lung Cap")


```


### c) Compare the mean lung capacities for smokers and non-smokers. Does it make sense?
```{r}
mean_smoke <- df %>%
  group_by(Smoke) %>%
  summarise(mean = mean(LungCap))
mean_smoke

```
According to this mean, it doesn't make sense that lung capacities of smokers is greater than that of non-smokers. 


### d) Examine the relationship between Smoking and Lung Capacity within age groups: “less than or
equal to 13”, “14 to 15”, “16 to 17”, and “greater than or equal to 18”.
```{r}

df <- mutate(df, AgeGrp = case_when(Age <= 13 ~ "less than or equal to 13",
                                    Age == 14 | Age == 15 ~ "14 to 15",
                                    Age == 16 | Age == 17 ~ "16 to 17",
                                    Age >= 18 ~ "greater than or equal to 18"))

df %>%
  ggplot(aes(y = LungCap, color = Smoke)) +
  geom_histogram(bins = 25) +
  facet_wrap(vars(AgeGrp)) +
  labs(title = "Relationship of LungCap and Smoke based on Age", y = "Lung Capacity", x = "Frequency")

```


### e) Compare the lung capacities for smokers and non-smokers within each age group. Is your answer different from the one in part c. What could possibly be going on here?
```{r}
df %>%
  ggplot(aes(x = Age, y = LungCap, color = Smoke)) +
  geom_line() +
  theme_classic() + 
  facet_wrap(vars(Smoke)) +
  labs(title = "Relationship of LungCap and Smoke v/s Age", y = "Lung Capacity", x = "Age")

```


## Question 2

### Reading the table

```{r}
Prior_convitions <- c(0:4)
Inmate_count <- c(128, 434, 160, 64, 24)
Pc <- data_frame(Prior_convitions, Inmate_count)
Pc
```

```{r}
Pc <- mutate(Pc, Probability = Inmate_count/sum(Inmate_count))
Pc
```


### 2a - Probability that a randomly selected inmate has exactly 2 prior convictions:

```{r}
Pc %>%
  filter(Prior_convitions == 2) %>%
  select(Probability)
```

### 2b - Probability that a randomly selected inmate has fewer than 2 convictions:

```{r}
temp <- Pc %>%
  filter(Prior_convitions < 2)
sum(temp$Probability)
```

### 2c - Probability that a randomly selected inmate has 2 or fewer prior convictions:

```{r}
temp <- Pc %>%
  filter(Prior_convitions <= 2)
sum(temp$Probability)
```

### 2d - Probability that a randomly selected inmate has more than 2 prior convictions:

```{r}
temp <- Pc %>%
  filter(Prior_convitions > 2)
sum(temp$Probability)
```

### 2e - Expected value for the number of prior convictions:

```{r}
Pc <- mutate(Pc, Wm = Prior_convitions*Probability)
e <- sum(Pc$Wm)
e
```

### 2f - Variance for the Prior Convictions:

```{r}
v <-sum(((Pc$Prior_convitions-e)^2)*Pc$Probability)
v
```

standard deviation for the Prior Convictions:

```{r}
sqrt(v)
```

