---
title: "Homework 2"
author: "Asch Harwood"
description: "He"
date: "02/05/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Q1

1. The time between the date a patient was recommended for heart surgery and the surgery date
for cardiac patients in Ontario was collected by the Cardiac Care Network (“Wait Times Data
Guide,” Ministry of Health and Long-Term Care, Ontario, Canada, 2006). The sample mean
and sample standard deviation for wait times (in days) of patients for two cardiac procedures
are given in the accompanying table. Assume that the sample is representative of the Ontario
population


Construct the 90% confidence interval to estimate the actual mean wait time for each of the two
procedures. Is the confidence interval narrower for angiography or bypass surgery?

```{r}
# create data frame
df <- data.frame(
  "Surgical Procedure" = c("Bypass", "Angiography"),
  "Sample Size" = c(539, 847),
  "Mean Wait Time" = c(19, 18),
  "Standard Deviation" = c(10, 9)
)

# print data frame
df

```
```{r}
?t.test
t.test(x = 19, n = 539, sd = 10)
```

