---
title: "Estimating National Food Waste with State/Local Waste Characterization Studies in the United States"
author: "Asch Harwood"
description: "Final Project Part 1: Proposal"
date: "03/20/23"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart1
editor: 
  markdown: 
    wrap: 72
---

```{r message= FALSE}
library("dplyr")
library("knitr")
library(kableExtra)
library(xtable)
```

# Background

ReFED, a national nonprofit that advocates for the reduction of food
waste, maintains several models to estimate the amount of wasted food
and its greenhouse gas emissions in the United States. Collectively,
U.S. residents and businesses throw out roughly 90 million tons of food
annually, contributing nearly 5 percent of the total net US greenhouse
gas emissions.

Creating system-level estimates in the food supply chain is notoriously
difficult. To produce these estimates, ReFED employs a 'mass balance'
accounting approach. For each sector, ReFED estimates the value of the
total supply of food in US dollars. It then applies a price per lb
conversion factor to convert dollars to tons to estimate the total suppy
of food in weight. ReFED then draws on academic and private-sector
studies to determine 'surplus rates'--the share of food that is not
going to its intended destination. The supply of food is multiplied by
these surplus rates to determine the total amount of surplus food in a
given place and year.

A limitation of this approach, however, is that the research to estimate
surplus rates is rarely, if ever, repeated at regular intervals. As a
result, the lack of year-over-year and geographic variation in surplus
rates makes it incredibly difficult to identify systematic changes in
the food waste behaviors.

# Proposal

I propose utilizing state and local waste characterization studies
combined with 'easily observable' and 'widely available' predictors
(i.e. food supply, population, population density, GDP, unemployment,
food prices, consumer confidence, investment/spending in food waste
reduction technologies, knowledge and attitudes about food
waste/environmental issues) to build a weighted mixed effects linear
regression model. The goal is estimate the share of food waste of total
municipal solid waste at the national level in a given year.

State and local governments periodically conduct inferential waste
characterization studies, where they randomly select several 200 - 300
lb samples from landfills and other waste management facilities, and
separate the garbage to determine the composition of municipal solid
waste in a particular area over a predetermined time. These studies are
then used to set waste management budgets, 'right-size' landfill
capacity, and create/evaluate programs meant to divert recyclable and
compostable materials from landfills. However, they are not conducted
regularly or consistently (or at all in many places), and, to my
knowledge, there are no nationally-representative waste characterization
studies.

Nevertheless, I believe there is the potential to creatively use these
studies to estimate the share of food waste in landfills at the
national-level over time, and, hopefully, to determine if there are any
systematic changes in food waste management behavior in the commercial
and residential sectors.

### Research Questions

1.  Is it possible to use state and local waste characterization studies
    to estimate the share of food waste in municipal solid waste at the
    national-level year over year?

2.  What 'easily-observable' and 'widely-available' predictors allow us
    to estimate the share of food waste in the absence of
    regularly-conducted, national-level samples?

3.  How 'good' does the model need to be to be "operationally" useful by
    food businesses, policymakers, and households?

4.  How do food waste estimates produced using a bottom-up method
    proposed here compare to top-down, mass balanced-based approaches?

### Hypothesis

**Population**

H0: Population is not correlated with percent food waste

H1: Population is correlated with percent food waste

**Food Supply**

H0: Food supply is not correlated with percent food waste

H1: Food supply is correlated with percent food waste

**GDP**

H0: GDP growth is not correlated with percent food waste

H1: GDP growth is positively correlated with percent food waste

**Unemployment**

H0: Unemployment is not correlated with percent food waste

H1: Unemployment is negatively correlated with percent food waste

**Food Prices**

H0: Food prices are not correlated with percent food waste

H1: Food prices are negatively correlated with percent food waste

**Consumer Confidence**

H0: Consumer confidence is not correlated with percent food waste

H1: Consumer confidence is positively correlated with percent food waste

**Consumer Knowledge**

H0: Consumer knowledge/concern about climate change is not correlated
with percent food waste

H1: Consumer knowledge/concern is negatively correlated with percent
food waste

**Investment**

H0: Investment in food waste reduction technologies is not correlated
with percent food waste

H1: Investment in food waste reduction technologies is positively
correlated with percent food waste

### Exploratory Data Analysis

Currently, no comprehensive aggregate municipal solid waste
characterization datasets exisit for the United States in the public
domain. To address this, I have compiled roughly 300 waste
characterization studies, of which I have parsed and coded roughly 50 to
date, based on the following (rough) criteria[^1]. To be eligible to be
included in my dataset, each study must be:

[^1]: I recognize that we were encouraged to work with pre-existing
    datasets. Nevertheless, I view this as an opportunity to prototype a
    project I believe might have real-world applicability in my
    professional life.

-   conducted in the United States and be representative of the state,
    county, or local level between the years of 1999 and 2023
-   infer the amount of total food as a share of the total amount of
    municipal solid waste for an entire year
-   report the number of samples, which need to be between 200 - 300 lbs
    each

```{r}
waste <- read.csv(url("https://raw.githubusercontent.com/AschHarwood/603_Spring_2023/asch_603_spring_2023/posts/_data/wasteCharacterizationStudies_v1.csv"))
waste$percent_food <- round(waste$percent_food, digits=4)
table <- na.omit(waste) %>%
  sample_n(10)
kable(table, caption = "Waste Characterization Dataset", format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  row_spec(0:nrow(table), font_size = 12)


```

```{r}
cat("Summary of Waste Characterization\n")
cat('\n')
summary(waste)


```

```{r}
cat('Summary of Waste Characterization\n')
cat('\n')
str(waste)
```

```{r}
# Create a data frame with the data dictionary
data_dictionary <- data.frame(
  `col_name` = c("year", "state", "geographic_scope", "geographic_scope_name", "sector_scope", "sample_size", "percent_food"),
  `d_type` = c("int", "chr", "chr", "chr", "chr", "int", "dbl"),
  description = c(
    "The year when the study was conducted",
    "The U.S. state where the study was conducted",
    "The scope of the geographic area covered by the study (e.g., state, county, or city)",
    "The name of the specific geographic area covered by the study (e.g., state name or city name)",
    "The sector scope of the study (e.g., commercial, residential, or transfer-truck)",
    "The number of samples included in the study",
    "The percentage of food waste in the studied area"
  )
)
kable(data_dictionary)
```
