---
title: "Homework 4"
author: "Asch Harwood"
description: "Homework 4, Intro to Quant"
date: "4/25/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw4
---

```{r}
library("dplyr")
library("knitr")
library(kableExtra)
library(xtable)
library(ggplot2)
library(GGally)
library(lme4)
library(car)
library(dplyr)
```

# Question 1

For recent data in Jacksonville, Florida, on - y = selling price of home (in dollars)

-   x1 = size of home(in square feet)

-   x2 = lot size (in square feet),

-   prediction equation \<- ŷ = −10,536 + 53.8x1 + 2.84x2.

### A

A particular home of 1240 square feet on a lot of 18,000 square feet sold for \$145,000. Find the predicted selling price and the residual, and interpret.

```{r}
x1 <-  1240 #house size
x2 <- 18000 #lot size
y_observed <- 145000
```

```{r}
# find predicted selling price
y_hat <- (-10536) + (53.8*x1) + (2.84*x2)
residual <- y_observed - y_hat
cat('Predict sale price: $',  y_hat)
cat('\n')
cat('Residual: $', residual)
cat('\n')

cat('The predicted sale price of', y_hat, 'dollars is', residual, 'dollars less than the actual observed house price of', y_observed, '.')
```

### B

Holding the lot size fixed, each unit increase in square footage increases the house value by 53.8 dollars. This is because the square footage coefficient is 53.8. 

###C

The lot size would need to increase by 19.94 square feet to increase the home value by 53.8 dollars.

# Question 2

(Data file: salary in alr4 R package). The data file concerns salary and other characteristics of all
faculty in a small Midwestern college collected in the early 1980s for presentation in legal
proceedings for which discrimination against women in salary was at issue. All persons in the data
hold tenured or tenure track positions; temporary faculty are not included. The variables include
degree, a factor with levels PhD and MS; rank, a factor with levels Asst, Assoc, and Prof; sex, a
factor with levels Male and Female; Year, years in current rank; ysdeg, years since highest degree,
and salary, academic year salary in dollars.

### A 

Test the hypothesis that the mean salary for men and women is the same, without regard to
any other variable but sex. Explain your findings.

### B 

Run a multiple linear regression with salary as the outcome variable and everything else as
predictors, including sex. Assuming no interactions between sex and the other predictors,
obtain a 95% confidence interval for the difference in salary between males and females.

### C

Interpret your finding for each predictor variable; discuss (a) statistical significance, (b)
interpretation of the coefficient / slope in relation to the outcome variable and other variables


###D 

Change the baseline category for the rank variable. Interpret the coefficients related to rank
again.

###E 

Finkelstein (1980), in a discussion of the use of regression in discrimination cases, wrote,
“[a] variable may reflect a position or status bestowed by the employer, in which case if there
is discrimination in the award of the position or status, the variable may be ‘tainted.’ ” Thus,
for example, if discrimination is at work in promotion of faculty to higher ranks, using rank to
adjust salaries before comparing the sexes may not be acceptable to the courts.
Exclude the variable rank, refit, and summarize how your findings changed, if they did.

### F 

Everyone in this dataset was hired the year they earned their highest degree. It is also
known that a new Dean was appointed 15 years ago, and everyone in the dataset who
earned their highest degree 15 years ago or less than that has been hired by the new Dean.
Some people have argued that the new Dean has been making offers that are a lot more
generous to newly hired faculty than the previous one and that this might explain some of
the variation in Salary.

Create a new variable that would allow you to test this hypothesis and run another multiple
regression model to test this. Select variables carefully to make sure there is no
multicollinearity. Explain why multicollinearity would be a concern in this case and how
you avoided it. Do you find support for the hypothesis that the people hired by the new
Dean are making higher than those that were not?

# Question 3
(Data file: house.selling.price in smss R package)

### A 

Using the house.selling.price data, run and report regression results modeling y = selling
price (in dollars) in terms of size of home (in square feet) and whether the home is new (1 =
yes; 0 = no). In particular, for each variable; discuss statistical significance and interpret the
meaning of the coefficient.

### B

Report and interpret the prediction equation, and form separate equations relating selling
price to size for new and for not new homes.

### C

Find the predicted selling price for a home of 3000 square feet that is (i) new, (ii) not new.

### D 

Fit another model, this time with an interaction term allowing interaction between size and
new, and report the regression results

### E 

Report the lines relating the predicted selling price to the size for homes that are (i) new,
(ii) not new.

### F 

Find the predicted selling price for a home of 3000 square feet that is (i) new, (ii) not new.

###G 

Find the predicted selling price for a home of 1500 square feet that is (i) new, (ii) not new.
Comparing to (F), explain how the difference in predicted selling prices changes as the size
of home increases.

### H 

Do you think the model with interaction or the one without it represents the relationship of
size and new to the outcome price? What makes you prefer one model over another?


