---
title: "Homework 4"
author: "Asch Harwood"
description: "Homework 4, Intro to Quant"
date: "4/25/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw4
---

```{r}
library("dplyr")
library("knitr")
library(kableExtra)
library(xtable)
library(ggplot2)
library(GGally)
library(lme4)
library(car)
library(dplyr)
```

# Question 1

For recent data in Jacksonville, Florida, on - y = selling price of home (in dollars)

-   x1 = size of home(in square feet)

-   x2 = lot size (in square feet),

-   prediction equation \<- ŷ = −10,536 + 53.8x1 + 2.84x2.

### A

A particular home of 1240 square feet on a lot of 18,000 square feet sold for \$145,000. Find the predicted selling price and the residual, and interpret.

```{r}
x1 <-  1240 #house size
x2 <- 18000 #lot size
y_observed <- 145000
```

```{r}
# find predicted selling price
y_hat <- (-10536) + (53.8*x1) + (2.84*x2)
residual <- y_observed - y_hat
cat('Predict sale price: $',  y_hat)
cat('\n')
cat('Residual: $', residual)
cat('\n')

cat('The predicted sale price of', y_hat, 'dollars is', residual, 'dollars less than the actual observed house price of', y_observed, '.')
```

### B

Holding the lot size fixed, each unit increase in square footage increases the house value by 53.8 dollars. This is because the square footage coefficient is 53.8. 

###C

The lot size would need to increase by 19.94 square feet to increase the home value by 53.8 dollars.



