---
title: "Homework 3"
author: "Asch Harwood"
description: "Homework 3"
date: "4/10/23"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw3

---


```{r message = FALSE}
library(alr4)
library(smss)
library(ggplot2)
library(dplyr)
data(UN11)
data(water)
data("Rateprof")
data("student.survey")
```

# Q1

### a

Predictor: PPGDP
Response:  Fertility

### b

- A straight line does not seem like a good fit for this model as it is heavily skewed to the right

```{r}
ggplot(UN11, aes(x = ppgdp, y = fertility)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE)
```

### c

- By applying a log transformation to the x and y variables, there now appears to be a linear relationship between log(ppgdp) and log(fertility).

```{r}

ggplot(UN11, aes(x = log(ppgdp), y = log(fertility))) +
  geom_point(color = 'blue') +
  geom_smooth(method = 'lm', color = 'blue', se = FALSE) 
```


# Q2

### a

As we can see in both the scatter plot and the regression coefficients, the slope of the equation does not change. We would not expect it to since we are not altering the relationship between between our variables of interest, simply the unit it is denoted in. The only thing that changes is the intercept, since we shift the x values to the left by converting their units. 

### b

Likewise, the correlation does not change because we do not change the relationship by changing the unit of measurement.

```{r}
fit<- lm(log(fertility) ~ log(ppgdp), data = UN11)
summary(fit)
cor(log(UN11$ppgdp), log(UN11$fertility))
```

```{r}
fit<- lm(log(fertility) ~ log(ppgdp/1.33), data = UN11)
summary(fit)
cor(log(UN11$ppgdp/1.33), log(UN11$fertility))
```

```{r}
ggplot(UN11, aes(x = log(ppgdp), y = log(fertility))) +
  geom_point(color = 'blue') +
  geom_smooth(method = 'lm', color = 'blue', se = FALSE) +
  geom_point(aes(x=log(ppgdp/1.33)), color = 'red') +
  geom_smooth(aes(x=log(ppgdp/1.33)), method = 'lm', color = 'red', se = FALSE)
```

# Q3

There are clear, linear relationships between runoff (BSAAM) and precipitation at OPBPC, OPRC, and OPSLAKE. There also may be linear relationships between APMAM, APSB, and APSLake, but they appear to be right-skewed. 

These same patterns also hold when comparing precipitation locations. The O locations display multicollinearity with each other, as do the A locations. Similarly, when compared to each other, the A locations and O locations might also be related, but those relations appear to be skewed. 

When comparing year to precipitation and runoff, there doesn't appear to a be a linear relationship. 


```{r}
pairs(water)
```


# Q4

- Quality, helpfulness, and clarity all exhibit clear, linear relationships
- Easiness also exhibits linear relationships with quality, helpfulness, and clarity but the strength of that relationship appears to be much weaker
- While we cannot rule out a linear relationship between raterInterest and the other four variables based on a visual inspection, that relationship, if it exists, is much weaker than the other observed relationships.
```{r}
col_list <- c('quality', 'helpfulness', 'clarity', 'easiness', 'raterInterest')
pairs(Rateprof[col_list])
```


# Q5

### i

Visually, there does seem to be a linear relationship between religiosity and political ideology. As religious participation increases, respondents move right on the political spectrum. This observed relationship is supported by our regression model. Compared to those who never attend church, and assuming the very liberal is the baseline, those who attend occasionally move 0.25 points to the right. Those who attend most weeks weeks are 2.16 points to the right on the political spectrum, and those who attend every week 2.6 points to the right from those who never attend church. However, occasionally is not statistically significant while most weeks and every week are statistically significant. The R-Squared metric shows us that religiosity explains about 39 percent of variation in political ideology. The entire model is statistically significant.


```{r}
ggplot(data = student.survey, aes(x = re, y = pi)) + 
  geom_count()
```



```{r}
re_dummies <- model.matrix(~ re - 1, data = student.survey)
re_dummies_df <- as.data.frame(re_dummies)
re_pi_df <- cbind(student.survey, re_dummies_df)
colnames(re_pi_df) <- gsub(" ", "_", colnames(re_pi_df))
pi_mapping <- c("very liberal" = -3, "liberal" = -2, "slightly liberal" = -1, "moderate" = 0, 
                "slightly conservative" = 1, "conservative" = 2, "very conservative" = 3)
re_pi_df$pi_numeric <- as.numeric(pi_mapping[as.character(re_pi_df$pi)])
re_pi_df$re <- factor(re_pi_df$re, ordered = FALSE)
re_pi_df$re <- relevel(factor(re_pi_df$re), ref = "never")
fit <- lm(pi_numeric ~ re, data = re_pi_df)
summary(fit)
```

### ii

Visually, a relationship is plausible. However, there appear to be a few outliers, which may be exerting undue influence. After dropping those outliers, a relationship seems less likely. The model suggests that for every 1 hour increase in tv, gpa drops by 0.02. However, this finding is not significant.

```{r}
data("student.survey")
ggplot(data = student.survey, aes(x = tv, y = hi)) +
  geom_point() +
  geom_smooth(method='lm')
```
```{r}
#dropping outliers
s_survey <- student.survey[student.survey$tv<=20,]
ggplot(data = s_survey, aes(x = tv, y = hi)) +
  geom_point() +
  geom_smooth(method='lm')
```


```{r}
fit <- lm(hi ~ tv, data = s_survey)
summary(fit)
```




