---
title: "Homework 4"
author: "Asch Harwood"
description: "Homework 4, Intro to Quant"
date: "4/25/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw4
---

```{r}
library("dplyr")
library("knitr")
library(kableExtra)
library(xtable)
library(ggplot2)
library(GGally)
library(lme4)
library(car)
library(dplyr)
```

# Question 1

For recent data in Jacksonville, Florida, on - y = selling price of home (in dollars)

-   x1 = size of home(in square feet)

-   x2 = lot size (in square feet),

-   prediction equation \<- ŷ = −10,536 + 53.8x1 + 2.84x2.

### A

A particular home of 1240 square feet on a lot of 18,000 square feet sold for \$145,000. Find the predicted selling price and the residual, and interpret.

```{r}
x1 <-  1240
x2 <- 18000
y_observed <- 145000
```

```{r}
# find predicted selling price
y_hat <- (-10536) + (53.8*x1) + (2.84*x2)
y_hat #selling price
```

Predict selling price: \${y_hat} .

### B

For fixed lot size, how much is the house selling price pre\`dicted to increase for each squarefoot increase in home size? Why?

###C

According to this prediction equation, for fixed home size, how much would lot size need to increase to have the same impact as a one-square-foot increase in home size
