---
title: "Logistic_Regression_Bullying"
format: html
editor: visual
---

## Research Question

Identifying the Factors which makes Students most vulnerable to be a victim of Bullying.

Calling the libraries

```{r}
library(tidyverse)
library(stringr)
library(rmarkdown)
library(knitr)

```

Reading the data

```{r}
data <- read.csv("")
```

Replacing the White Spaces with "NA"

```{r}
data[data == " "] <- NA
```

returning Sub-Strings of **"years old"** from Column "Custom *Age" && **"time" , "times", " or..." f**rom Column "Physically\_* attacked", "Physical_fighting", "Close_friends" && **"day", "days", " \-\--to....", "\--or..."** from the column Miss_school_no_permission.

Using both str_sub and sub

```{r}
data$Custom_Age <- str_sub(data$Custom_Age,1,2)
data$Physically_attacked <- sub(" .*", "", data$Physically_attacked)
data$Physical_fighting <- sub(" .*","",data$Physical_fighting)
data$Close_friends <- str_sub(data$Close_friends,1,2)
data$Miss_school_no_permission <- sub(" .*","",data$Miss_school_no_permission)
```

Columns Such as "Felt_lonely", "Other_students_kind_and_helpful", "Parents_understand_problems" have 5 level of text responses which are replaced into numeric as follows:

1\) "Never" \<-1,

2\) "Rarely" \<- 2,

3\) "Sometimes" \<- 3,

4\) "Most of the time" \<- 4,

5\) "Always" \<- 5.

```{r}
data[data == "Never"] <- 1
data[data == "Rarely"] <- 2
data[data == "Sometimes"] <- 3
data[data == "Most of the time"] <- 4
data[data == "Always"] <- 5
```

Replacing ***"YES" with 1*** and ***"No" with 0*** in the columns "Bullied_on_school_property_in_past_12_months" , "Bullied_not_on_school_property_in_past_12_months", "Cyber_bullied_in_past_12_months" , "Most_of_the_time_or_always_felt_lonely", "Missed_classes_or_school_without_permission".

```{r}
data[data == "Yes"] <- 1
data[data == "No"] <- 0
```

Assigning data-types into Factors depending upon the requirement of the model

```{r}
data$Bullied_on_school_property_in_past_12_months <- as.factor(data$Bullied_on_school_property_in_past_12_months)
data$Bullied_not_on_school_property_in_past_12_months <- as.factor(data$Bullied_not_on_school_property_in_past_12_months)
data$Cyber_bullied_in_past_12_months <- as.factor(data$Cyber_bullied_in_past_12_months)
data$Sex <- as.factor(data$Sex)
data$Felt_lonely <- as.factor(data$Felt_lonely)
data$Other_students_kind_and_helpful <- as.factor(data$Other_students_kind_and_helpful)
data$Parents_understand_problems <- as.factor(data$Parents_understand_problems)
data$Most_of_the_time_or_always_felt_lonely <- as.factor(data$Most_of_the_time_or_always_felt_lonely)
data$Missed_classes_or_school_without_permission <- as.factor(data$Missed_classes_or_school_without_permission)

```

Assigning the data-types into integers depending upon the requirement of the model.

```{r}
data$Custom_Age <- as.integer(data$Custom_Age)
data$Physically_attacked <- as.integer(data$Physically_attacked)
data$Physical_fighting <- as.integer(data$Physical_fighting)
data$Close_friends <- as.integer(data$Close_friends)
data$Miss_school_no_permission <- as.integer(data$Miss_school_no_permission)

```

Deleting the columns which seems irrelevant to the model or they have high number of missing values

```{r}
data <- data[, -c(16:18)]
```

Deleting "NA" from columns 2 to 15

```{r}
data <- data[complete.cases(data),]
```

Coding the Logistic Regression Model

Null Hypothesis :- Factors such as:

1)Custom_Age

2\) Sex

3)Degree to which the student was feeling lonely there are 5 levels in the dataset related to it

4\) Number of Close friends the student has

5\) Number of days the student missed school without permission there are 5 levels mentioned in the dataset related to it

6\) Degree to which other students are kind and helpful

7\) Degree to which parents understand their problems

8\) Whether they felt lonely Most of the time or always.

```{r}
logistic <- glm(Bullied_on_school_property_in_past_12_months ~ Custom_Age + Sex + Felt_lonely + Close_friends + Miss_school_no_permission + Other_students_kind_and_helpful + Parents_understand_problems + Most_of_the_time_or_always_felt_lonely, family = binomial, data = data)
summary(logistic)

```

Model-2 deleting variable Sex from the previous variable

```{r}
logistic_2 <- glm(Bullied_on_school_property_in_past_12_months ~ Custom_Age  + Felt_lonely + Close_friends + Miss_school_no_permission + Other_students_kind_and_helpful + Parents_understand_problems + Most_of_the_time_or_always_felt_lonely, family = binomial, data = data)
summary(logistic_2)
```

Model-3 generating model to learn about the factors affecting cyber bullying

```{r}
logistic3 <- glm(Cyber_bullied_in_past_12_months ~ Custom_Age + Sex + Felt_lonely + Close_friends + Miss_school_no_permission + Other_students_kind_and_helpful + Parents_understand_problems + Most_of_the_time_or_always_felt_lonely, family = binomial, data = data)
summary(logistic3)

```

```{r}
#render("Check_in_1&2.qmd", output_file = "my_document.html", output_format = "html_document")

```
