---
title: "Check-in 2: DACSS 603 Final Project"
author: "Laura Collazo"
desription: ""
date: "2023-4-15"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart2
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(psych)
library(lattice)
library(FSA)
library(kableExtra)
library(lmtest)
library(stargazer)


knitr::opts_chunk$set(echo = TRUE)
```
# Research Question

Many studies exist on how belonging influences in-person university students yet there is an absence of literature related to this topic for online learners. This is surprising given that even before the COVID-19 pandemic students studying exclusively online have experienced year-over-year growth since at least 2016 (National Center for Educational Statistics, n.d.). High levels of social belonging have repeatedly being found to influence the persistence of in-person students (Fleming et al., 2017; Johnson et al., 2007; Sriram, 2017). With this in mind, the role belonging plays among online students should not be overlooked.

There is no universal definition of belonging, but one way to think about it is a "students' sense of being accepted, valued, included, and encouraged by others (teacher and peers) in the academic classroom setting and of feeling oneself to be an important part of the life and activity of the class" (Goodenow, 1993). In studies of in-person students, it has been found that not all students experience belonging equally. Some of the student populations that report lower levels of belonging include:

* students of color compared to their White peers (Johnson et al., 2007)
* women and students of color studying in STEM (Thacker et al., 2022)
* first-generation students (Longwell-Grice et al., 2016)
* students with disabilities (Fleming et al., 2017).

Experiencing a lack of belonging is not innate to these students, though, but rather a reflection of social structural inequalities present in educational systems (Brock, 2019). This reality makes belonging important to explore among students not just generally, but specifically, and is why this study will ask:

**How does race influence belonging of online undergraduate students?**

# Hypothesis

Literature on how demographics influence the social belonging of online students is elusive making it difficult to develop a hypothesis rooted in previous literature. Therefore, since the data for this study comes from students in an online statistics course, the hypotheses for this study will be based on the notion that underrepresented minorities (URM) in STEM experience a lower sense of belonging that students who are not URMs (Thacker et al., 2022). 

To understand representation in STEM by race it is helpful to examine the data in this table which shows the breakdown of workers in all jobs compared to all STEM jobs (Fry et al., 2021).

```{r}

race <- c("Black", "Hispanic", "Asian", "Other", "White")

all_jobs <- c("11%", "17%", "6%", "3%", "63%")

all_stem_jobs <- c("9%", "8%", "13%", "3%", "67%")

jobs <- data.frame(race, all_jobs, all_stem_jobs)

kable(jobs, caption = "Employed adults age 25+ in the 2017-19 American Community Survey") %>% 
  kable_styling("striped")

```

Employees who are White make up the majority of workers in STEM, yet employees who are Asian hold more than double the percentage of STEM jobs compared to the percentage of all jobs held by workers who are Asian (Fry et al., 2021). For this reason, Asian students will not be included as an underrepresented minority group in STEM. However, it is important to be mindful that many ethnic groups fall under a racial category and not all are equally represented in STEM. For instance, although Asians as a whole racial group are not underrepresented in STEM, those who are from Southeast Asia are underrepresented (Shivaram, 2021). Future studies should be mindful of this and seek to explore for both racial and ethnic demographic data.    

**Hypothesis**

* Students who are part of an underrepresented minority group in STEM (Black or Hispanic) will report lower levels of social belonging than students who are not part of an underrepresented racial minority group in STEM (White or Asian).

# Descriptive Statistics

## Dataset

The data for this study was collected during the last month of the Fall 2021 semester at a large public university in the mid-west of the United States (Jeng et al., 2023). Students were recruited through voluntary response sampling in an online introduction to statistics course and received extra credit for participating in the study. The instructor of this course was not a member of the research team. A total of 240 students completed the survey and responses from 17 students were removed due to either missing demographics, the same rating for every example, or identical response made to more than 50% of open-ended questions. In total, responses from 223 students are included.

The full dataset can be found [here](https://databank.illinois.edu/datasets/IDB-2872989).

### Variables

To answer the research question for the present study, the following variables have been selected. Details on these variables are included in following sections.

* `race`
* `gender`
* `year`
* `bel_1`
* `bel_2`
* `bel_2_r`
* `bel_3`
* `bel_4`
* `bel_4_r`
* `bel_5`
* `bel_5_r`
* `bel_6`

Using the above variables, additional variables were created to aid in analysis. The variable `bel_sum` sums each question in the belonging scale to create an overall belonging score for every participant and the variable `urm` indicates if a student belongs to a represented or underrepresented minority (URM) group in STEM given their reported race.

#### Race/Ethnicity

The response options for `race` include "Asian or Asian American" (1), "Black or African American" (2), "Hispanic or Latino" (3), "White" (4), and "Other" (5). Students who reported their race as "Other" (n = 10) have been removed from this analysis as it cannot be determined if they belong to an underrepresented minority group or not.

#### Gender

The responses for `gender` include "Man" (2) and "Woman or non-binary" (1). The researchers who collected this data explain that their were so few students who identified as non-binary that the sample was too small for a separate analysis (Jeng et al., 2023). What they did do was run all analysis twice, once with non-binary students excluded and once with this group combine with respondents who identified as women. They found their findings to be the same in both instances and therefore choose to combine these two groups. For the purpose of this study, this variable will be coded as "Male" and "Female."

#### Year

The variable `year` includes four response options: 1, 2, 3 or 4. These correspond to being a Freshman, Sophomore, Junior, or Senior, and values have been coded as such.

#### Belonging

A total of 6 Likert questions were asked to measure social belonging which were adapted from Goodenow's (1993) Psychological Sense of School Membership (PSSM) scale. Five response options ("Not at all true", "Slightly true", "Moderately true", "Mostly true", and "Completely true") were provided to students to answer the below questions. Questions 2, 4, and 5 were reversed scored which is why 2 variables exist for each of these questions.

1. I feel like a real part of this class (`bel_1`)
2. Sometimes I feel as if I don’t belong in this class(`bel_2`/`bel_2_r`)
3. I am included in lots of activities in this class(`bel_3`)
4. I feel very different from most other students in this class (`bel_4`/ `bel_4_r`)
5. I wish I were in a different class (`bel_5`/`bel_5_r`)
6. I feel proud of belonging to this class (`bel_6`)

#### Explore Data

##### Prepare Data

```{r tidy data, echo=TRUE}

# read in data

belonging_data <- read_csv("_data/belonging_survey_2022-07-08.csv", show_col_types = FALSE)

# tidy data

data <- belonging_data %>%
  
# filter to include only 1 set of student belonging responses
  
  filter(example_num == 1) %>% 
  
# filter to include remove the racial category "Other"
  
  filter(race != 5) %>% 
  
# select needed columns
  
  select(race, gender, year, bel_1, bel_2, bel_2_r, bel_3, bel_4, bel_4_r, bel_5, bel_5_r, bel_6) %>%

# create variable bel_sum which sums all belonging responses
  
  mutate(bel_sum = rowSums(across(c(bel_1, bel_2_r, bel_3, bel_4_r, bel_5_r, bel_6)))) %>% 

# create variable is_urm which turns race into a binary variable

  mutate(urm = case_when(race == 1 ~ "not underrepresented",
                         race == 4 ~ "not underrepresented",
                         race == 2 ~ "underrepresented",
                         race == 3 ~ "underrepresented",))

### Create new variables with Likert scores as an ordered factor

data$f_bel_1 = factor(data$bel_1,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_2 = factor(data$bel_2,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_3 = factor(data$bel_3,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_4 = factor(data$bel_4,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_5 = factor(data$bel_5,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_6 = factor(data$bel_6,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

# recode data
  
data <- data %>% 
  mutate(gender = recode(gender,
                            `1` = "Female",
                            `2` = "Male")) %>%
  mutate(race = recode(race,
                            `1` = "Asian",
                            `2` = "Black",
                            `3` = "Hispanic",
                            `4` = "White")) %>%
  mutate(across(bel_1:bel_6, 
                ~ recode(.x, `1` = "Not at all true",
                             `2` = "Slightly true",
                             `3` = "Moderately true",
                             `4` = "Mostly true",
                             `5` = "Completely true"))) %>% 
  mutate(year = recode(year,
                        `1` = "Freshman",
                        `2` = "Sophmore",
                        `3` = "Junior",
                        `4` = "Senior")) 

```

##### Examine Data

```{r examine data, echo=TRUE}

# examine data

describe_data <- describe(x=data) %>% 
  select(c(vars, n, mean, sd, median, min, max, range))

kable(describe_data) %>% 
  kable_styling("striped")

str(data)

```

##### Examine Variables

```{r visualizing variables}

# create xtabs and bar plot to visualize variables

# race

xt_race <- xtabs(~race, data = data)

kable(xt_race) %>% 
  kable_styling("striped")

barplot(xt_race, 
        xlab = "Race",
        ylab = "Frequency")

# urm

xt_urm <- xtabs(~urm, data = data)

kable(xt_urm) %>% 
  kable_styling("striped")

barplot(xt_urm, 
        xlab = "URM",
        ylab = "Frequency")

# gender

xt_gender <- xtabs(~gender, data = data)

kable(xt_gender) %>% 
  kable_styling("striped")

barplot(xt_gender, 
        xlab = "Gender",
        ylab = "Frequency")

# bel_sum

xt_sum <- xtabs(~bel_sum, data = data)

kable(xt_sum) %>% 
  kable_styling("striped")

barplot(xt_sum, 
        xlab = "Belonging",
        ylab = "Frequency")

# f_bel_1

xt_1 <- xtabs(~f_bel_1, data = data)

kable(xt_1) %>% 
  kable_styling("striped")

barplot(xt_1, 
        xlab = "I feel like a real part of this class",
        ylab = "Frequency")

# f_bel_2

xt_2 <- xtabs(~f_bel_2, data = data)

kable(xt_2) %>% 
  kable_styling("striped")

barplot(xt_2, 
        xlab = "Sometimes I feel as if I don’t belong in this class",
        ylab = "Frequency")

# f_bel_3

xt_3 <- xtabs(~f_bel_3, data = data)

kable(xt_3) %>% 
  kable_styling("striped")

barplot(xt_3, 
        xlab = "I am included in lots of activities in this class",
        ylab = "Frequency")

# f_bel_4

xt_4 <- xtabs(~f_bel_4, data = data)

kable(xt_4) %>% 
  kable_styling("striped")

barplot(xt_4, 
        xlab = "I feel very different from most other students in this class",
        ylab = "Frequency")

# f_bel_5

xt_5 <- xtabs(~f_bel_5, data = data)

kable(xt_5) %>% 
  kable_styling("striped")

barplot(xt_5, 
        xlab = "I wish I were in a different class",
        ylab = "Frequency")

# f_bel_6

xt_6 <- xtabs(~f_bel_6, data = data)

kable(xt_6) %>% 
  kable_styling("striped")

barplot(xt_6, 
        xlab = "I feel proud of belonging to this class",
        ylab = "Frequency")

```

# Analysis

Before conducting hypothesis testing and fitting regression models, it is helpful to examine key variables through visualization. 

The first visualization is a boxplot which examines the dependent variable `bel_sum` and the independent variable `urm`. It reveals that the mean belonging score is similar for both groups, however the range of responses is greater among students who belong to a URM group. 

```{r  echo=TRUE, warning=FALSE}

# bel_sum, urm

ggplot(data, aes(x = urm, y = bel_sum, color = urm))+
    geom_boxplot()+
    ylab("Belonging")+xlab("URM")

```
The second boxplot adds `gender`, one of the control variables selected for this study. It shows the mean belonging score is similar for females in both groups as well as  males. It can also be observed that males, compared to females, have lower means among both groups and that the range of belonging scores is greater for URM students than non-URM students. 

```{r  echo=TRUE, warning=FALSE}

# bel_sum, urm, gender

ggplot(data, aes(x = urm, y = bel_sum, color = gender))+
    geom_boxplot()+
    ylab("Belonging")+xlab("URM")

```

This final boxplot examines `year`, another control variable. This visualization indicates more variability between the two `urm` groups than seen in previous visualizations. Where students are URMs, the mean belonging score decreases as year in school increases. An additional observation among students who are non-URMs is that the mean belonging score appears to be consistent for students in their 1st, 2nd, and 4th year of studies, but drops for students in their 3rd year. 

```{r  echo=TRUE, warning=FALSE}

# bel_sum, urm, year

data %>% 
  mutate(year = as.factor(year)) %>% 
ggplot(aes(x = urm, y = bel_sum, color = factor(year, levels = c("Freshman", "Sophmore", "Junior", "Senior")))) +
    geom_boxplot() +
    ylab("Belonging") +
      xlab("URM") +
  labs(color = "Year")
```

In the next section of visualizations, stacked bar charts have been selected to examine the explanatory variable `urm` against each of the 6 individual belonging questions which have been selected as response variables (`bel_1`:`bel_6`). A couple initial observations that stand out are:

* There are no non-URM students who responded "Completely true" to the statement "Sometimes I feel as if I don't belong in this class" (`bel_2`).
* A much higher percentage of URM students responded "Completely true" to the statement "I feel very different from most other students in this class" (`bel_4`) than non-URM students.
*  A much higher percentage of URM students responded "Completely true" to the statement "I wish I were in a different class" (`bel_5`) than non-URM students.

Hypothesis testing and regression will determine if these observations, and those mentioned previously, are significant.

```{r, message=FALSE}

bel_plot <-function(myxvar, myxlabel) {
  data %>%
    group_by({{myxvar}}, urm) %>%
  summarise(count = n()) %>%
    mutate(percent = count/sum(count) * 100) %>% 
    ggplot(aes(x = factor ({{myxvar}}, levels = c("Not at all true", "Slightly true", "Moderately true", "Mostly true", "Completely true")),
             y = percent, 
             fill = urm)) + 
    geom_bar(stat = "identity", position = "fill") + 
    labs(y = "Percent", 
         x = myxlabel,
         title = "Sense of belonging by whether students' race is underrepresented in STEM " )
}

bel_plot(bel_1, "I feel like a real part of this class")

bel_plot(bel_2, "Sometimes I feel as if I don’t belong in this class")

bel_plot(bel_3, "I am included in lots of activities in this class")

bel_plot(bel_4, "I feel very different from most other students in this class")

bel_plot(bel_5, "I wish I were in a different class")

bel_plot(bel_6, "I feel proud of belonging to this class")

```

# Hypothesis Testing

## ANOVA

The first hypothesis test conducted is an ANOVA. This test has been selected to compare the means of URM and non-URM students.

```{r}

fit_sum <- aov(bel_sum ~ urm, data = data)

summary(fit_sum)

```

The results of this ANOVA reveal that Pr(>F) is 0.0584. This is just slightly greater than p < .05, therefore the effect size should also be examined before deciding if the null hypothesis should or shouldn't be rejected. Eta squared has been selected to measure the effect size.

```{r}

effectsize::eta_squared(fit_sum)

```
The effect size for the ANOVA between `bel_sum` and `urm` is 0.02 which is small. This indicates the null hypothesis should not be rejected and means the difference in mean of belonging between students in this study who are and are not URM is not statistically significant.

## Chi-squared

The next round of hypothesis testing examines the explanatory outcome variable `urm` against each of the six individual belonging questions that make up the belonging scale. As these are all categorical variables, a chi-squared test will be used.

The first chi-squared test examines `bel_1`, which includes responses to the statement "I feel like a real part of this class." It results in a p-value of 0.6553 which is not statistically significant.

```{r, warning=FALSE}

chi_bel_plot <- function(mydata, myxvar, mytitle) {
  ggplot({{mydata}}, aes(x=factor ({{myxvar}}, levels = c("Not at all true", "Slightly true", "Moderately true", "Mostly true", "Completely true")),y=Freq, fill=urm)) +
  geom_bar(stat="identity",position="dodge") +
  labs(x = mytitle)
}

# chisq and bel_1

chisq.test(data$bel_1, data$urm)

bel_1_urm <- data.frame(with(data, table(bel_1,urm)))

chi_bel_plot(bel_1_urm, bel_1, "I feel like a real part of this class")

```

The second chi-squared test examines `bel_2`, which includes responses to the statement "Sometimes I feel as if I don’t belong in this class." It results in a p-value of 0.01644 which is statistically significant. 

```{r, warning=FALSE}

# chisq and bel_2

chisq.test(data$bel_2, data$urm)

bel_2_urm <- data.frame(with(data, table(bel_2,urm)))

chi_bel_plot(bel_2_urm, bel_2, "Sometimes I feel as if I don’t belong in this class")

```

This chi-squared test examines `bel_3`, which includes responses to the statement "I am included in lots of activities in this class." It results in a p-value of 0.9815 which is not statistically significant. 

```{r, warning=FALSE}

# chisq and bel_3 (SIGNIFICANT P)

chisq.test(data$bel_3, data$urm)

bel_3_urm <- data.frame(with(data, table(bel_3,urm)))

chi_bel_plot(bel_3_urm, bel_3, "I am included in lots of activities in this class")

```

This next chi-squared test examines `bel_4`, which includes responses to the statement "I feel very different from most other students in this class." It results in a p-value of 0.1081 which is not statistically significant.

```{r, warning=FALSE}

# chisq and bel_4

chisq.test(data$bel_4, data$urm)

bel_4_urm <- data.frame(with(data, table(bel_4,urm)))

chi_bel_plot(bel_4_urm, bel_4, "I feel very different from most other students in this class")

```

The following chi-squared test examines `bel_5`, which includes responses to the statement "I wish I were in a different class." It results in a p-value of 0.02648 which is statistically significant.

```{r, warning=FALSE}

# chisq and bel_5 (SIGNIFICANT P)

chisq.test(data$bel_5, data$urm)

bel_5_urm <- data.frame(with(data, table(bel_5,urm)))

chi_bel_plot(bel_5_urm, bel_5, "I wish I were in a different class")

```

This final chi-squared test examines `bel_6`, which includes responses to the statement "I feel proud of belonging to this class." It results in a p-value of 0.2512 which is not statistically significant.

```{r, warning=FALSE}

# chisq and bel_6

chisq.test(data$bel_6, data$urm)

bel_6_urm <- data.frame(with(data, table(bel_6,urm)))

chi_bel_plot(bel_6_urm, bel_6, "I feel proud of belonging to this class")

```

To recap the results of these six chi-squared tests, `bel_2`, "Sometimes I feel as if I don’t belong in this class", and `bel_5`, "I wish I were in a different class", were both statistically significant. This indicates a relationship between each of these two specific measure of belonging and `urm`.

## Linear Regression

In this section regression models will be compared to determine which has the best fit. In addition to the variables used in hypothesis testing, `gender` and `year` will be include as control variables and `urm` * `gender` as an interaction term. This interaction term has been included as previous research indicates the intersection of gender and racial identities can reveal significant differences (Rainey et al., 2018).

The first regression model examines `bel_sum` and `urm`. The adjusted R-squared of this model is 0.01221, and as this is very close to zero it indicates the response variable `bel_sum` is not explained by the predictor variable `urm`. Another way to think of this is that belonging to an URM group only explains 1.22% of the variation in a student's sense of belonging. 

From this model it can be concluded that being part of a URM group leads to a -1.34 lower belonging score compared to those who are non-URM students.

```{r}

# bel_sum & urm

model_urm <- lm(bel_sum ~ urm, data = data)

summary(lm(bel_sum ~ urm, data = data))

```

The second regression model examines `bel_sum`, `urm`, and `gender`. The adjusted R-square is low, 0.03259, which indicates the relationship between these variables is not strong enough to explain the variation between the response and predictor variables. 

From this model it can be concluded that `gender` is statistically significant. Identifying as male leads to a -1.65 lower belonging score compared to females when other variables are held constant.

```{r}

# bel_sum, urm & gender

model_gender <- lm(bel_sum ~ urm + gender, data = data)

summary(lm(bel_sum ~ urm + gender, data = data))

```

The third regression model replaces `gender` with `year`. The adjusted R-squared in this model is still low. This model explains just 5.47% of the variation between the response and predictor variables.

From this model it can be concluded that being a Junior or Senior is statistically significant when compared to Freshman. Seniors have a belonging score -3.2 points lower than Freshman and Juniors have a belonging score -2.6 points lower than Freshman when holding other variables constant. 

```{r}

# bel_sum, urm & year

model_year <- lm(bel_sum ~ urm + year, data = data)

summary(lm(bel_sum ~ urm + year, data = data))

```

This next regression model includes `bel_sum`, `urm`, `gender` and `year`. The adjusted R-squared of this model is higher than others, however it only explains 6.63% of the variation which is still rather low. 

This model also shows that when variables are held constant being a Junior or Senior is statistically significant. Identifying as Male has a p-value of .0576 which is just slightly above p < .05. 

```{r}

# bel_sum, urm, gender & year

model_all <- lm(bel_sum ~ urm + gender + year, data = data)

summary(lm(bel_sum ~ urm + gender + year, data = data))

```

The final regression model using `bel_sum` includes the interaction between `urm` and `gender` as the explanatory variable. The adjusted R-squared of this model is again low explaining just 4.1% of the variation between variables. No variables are significant in this model at the 5% level, however the interaction between being underrepresented and male is significant at the 10% level. 

```{r}

# bel_sum, urm * gender

model_interaction <- lm(bel_sum ~ urm * gender, data = data)

summary(lm(bel_sum ~ urm * gender, data = data))

```

# Model Comparison

To determine which of the five linear regression models has the best fit, AIC and BIC is examined alongside adjusted R-squared.

```{r, warning=FALSE}

models <- list(model_urm, model_gender, model_year, model_all, model_interaction)

column_labels <- c("urm", "urm + gender", "urm + year", "urm + gender + year", "urm * gender" )

stargazer(models, 
          title = "Linear Regression Models", 
          align = TRUE, 
          type = "text", 
          column.labels = column_labels, 
          add.lines = list(c("AIC", 
                             round(AIC(model_urm), 2), 
                             round(AIC(model_gender), 2), 
                             round(AIC(model_year), 2), 
                             round(AIC(model_all), 2), 
                             round(AIC(model_interaction), 2)), 
                           c("BIC", 
                             round(BIC(model_urm), 2), 
                             round(BIC(model_gender), 2), 
                             round(BIC(model_year), 2), 
                             round(BIC(model_all), 2), 
                             round(BIC(model_interaction), 2))))


```

After examining the results, it is clear that model four, `bel_sum ~ urm + gender + year`, is the best fitting model of those tested. This can be inferred as it has both the highest adjusted R-squared and lowest AIC. Although this model does has the highest BIC, this is not surprising as BIC can penalize models with more complexity.    

# Diagnostics

The final step of analysis is performing diagnostics on model four, `bel_sum ~ urm + gender + year`, which has been selected as the best fitting model.

## Residuals vs Fitted

In the Residuals vs Fitted plot both the linearity and constant variance assumptions are violated. These violations are evident as the line is not linear and the points are not equally horizontal around the line at 0. There are also residuals which stand out from the rest indicating outliers.

```{r}

plot(lm(bel_sum ~ urm + gender + year, data = data), which = 1)

```

## Normal Q-Q

In the QQ plot, the residuals fall on the line towards the center but curve towards the ends. This indicates the data isn't part of a normal distribution, therefore the normality assumption has been violated (Ford, 2015). 

```{r}

plot(lm(bel_sum ~ urm + gender + year, data = data), which = 2)

```

## Scale-Location

In the Scale-Location plot the red line is decreasing rather than being horizontal and the variance of points do not appear equal. 

```{r}

plot(lm(bel_sum ~ urm + gender + year, data = data), which = 3)

```

To test for heteroscedasticity a Breusch-Pagan Test is used. This test results in a p-value less than .05 allowing the null hypothesis, that the residuals are homoscedastic, to be rejected. This is a violation of the constant variance assumption.

```{r}

model_all <- lm(bel_sum ~ urm + gender + year, data = data)

# Breusch-Pagan test

bp_test <- bptest(model_all)

bp_test

```

## Cook's distance

The Cook's Distance plot shows a violation of the influential observation assumption. This is because there are points with a Cook's distance larger than 4/n, which in this study is 4/213 = 0.019.

```{r}

plot(lm(bel_sum ~ urm + gender + year, data = data), which = 4)

```

# References

Borck, C.R. (2020). “I belong here.”: Culturally sustaining pedagogical praxes from an alternative high school in Brooklyn. The Urban Review, 52(2), 376-391. [https://doi.org/10.1007/s11256-019-00536-z](https://doi.org/10.1007/s11256-019-00536-z)

Fleming, A.R., Oertle, K.M., Plotner, A.J., & Hakun, J.G. (2017). Influence of social factors on student satisfaction among college students with disabilities. Journal of College Student Development, 58(2), 215-228. [https://doi.org/10.1353/csd.2017.0016](https://doi.org/10.1353/csd.2017.0016)

Ford, C. (2015, August 26). Understanding Q-Q plots. University of Virgina Library. [https://data.library.virginia.edu/understanding-q-q-plots/](https://data.library.virginia.edu/understanding-q-q-plots/)

Fry, R., Kennedy, B., & Funk, C. (2021, April 1). STEM jobs see uneven progress in increasing gender, racial and ethnic diversity. Pew Research Center. [https://www.pewresearch.org/science/2021/04/01/stem-jobs-see-uneven-progress-in-increasing-gender-racial-and-ethnic-diversity](https://www.pewresearch.org/science/2021/04/01/stem-jobs-see-uneven-progress-in-increasing-gender-racial-and-ethnic-diversity) 

Goodenow. (1993). Classroom belonging among early adolescent students: relationships to
motivation and achievement. The Journal of Early Adolescence, 13(1), 21–43.
[https://doi.org/10.1177/0272431693013001002](https://doi.org/10.1177/0272431693013001002)

Jeng, A., Bosch, N., & Perry, M. (2023). Sense of belonging predicts perceived helpfulness in online peer help-giving interactions. The Internet and Higher Education, 57, 100901. [https://doi.org/10.1016/j.iheduc.2022.100901](https://doi.org/10.1016/j.iheduc.2022.100901) 

Johnson, D.R., Soldner, M., Leonard, J. B., Alvarez, P., Inkelas, K. K., Rowan-Kenyon, H., & Longerbeam, S. (2007). Examining sense of belonging among first-year undergraduates from different racial/ethnic groups. Journal of College Student Development, 48(5), 525–542. [https://doi.org/10.1353/csd.2007.0054](https://doi.org/10.1353/csd.2007.0054)

Longwell-Grice, R., Adsitt, N.Z., Mullins, K., & Serrata, W. (2016). The first ones: Three studies on first-generation college students. NACADA Journal, 36(2), 34-46. [https://doi.org/10.12930/NACADA-13-028](https://doi.org/10.12930/NACADA-13-028)

National Center for Educational Statistics. (n.d.). IPEDS Data Explorer.
[https://nces.ed.gov/ipeds/search/ViewTable?tableId=29450](https://nces.ed.gov/ipeds/search/ViewTable?tableId=29450) 

Rainey, K., Dancy, M., Mickelson, R. et al. (2018). Race and gender differences in how sense of belonging influences decisions to major in STEM. International Journal of STEM Education, 5(1), 10-14. [https://doi.org/10.1186/s40594-018-0115-6](https://doi.org/10.1186/s40594-018-0115-6) 

Shivaram, D. (2021, December 12). Southeast Asians are underrepresented in STEM. The label 'Asian' boxes them out more. NPR. [https://www.npr.org/2021/12/12/1054933519/southeast-asian-representation-science](https://www.npr.org/2021/12/12/1054933519/southeast-asian-representation-science)

Sriram, R. (2017). Student affairs by the numbers: Quantitative research and statistics for professionals. Stylus Publishing, LLC.

Thacker, I., Seyranian, V., Madva, A., Duong, N.T., & Beardsley, P. (2022). Social connectedness in physical isolation: Online teaching practices that support under-represented undergraduate students’ feelings of belonging and engagement in STEM. Education Sciences, 12(2), 61-83. [https://doi.org/10.3390/educsci12020061](https://doi.org/10.3390/educsci12020061)
