---
title: "DACSS 603 Final Project"
author: "Laura"
desription: "Something to describe what I did"
date: "08/02/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - final_project
  - LauraCollazo
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(lsr)
library(psych)
library(lattice)
library(FSA)

knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

This document provides yaml header inforamtion you will need to replicate each week to submit your homework or other blog posts. Please observe the following conventions:

-   Save your own copy of this template as a blog post in the `posts` folder, naming it `FirstLast_hwX.qmd`
-   Edit the yaml header to change your author name - use the same name each week
-   include a *description* that is reader friendly
-   update the *category* list to indicate the type of submission, the data used, the main packages or techniques, your name, or any thing else to make your document easy to find
-   edit as a normal qmd/rmd file


```{r}
x <- c(2,3,4,5)
mean(x)
```

## Rendering your post

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. 

:::{.callout-warning}
Be sure that you have moved your `*.qmd` file into the `posts` folder **BEFORE** you render it, so that all files are stored in the correct location.
:::

:::{.callout-important}
Only render a single file - don't try to render the whole website!
:::

:::{.callout-note}
## Pilot Student Blogs

We are piloting a workflow including individual student websites with direted and limited pull requests back to course blogs. Please let us know if you would like to participate.
:::

## Reading in data files

The easiest data source to use - at least initially - is to choose something easily accessible, either from our `_data` folder provided, or from an online source that is publicly available.

:::{.callout-tip}
## Using Other Data
If you would like to use a source that you have access to and it is small enough and you don't mind making it public, you can copy it into the `_data` file and include in your *commit* and *pull request*.

```{r}

belonging_data <- read_csv("_data/belonging_survey_2022-07-08.csv")

# tidy data
data <- belonging_data %>%
  
  # filter one set of student belonging responses
  filter(example_num == 1) %>%
  
  # select relevant columns
  select(year, gender, race, bel_1, bel_2, bel_2_r, bel_3, bel_4, bel_4_r, bel_5, bel_5_r, bel_6) %>%
  
  
  # sum belonging for total score
  mutate(bel_sum = rowSums(across(c(bel_1, bel_2_r, bel_3, bel_4_r, bel_5_r, bel_6)))) %>% 
  
  # code
  
  mutate(uc_year = recode(year,
                          `1` = "Freshman",
                          `2` = "Sophomore",
                          `3` = "Junior",
                          `4` = "Senior")) %>%
  mutate(uc_gender = recode(gender,
                            `1` = "Female",
                            `2` = "Male")) %>%
  mutate(uc_race = recode(race,
                          `1` = "Asian",
                          `2` = "Black",
                          `3` = "Hispanic",
                          `4` = "White",
                          `5` = "Other")) %>%
  mutate(uc_bel_1 = recode(bel_1,
                           `1` = "Not at all true",
                           `2` = "Slightly true",
                           `3` = "Moderately true",
                           `4` = "Mostly true",
                           `5` = "Completely true")) %>%
  mutate(uc_bel_2 = recode(bel_2,
                             `1` = "Not at all true",
                             `2` = "Slightly true",
                             `3` = "Moderately true",
                             `4` = "Mostly true",
                             `5` = "Completely true")) %>% 
  mutate(uc_bel_2_r = recode(bel_2_r,
                             `1` = "Not at all true",
                             `2` = "Slightly true",
                             `3` = "Moderately true",
                             `4` = "Mostly true",
                             `5` = "Completely true")) %>%
  mutate(uc_bel_3 = recode(bel_3,
                           `1` = "Not at all true",
                           `2` = "Slightly true",
                           `3` = "Moderately true",
                           `4` = "Mostly true",
                           `5` = "Completely true")) %>%
   mutate(uc_bel_4 = recode(bel_4,
                             `1` = "Not at all true",
                             `2` = "Slightly true",
                             `3` = "Moderately true",
                             `4` = "Mostly true",
                             `5` = "Completely true")) %>%
  mutate(uc_bel_4_r = recode(bel_4_r,
                             `1` = "Not at all true",
                             `2` = "Slightly true",
                             `3` = "Moderately true",
                             `4` = "Mostly true",
                             `5` = "Completely true")) %>%
    mutate(uc_bel_5 = recode(bel_5,
                             `1` = "Not at all true",
                             `2` = "Slightly true",
                             `3` = "Moderately true",
                             `4` = "Mostly true",
                             `5` = "Completely true")) %>%
  mutate(uc_bel_5_r = recode(bel_5_r,
                             `1` = "Not at all true",
                             `2` = "Slightly true",
                             `3` = "Moderately true",
                             `4` = "Mostly true",
                             `5` = "Completely true")) %>%
  mutate(uc_bel_6 = recode(bel_6,
                           `1` = "Not at all true",
                           `2` = "Slightly true",
                           `3` = "Moderately true",
                           `4` = "Mostly true",
                           `5` = "Completely true"))

### Create a new variable which is the Likert scores as an ordered factor

data$f_bel_1 = factor(data$bel_1,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_2 = factor(data$bel_2,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_3 = factor(data$bel_3,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_4 = factor(data$bel_4,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_5 = factor(data$bel_5,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

data$f_bel_6 = factor(data$bel_6,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4", "5")
                       )

# order year variable

data$f_year = factor(data$year,
                       ordered = TRUE,
                       levels = c("1", "2", "3", "4"))




view(data)
```

```{r}

# examine data

summary(data)

describe(x=data)

headTail(data)

str(data)

```

```{r}

# create xtabs and bar plot for individual variables

# f_bel_1

xt_1 <- xtabs(~f_bel_1, data = data)

xt_1

barplot(xt_1, 
        xlab = "I feel like a real part of this class",
        ylab = "Frequency")

# f_bel_2

xt_2 <- xtabs(~f_bel_2, data = data)

xt_2

barplot(xt_2, 
        xlab = "Sometimes I feel as if I don’t belong in this class",
        ylab = "Frequency")

# f_bel_3

xt_3 <- xtabs(~f_bel_3, data = data)

xt_3

barplot(xt_3, 
        xlab = "I am included in lots of activities in this class",
        ylab = "Frequency")

# f_bel_4

xt_4 <- xtabs(~f_bel_4, data = data)

xt_4

barplot(xt_4, 
        xlab = "I feel very different from most other students in this class",
        ylab = "Frequency")

# f_bel_5

xt_5 <- xtabs(~f_bel_5, data = data)

xt_5

barplot(xt_5, 
        xlab = "I wish I were in a different class",
        ylab = "Frequency")

# f_bel_6

xt_6 <- xtabs(~f_bel_6, data = data)

xt_6

barplot(xt_6, 
        xlab = "I feel proud of belonging to this class",
        ylab = "Frequency")

# f_year

xt_year <- xtabs(~f_year, data = data)

xt_year

barplot(xt_year, 
        xlab = "Year in school",
        ylab = "Frequency")

# f_race

xt_race <- xtabs(~uc_race, data = data)

xt_race

barplot(xt_race, 
        xlab = "Race",
        ylab = "Frequency")

# f_gender

xt_gender <- xtabs(~uc_gender, data = data)

xt_gender

barplot(xt_gender, 
        xlab = "Gender",
        ylab = "Frequency")

```
```{r}

# create box plots

boxplot(data$bel_1,
        ylab="I feel like a real part of this class")

boxplot(data$bel_2_r,
        ylab="REVERSE: Sometimes I feel as if I don’t belong in this class")

boxplot(data$bel_3,
        ylab="I am included in lots of activities in this class")

boxplot(data$bel_4_r,
        ylab= "REVERSE: I feel very different from most other students in this class")

boxplot(data$bel_5_r,
        ylab= "REVERSE: I wish I were in a different class")

boxplot(data$bel_6,
        ylab= "I feel proud of belonging to this class")

```

```{r}

# xtabs and bar plots by race

# bel_1

xt_race_1 = xtabs(~ uc_race + f_bel_1, data=data)

xt_race_1

barplot(xt_race_1,
        beside=TRUE,
        legend=TRUE,
        xlab="I feel like a real part of this class",
        ylab="Frequency")

# bel_2

xt_race_2 = xtabs(~ uc_race + f_bel_2, data=data)

xt_race_2

barplot(xt_race_2,
        beside=TRUE,
        legend=TRUE,
        xlab="Sometimes I feel as if I don’t belong in this class",
        ylab="Frequency")

# bel_3

xt_race_3 = xtabs(~ uc_race + f_bel_3, data=data)

xt_race_3

barplot(xt_race_3,
        beside=TRUE,
        legend=TRUE,
        xlab="I am included in lots of activities in this class",
        ylab="Frequency")

# bel_4

xt_race_4 = xtabs(~ uc_race + f_bel_4, data=data)

xt_race_4

barplot(xt_race_4,
        beside=TRUE,
        legend=TRUE,
        xlab= "I feel very different from most other students in this class",
        ylab="Frequency")

# bel_5

xt_race_5 = xtabs(~ uc_race + f_bel_5, data=data)

xt_race_5

barplot(xt_race_5,
        beside=TRUE,
        legend=TRUE,
        xlab= "I wish I were in a different class",
        ylab="Frequency")

# bel_6

xt_race_6 = xtabs(~ uc_race + f_bel_6, data=data)

xt_race_6

barplot(xt_race_6,
        beside=TRUE,
        legend=TRUE,
        xlab= "I feel proud of belonging to this class",
        ylab="Frequency")


```

```{r}

# xtabs and bar plots by year

# bel_1

xt_year_1 = xtabs(~ f_year + f_bel_1, data=data)

xt_year_1

barplot(xt_year_1,
        beside=TRUE,
        legend=TRUE,
        xlab="I feel like a real part of this class",
        ylab="Frequency")

# bel_2

xt_year_2 = xtabs(~ f_year + f_bel_2, data=data)

xt_year_2

barplot(xt_year_2,
        beside=TRUE,
        legend=TRUE,
        xlab="Sometimes I feel as if I don’t belong in this class",
        ylab="Frequency")

# bel_3

xt_year_3 = xtabs(~ f_year + f_bel_3, data=data)

xt_year_3

barplot(xt_year_3,
        beside=TRUE,
        legend=TRUE,
        xlab="I am included in lots of activities in this class",
        ylab="Frequency")

# bel_4

xt_year_4 = xtabs(~ f_year + f_bel_4, data=data)

xt_year_4

barplot(xt_year_4,
        beside=TRUE,
        legend=TRUE,
        xlab= "I feel very different from most other students in this class",
        ylab="Frequency")

# bel_5

xt_year_5 = xtabs(~ f_year + f_bel_5, data=data)

xt_year_5

barplot(xt_year_5,
        beside=TRUE,
        legend=TRUE,
        xlab= "I wish I were in a different class",
        ylab="Frequency")

# bel_6

xt_year_6 = xtabs(~ f_year + f_bel_6, data=data)

xt_year_6

barplot(xt_year_6,
        beside=TRUE,
        legend=TRUE,
        xlab= "I feel proud of belonging to this class",
        ylab="Frequency")


```

```{r}

# xtabs and bar plots by gender

# bel_1

xt_gender_1 = xtabs(~ uc_gender + f_bel_1, data=data)

xt_gender_1

barplot(xt_gender_1,
        beside=TRUE,
        legend=TRUE,
        xlab="I feel like a real part of this class",
        ylab="Frequency")

# bel_2

xt_gender_2 = xtabs(~ uc_gender + f_bel_2, data=data)

xt_gender_2

barplot(xt_gender_2,
        beside=TRUE,
        legend=TRUE,
        xlab="Sometimes I feel as if I don’t belong in this class",
        ylab="Frequency")

# bel_3

xt_gender_3 = xtabs(~ uc_gender + f_bel_3, data=data)

xt_gender_3

barplot(xt_gender_3,
        beside=TRUE,
        legend=TRUE,
        xlab="I am included in lots of activities in this class",
        ylab="Frequency")

# bel_4

xt_gender_4 = xtabs(~ uc_gender + f_bel_4, data=data)

xt_gender_4

barplot(xt_gender_4,
        beside=TRUE,
        legend=TRUE,
        xlab= "I feel very different from most other students in this class",
        ylab="Frequency")

# bel_5

xt_gender_5 = xtabs(~ uc_gender + f_bel_5, data=data)

xt_gender_5

barplot(xt_gender_5,
        beside=TRUE,
        legend=TRUE,
        xlab= "I wish I were in a different class",
        ylab="Frequency")

# bel_6

xt_gender_6 = xtabs(~ uc_gender + f_bel_6, data=data)

xt_gender_6

barplot(xt_gender_6,
        beside=TRUE,
        legend=TRUE,
        xlab= "I feel proud of belonging to this class",
        ylab="Frequency")


```
```{r}

# histograms

histogram(~ f_bel_1 | uc_gender,
          data=data,
          layout=c(1,2))

histogram(~ f_bel_1 | uc_gender + uc_race,
          data=data,
          layout=c(2,5))

```
```{r}

# boxplots

boxplot(bel_1 ~ uc_gender + uc_race,
        data=data,
        names=c("a.f", "a.m", "b.f", "b.m", "h.f", "h.m", "o.f", "o.m", "w.f", "w.m"),
        ylab="Value",
        xlab="I feel like a real part of this class")

boxplot(f_bel_1 ~ uc_gender + uc_race,
        data=data,
        names=c("a.f", "a.m", "b.f", "b.m", "h.f", "h.m", "o.f", "o.m", "w.f", "w.m"),
        ylab="Value",
        xlab="I feel like a real part of this class")

boxplot(bel_2_r ~ uc_gender + uc_race,
        data=data,
        names=c("a.f", "a.m", "b.f", "b.m", "h.f", "h.m", "o.f", "o.m", "w.f", "w.m"),
        ylab="Value",
        xlab="REVERSE:Sometimes I feel as if I don’t belong in this class")

boxplot(bel_3 ~ uc_gender + uc_race,
        data=data,
        names=c("a.f", "a.m", "b.f", "b.m", "h.f", "h.m", "o.f", "o.m", "w.f", "w.m"),
        ylab="Value",
        xlab="I am included in lots of activities in this class")

boxplot(bel_4_r ~ uc_gender + uc_race,
        data=data,
        names=c("a.f", "a.m", "b.f", "b.m", "h.f", "h.m", "o.f", "o.m", "w.f", "w.m"),
        ylab="Value",
        xlab= "REVERSE: I feel very different from most other students in this class")

boxplot(bel_5_r ~ uc_gender + uc_race,
        data=data,
        names=c("a.f", "a.m", "b.f", "b.m", "h.f", "h.m", "o.f", "o.m", "w.f", "w.m"),
        ylab="Value",
        xlab= "REVERSE: I wish I were in a different class")

boxplot(bel_6 ~ uc_gender + uc_race,
        data=data,
        names=c("a.f", "a.m", "b.f", "b.m", "h.f", "h.m", "o.f", "o.m", "w.f", "w.m"),
        ylab="Value",
        xlab= "I feel proud of belonging to this class")

```


```{r}
# plots with median and quartile

# bel_1

sum_bel_1 = Summarize(bel_1 ~ uc_race + uc_gender,
                data=data,
                digits=3)

sum_bel_1

ggplot(sum_bel_1, aes(x = uc_race,
                y = median,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: I feel like a real part of this class") +
    xlab("Race")

ggplot(sum_bel_1, aes(x = uc_race,
                y = mean,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("mean: I feel like a real part of this class") +
    xlab("Race")

# bel_2

sum_bel_2 = Summarize(bel_2_r ~ uc_race + uc_gender,
                data=data,
                digits=3)

sum_bel_2

ggplot(sum_bel_2, aes(x = uc_race,
                y = median,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: Reverse-Sometimes I feel as if I don't belong in this class") +
    xlab("Race")

ggplot(sum_bel_2, aes(x = uc_race,
                y = mean,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean: Reverse-Sometimes I feel as if I don't belong in this class") +
    xlab("Race")

# bel_3

sum_bel_3 = Summarize(bel_3 ~ uc_race + uc_gender,
                data=data,
                digits=3)

sum_bel_3

ggplot(sum_bel_3, aes(x = uc_race,
                y = median,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: I am included in lots of activities in this class") +
    xlab("Race")

ggplot(sum_bel_3, aes(x = uc_race,
                y = mean,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean: I am included in lots of activities in this class") +
    xlab("Race")

# bel_4

sum_bel_4 = Summarize(bel_4_r ~ uc_race + uc_gender,
                data=data,
                digits=3)

sum_bel_4

ggplot(sum_bel_4, aes(x = uc_race,
                y = median,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: Reverse-I feel very different from most other students in this class") +
    xlab("Race")

ggplot(sum_bel_4, aes(x = uc_race,
                y = mean,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean: Reverse-I feel very different from most other students in this class") +
    xlab("Race")

# bel_5

sum_bel_5 = Summarize(bel_5_r ~ uc_race + uc_gender,
                data=data,
                digits=3)

sum_bel_5

ggplot(sum_bel_5, aes(x = uc_race,
                y = median,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: Reverse-I wish I were in a different class") +
    xlab("Race")

ggplot(sum_bel_5, aes(x = uc_race,
                y = mean,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean: Reverse-I wish I were in a different class") +
    xlab("Race")

# bel_6

sum_bel_6 = Summarize(bel_6 ~ uc_race + uc_gender,
                data=data,
                digits=3)

sum_bel_6

ggplot(sum_bel_6, aes(x = uc_race,
                y = median,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: I feel proud of belonging to this class") +
    xlab("Race")

ggplot(sum_bel_6, aes(x = uc_race,
                y = mean,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean: I feel proud of belonging to this class") +
    xlab("Race")

# sum of bel

sum_bel_sum = Summarize(bel_sum ~ uc_race + uc_gender,
                data=data,
                digits=3)

sum_bel_sum

ggplot(sum_bel_sum, aes(x = uc_race,
                y = median,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: Total Belonging Score") +
    xlab("Race")

ggplot(sum_bel_sum, aes(x = uc_race,
                y = mean,
                color = uc_gender)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean: Total Belonging Score") +
    xlab("Race")

```

```{r}

# sum of bel

sum_bel_sum_race = Summarize(bel_sum ~ uc_race,
                data=data,
                digits=3)

sum_bel_sum_race

ggplot(sum_bel_sum_race, aes(x = uc_race,
                y = median)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: Total Belonging Score") +
    xlab("Race")

ggplot(sum_bel_sum_race, aes(x = uc_race,
                y = mean)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean: Total Belonging Score") +
    xlab("Race")

sum_bel_sum_gender = Summarize(bel_sum ~ uc_gender,
                data=data,
                digits=3)

sum_bel_sum_gender

ggplot(sum_bel_sum_gender, aes(x = uc_gender,
                y = median)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Median: Total Belonging Score") +
    xlab("Race")

ggplot(sum_bel_sum_gender, aes(x = uc_gender,
                y = mean)) +
    geom_errorbar(aes(ymin = Q1,
                      ymax = Q3),
                   width=.2, size=0.7, position = (position_dodge(.2))) +
    geom_point(shape=15, size=4, position = (position_dodge(.2))) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean: Total Belonging Score") +
    xlab("Race")

```






