---
title: "DACSS 603 Homework 2"
author: "Laura Collazo"
description: "Homework 2 for DASS 603."
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - desriptive statistics 
  - probability
---

# Libraries

```{r, echo=T, message=FALSE}

library(tidyverse)
library(kableExtra)
library(readxl)
library(lsr)

```

# Question 1

The data for the first question is shown below.

```{r, echo=T}

surgical_procedure <- c("Bypass", "Angiography")
sample_size <- c(539, 847)
mean_wait_time <- c(19, 18)
standard_deviation <- c(10, 9)

data <- data.frame(surgical_procedure, sample_size, mean_wait_time, standard_deviation)

kable(data) %>% 
  kable_styling()

confidence_level <- .90

```

The 90% confidence interval for the actual mean wait time for patients receiving a bypass is:

```{r, echo=T}

# calculate confidence interval for "Bypass"

bypass_se <- 10 / sqrt(539)

bypass_tail <- (1-confidence_level)/2

bypass_t <- qt(p = 1 - bypass_tail, df = 539 -1)

bypass_ci <- c(19 - bypass_t * bypass_se,
        19 + bypass_t * bypass_se)

bypass_ci

```

The 90% confidence interval for the actual mean wait time for patients receiving an angiography is:

```{r, echo=T}

# calculate confidence interval for "Angiography"

angi_se <- 9 / sqrt(847)

angi_tail <- (1-confidence_level)/2

angi_t <- qt(p = 1 - angi_tail, df = 847 -1)

angi_ci <- c(18 - angi_t * angi_se,
        18 + angi_t * angi_se)

angi_ci

```

The confidence interval is narrower for the angiography.

# Question 2

The data for question two is shown below.

```{r, echo=T}

sample_size2 <- 1031
voted_yes <- 567
voted_no <- (sample_size2 - voted_yes)

data2 <- data.frame(sample_size2, voted_yes, voted_no)

kable(data2) %>% 
  kable_styling()

```

The point estimate, p, for the proportion of all adult Americans who believe a college education is essential for success is:

```{r, echo=T}

p <- voted_yes/sample_size2

p

```

A 95% confidence interval for p is:

```{r, echo=T}

margin_of_error <- qnorm(0.975) * sqrt(p*(1-p)/sample_size2)

low <- p - margin_of_error

low

high <- p + margin_of_error

high

```
What this tells us is at the 95% confidence interval between 51.95839% and 58.03191% of the population of adult Americans would believe a college education is essential for success.

# Question 3

```{r, echo=T}

```

# Question 4

## A

The below shows the results of a one-sample t-test which tests whether the mean income of female employees ($410) differs from the mean income of all senior-level workers ($500).

One-sample t-tests assume that the population distribution is normal and also that the observations in the sample are generated independently of one another.

```{r, echo=T}

set.seed(27)

rnorm_fixed <- function(n, mean, sd) {
  as.vector(mean + sd * scale(rnorm(n)))
}

data4 <- rnorm_fixed(9, 410, 90)

oneSampleTTest(x=data4, mu=500)

```

The results of this T-test is statistically significant as the p-value is below .05. This means we can reject the null hypothesis.

## B

The p-value when the mean is less than 500 is below.

```{r, echo=T}

t.test(data4, mu = 500, alternative = "less")

```

The results of this t-test are significant as the p-value is below .05. This means we can reject the null hypothesis that the true mean is not less than 500.

## C

The p-value when the mean is greater than 500 is below.

```{r, echo=T}

t.test(data4, mu = 500, alternative = "greater")

```

The results of this test are not significant as the p-value is greater than .05. This means we cannot reject the null hypothesis that the true mean is not greater than 500.

# Question 5

## A

```{r, echo=T}

```

## B

```{r, echo=T}

```

## C

```{r, echo=T}

```

# Question 6

The data for question 6 is shown below. A response of yes indicates a healthy snack was chosen after school.

```{r, echo=T}

data6 <- read_xlsx("homework_2_data.xlsx")

table(data6$healthy_snack, data6$grade)

```

The null hypothesis for question 6 is that middle school students choosing a healthy snack is independent of grade level. The test that should be used to determine if we can reject the null hypothesis is a chi-square test. This test has been chosen as it's used to determine when there's an association between two categorical values. The results of the test are below.

```{r, echo=T}

chisq.test(data6$healthy_snack, data6$grade, correct = FALSE)

table <- data.frame((with(data6, table(healthy_snack, grade))))

ggplot(table, aes(x=healthy_snack,y=Freq, fill=grade))+
  geom_bar(stat="identity",position="dodge")

```

We can reject the null hypothesis because the p-value is below 0.05. This tells us that choosing a healthy snack is not independent of grade level.

Furthermore, the X-squared value, as seen below, is higher than the critical value so the null hypothesis can be rejected. 

```{r, echo=T}

qchisq(.05, 2, lower.tail = FALSE)

```

# Question 7

The data for question 7 is shown below. Each tuition value is the per-pupil cost in thousands of dollars for cyber charter school tuition.

```{r, echo=T}

data7 <- read_xlsx("homework_2_data.xlsx", sheet = "Question 7")

kable(data7) %>% 
  kable_styling()

```

The null hypothesis for question 7 is that the mean cyber charter school tuition cost per-pupil is the same in all areas. The test that should be used to determine if we can reject the null hypothesis is an ANOVA. This test has been chosen as it's used to compare the means of two or more independent groups. The results of the test are shown below.

```{r, echo=T}

ANOVA <- aov(tuition ~ area, data = data7)

summary(ANOVA)

ggplot(data7, aes(x = area, y = tuition, color=area))+
  geom_boxplot() +
  ylab("tuition") +
  xlab("area")

```

We can reject the null hypothesis because the p-value, Pr(>F), is less than .05. This tells us that at least one area has a tuition mean that is different from the others. 
