---
title: "DACSS 603 Homework 2"
author: "Laura Collazo"
description: "Homework 2 for DASS 603."
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - desriptive statistics 
  - probability
---

# Libraries

```{r, echo=T, message=FALSE}

library(tidyverse)
library(kableExtra)

```

# Question 1

The data for the first question is shown below.

```{r, echo=T}

surgical_procedure <- c("Bypass", "Angiography")
sample_size <- c(539, 847)
mean_wait_time <- c(19, 18)
standard_deviation <- c(10, 9)

data <- data.frame(surgical_procedure, sample_size, mean_wait_time, standard_deviation)

kable(data) %>% 
  kable_styling()

confidence_level <- 0.90

```

The 90% confidence interval for the actual mean wait time for patients receiving a bypass is:

```{r, echo=T}
# calculate confidence interval for "Bypass"

bypass_se <- 10 / sqrt(539)

bypass_tail <- (1-confidence_level)/2

bypass_t <- qt(p = 1 - bypass_tail, df = 539 -1)

bypass_ci <- c(19 - bypass_t * bypass_se,
        19 + bypass_t * bypass_se)

bypass_ci

```

The 90% confidence interval for the actual mean wait time for patients receiving an angiography is:

```{r, echo=T}

# calculate confidence interval for "Angiography"

angi_se <- 9 / sqrt(847)

angi_tail <- (1-confidence_level)/2

angi_t <- qt(p = 1 - angi_tail, df = 847 -1)

angi_ci <- c(18 - angi_t * angi_se,
        18 + angi_t * angi_se)

angi_ci

```

The confidence interval is narrower for the angiography.

# Question 2

The data for question two is shown below.

```{r, echo=T}

sample_size2 <- 1031
voted_yes <- 567
voted_no <- (sample_size2 - voted_yes)

data2 <- data.frame(sample_size2, voted_yes, voted_no)

kable(data2) %>% 
  kable_styling()

```

The point estimate, p, for the proportion of all adult Americans who believe a college education is essential for success is:

```{r, echo=T}

p <- voted_yes/sample_size2

p

```

A 95% confidence interval for p is:

```{r, echo=T}

margin_of_error <- qnorm(0.975) * sqrt(p*(1-p)/sample_size2)

low <- p - margin_of_error

low

high <- p + margin_of_error

high

```
What this tells us is at the 95% confidence interval between 51.95839% and 58.03191% of the population of adult Americans would believe a college education is essential for success.


# Question 3

```{r, echo=T}

```

# Question 4

```{r, echo=T}

```

# Question 5

```{r, echo=T}

```

# Question 6

```{r, echo=T}



```

# Question 7

```{r, echo=T}

```
