---
title: "Final Project check-in (1) "
author: "Diana Rinker"
description: "Final project (1)"
date: "3/27/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
editor: 
  markdown: 
    wrap: 72
---

## DACSS 603, spring 2023

## Final Project check-in (1), Diana Rinker.

# Online engagement "


It is well known that online engagement with the web resource is a 
highly valuable metric and is driving site revenue. However, engagement
and popularity might also be associated with other factors, that
websties are trying to avoid, such as online violence, inappropriate
behavior and misinformation.This research project is exploring whith factos impact readers engagement
in social media conversations.

To do that I will use the data from an online blog on the news
website. The author of this blog is posting weekly articles about
interpersonal relationships, that are formulated as a letter to to the
author, where the author gives an advice about the situation. Readers
are free to comment under each post, but cannot make their own posts.

Using this data set, I will explore how engagement of readers of a news
blog connected with variety of factors, such blogs's author engagement ,
topic of the post, source of readers and readers inappropriate online
behavior.

My research question is: Does the authors engagement in the conversation
around the post makes readers more engaged and promotes positive
interactions among them?

**DV:** My dependent construct is **"user's engagement**", I will
measure users' engagement at the level of individual post, using the
following variables

L1 - page viewers View page

L2 - page readers \* Reveal letter \* Reveal comments

L3 - logged in \* Login / sign up \* Up / down vote

L4 - commenter Comment

**IV:** My main independent variable is **Blog's author engagement.** I
will measure authors engagement as the factor variable, with the
following levels:

A. Unspecified comment

B. Featured comment

C. Engagement in conversation

To control for **confounders,** I will also measure the follwing
variables:

1.  Topic of the post ("post tag"), categorical variable.

2.  Source of the readers, also categorical variable.

3.  Mood of the conversation , derivative continuous variable calculated
    as the ratio of "likes" to "dislikes".

4.  Blocked and flagged comments.

```{r, echo=F}
#Loading necessary libraries: 
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

# Data

The data for analysis contains comments of all posts for 2022.

```{r, echo=T}
getwd()
raw <- as_tibble (read_csv("C:\\Users\\Diana\\OneDrive - University Of Massachusetts Medical School\\Documents\\R\\R working directory\\DACSS\\603\\my study files for dacss603\\globe\\LL Comment Data\\iaexport03242023nb_output.csv"))
comments.data<-raw 
colnames (comments.data)
head(comments.data$created_at)
head(comments.data$written_at)
comments.data <-comments.data%>%
              mutate(com.year = format(written_at,format = "%Y" ))
head(comments.data$com.year)
dim(comments.data)

comments.2022 <-comments.data  %>%
              filter(com.year =="2022" )
dim(comments.2022)


```

To answer my research question I will need three datasets that are
coming from a different sources.

-   The first dataset contains comments and their attributes.

-   The second dataset contains posts and their attributes

-   The third dataset contains website analytics, such as  page views, scrolls,
    viewers' sources and so on.

To get better understanding of my data, I will review variables:

### Number of comments, distributed by month

```{r, echo=T}

grouped<-comments.2022 %>%
  group_by (post_id)%>%
  summarise(n.comments=n(), 
            post.month = format(first(written_at),format = "%m" ))
grouped

ggplot(grouped, mapping = aes(x=post.month, y=n.comments, fill=post.month))+
  geom_boxplot() +
  labs(title = "Density distribution of comments per post ", y = "Number of comments" )+
  scale_y_continuous(breaks = seq (from=0, to= 10000, by= 100)) 
   # coord_flip()

```

This graph demonstrates significant variance in amount or comments per post  over 12 months of 2022. 



### Dependent variable. User engagement.

My dependent construct is **user's engagement**, I will measure users
engagement at the level of individual post.

Engagement metrics

L1 - page viewers View page

L2 - page readers

\* Reveal letter

\* Reveal comments

L3 - logged in

\* Login / sign up

\* Up / down vote

L4 - commenter Comment

### Mood of the post.

is a numerical variable.Since each comment has a certain number of "thumbs up" or "thumbs down". I will use this variable to calculate overall "mood" of the post.

```{r, echo=T}

# grouped<-comments.2022 %>%
#   group_by (post_id)%>%
#   summarise(n.comments=n(), 
#             post.month = format(first(written_at),format = "%m" ),
#             likes.sum = sum(absolute_likes), 
#             dislikes.sum=sum (absolute_dislikes))
# grouped <-  grouped%>%
#    mutate(mood =  likes.sum/dislikes.sum)  
#  grouped <- grouped %>%
#   mutate(mood = ifelse(is.nan(likes.sum/dislikes.sum), 0, likes.sum/dislikes.sum))
#  
 grouped<-comments.2022 %>%
      group_by (post_id)%>%
      summarise(n.comments=n(), 
            post.month = format(first(written_at),format = "%m" ),
            likes.sum = sum(absolute_likes), 
            dislikes.sum=sum (absolute_dislikes),
            blocked.sum= sum(approved=="blocked"),
            pct.positive =(sum(absolute_likes)/(sum(absolute_likes)+sum(absolute_dislikes)))*100,
            pct.negative =-1((sum(absolute_dislikes)/(sum(absolute_likes)+sum(absolute_dislikes)))*100)               
  )

ggplot(grouped, mapping = aes( x=pct.positive))+
  geom_baoxplot()  +
  labs(title = "Overall mood distribution" )

```
Than I visualize distribution of mood by months, nd overal variability of a mood score: 
```{r, echo=T}  
  
ggplot(grouped, mapping = aes(x= post.month, y=mood.score))+
  geom_boxplot()  +
  labs(title = "Distribution of mood by month ", y = "Post mood" )

ggplot(grouped, mapping = aes( y=mood.score))+
  geom_boxplot()  +
  labs(title = "Overall mood distribution " )
```
This is  showing that there are 







We can also check the connection between positive and negative sentiment:

```{r, echo=T} 

x.lm<-lm(pct.positive ~ pct.negative, data = grouped)
plot ( data = grouped, pct.negative ~ pct.positive )
```
It falls perfectly on  linear regression line, with positive and negative values strongly negatively correlated. 



### Blocked comments per post.
Now I will visualize amount of blocked comments per post: 

```{r, echo=T}
colnames(comments.2022)

ggplot(grouped, mapping = aes(x= post.month, y=blocked.sum))+
  geom_boxplot() +
 labs(title = "Blockeed comments per month")
```


### Blocked comments and Post mood: 
```{r, echo=T}

plot(blocked.sum ~ mood.score, data= grouped )
  
  
summary(grouped)  

fit <- lm(blocked.sum ~ mood.score, data = grouped)

sum(is.na(grouped$mood.score))

summary(fit)

  
```

### "created_at"

This variable indicates the date of the comment. Using the range of the
dates per post, I can estimate how long each post was in active
discussion. Later I can compare it with the views of the same post
coming from the third dataset.

```{r, echo=T}
# MG.comments<-comments.2022[grep("MeredithGoldstein", comments.2022$user_name, ignore.case = TRUE), ]

comments.2022$user_name<-  ifelse (is.na(comments.2022$user_name), 0, comments.2022$user_name)
comments.2022$mg.comment<-  ifelse (comments.2022$user_name=="MeredithGoldstein", 1, 0)

dim(comments.2022)
grouped<-comments.2022 %>%
  group_by (post_id)%>%
  summarise(n.comments=n(), 
            post.month = format(first(written_at),format = "%m" ),
            likes.sum = sum(absolute_likes), 
            dislikes.sum=sum (absolute_dislikes),
            blocked.sum= sum(approved=="blocked"), 
            pct.positive =(sum(absolute_likes)/(sum(absolute_likes)+sum(absolute_dislikes)))*100,
            pct.negative =(sum(absolute_dislikes)/(sum(absolute_likes)+sum(absolute_dislikes)))*100,
            mg.post =sum(mg.comment)               
  )
grouped<-grouped%>%
  mutate(mood.score = pct.positive/pct.negative)  
grouped

class(grouped$mg.post)

ggplot(grouped, mapping = aes(mg.post))+
  geom_boxplot()  +
  labs(title = "Authors comments ", y = "comments per post .per month" )

grouped$mg.com<-  ifelse (grouped$mg.post>0, 1, 0)

grouped$mg.com <- as.factor(grouped$mg.com)
ggplot(grouped, mapping = aes(y=mood.score, x=mg.com))+
  geom_boxplot() 

```

### "post_id"

Variable allowing me to connect this dataset with another one with the
post informnation,
