---
title: "Final_checkIn"
author: "Sai Padma  pothula"
desription: ""
date: "05/02/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Final checkIn 1
  - sai Pothula
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```


The purpose of this statistical research is to explore the relationship between mental health and academic performance among university students. The dataset was collected through a survey conducted, which asked students about their mental health and current academic situation. The survey collected data on a variety of variables, including students' grades, their reported mental health status, their level of stress, and their use of mental health resources.The dataset includes demographic information about the students, such as their age, gender, and ethnicity, as well as information about their academic performance, such as their CGPA. It also includes information about their mental health, such as whether they have ever been diagnosed with a mental health condition, whether they are currently experiencing symptoms of anxiety or depression, and whether they have sought out mental health resources in the past.

Research Questions

1.Is there a significant correlation between mental health status and CGPA among university students?
       2.How does the use of mental health resources, such as counselling or therapy, have a positive impact on academic performance?
       3.Why there are demographic differences in the relationship between mental health and academic performance, such as differences based on gender or ethnicity?

Hypothesis 	


There is a significant negative correlation between poor mental health status and CGPA among university students. Since CGPA is used as academic performance.

Demographic factors such as gender, ethnicity, and CGPA, can impact the relationship between mental health status and academic performance, with certain groups being more vulnerable to poor mental health outcomes.

Descriptive statistics        

This dataset has 101 rows, 11 columns and 44 missing values 

timestamp: indicates the date and time when the survey was completed by the respondent.

Choose your gender : is a categorical variable that indicates the respondent's gender.

Age: is a numerical variable that indicates the respondent's age in years.

What is your course: is a categorical variable that indicates the respondent's academic program or field of study.

Your current year of Study: is a categorical variable that indicates the respondent's current year of study in their academic program.

What is your CGPA: is a numerical variable that indicates the respondent's cumulative grade point average.

Marital status: is a categorical variable that indicates the respondent's current marital status.

Do you have Depression: is a categorical variable that indicates whether the respondent has ever been diagnosed with or experienced symptoms of depression.

Do you have Anxiety: is a categorical variable that indicates whether the respondent has ever been diagnosed with or experienced symptoms of anxiety.
Do you have Panic attack: is a categorical variable that indicates whether the respondent has ever been diagnosed with or experienced symptoms of panic attacks.

Did you seek any specialist for treatment: is a categorical variable that indicates whether the respondent has sought out any specialist for treatment related to their mental health issues. The variable includes a brief description of the types of specialists sought out by the respondents.


```{r}
library(tidyverse)
library(readr)
library(ggplot2)
```

```{r}
health <- read_csv("~/Desktop/practice/601/603_Spring_2023/posts/_data/student_mental.csv", show_col_types = FALSE)
```
```{r}
colnames(health)
```

```{r}
colnames(health)
```
```{r}
glimpse(health)


```
```{r}
head(health)
```

```{r}

```



