---
title: "Final Project Part 2"
author: "Alexa Potter"
description: "DACSS 603"
date: "04/20/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart2
  - nyctreecensus
---


# Research Question  


Greenery is widely recognized as a vital element to any public space. Plants and natural elements of all kinds can add immense benefits to urban areas, not only to the environment but to the humans who inhabit the space as well. Studies have shown that urban greening, specifically the planting of trees, can "combat challenges such as pollution, urban heat, and flooding, as well as to improve social cohesion, human health, and well-being."^[Cavender, N., & Donnelly, G. (2019). Intersecting Urban Forestry and botanical gardens to address big challenges for healthier trees, people, and cities. *PLANTS, PEOPLE, PLANET, 1*(4), 315–322. https://doi.org/10.1002/ppp3.38] The understanding and implementation of this information can lead cities, developers, and anyone with a stake in their community to consciously consider what elements they can incorporate into their own public spaces. The work does not end here though, greenery in public spaces requires maintenance and year-round management to make these efforts last long enough to see the benefits.


New York City, with a population of about 8.4 billion people and 300 square miles, is one of the largest urban spaces in the United States.^[U.S. Census Bureau . (n.d.). *U.S. Census Bureau quickfacts: New York City, New York*. Retrieved March 20, 2023, from https://www.census.gov/quickfacts/newyorkcitynewyork] In 1995, the New York City Department of Parks and Recreation conducted a city wide census of all the trees. They again conducted this survey in 2005 and 2015 to tackle their goal of enhancing and restoring urban forests.^[Merriman, D. (2017) *Volunteers count every street tree in New York City*. US Forest Service. Retrieved March 20, 2023, from https://www.fs.usda.gov/features/volunteers-count-every-street-tree-new-york-city-0] The information this survey collected has been used to create an interactive map of tree species around New York City. The Parks department then uses this data to calculate the related impacts and needs associated with the trees and tree maintenance.^[Cochran, C., & Greer, B. (2016, June 29). *Treescount! 2015: NYC's Third Street-Tree Census*. New York State Urban Forestry Council. Retrieved March 20, 2023, from https://nysufc.org/treescount/2016/04/26/]

While there is a large amount of data related to the tree census published, there is a gap in information on the relation between tree data to specific neighborhoods.  Critically analyzing this tree data on a neighborhood level can lead to further community involvement within and stemming from neighborhoods themselves.^[Ma, Q., Lin, J., Ju, Y. et al. Individual structure mapping over six million trees for New York City USA. *Sci Data 10*, 102 (2023). https://doi.org/10.1038/s41597-023-02000-w] Firsthand involvement can be used to develop connections between community members and foster ownership among members with the environment they inhabit.  

This informs my research question: 

#### What are the significant differences in tree characteristics across New York City neighborhoods? 


# Hypothesis  

New York City itself is a diverse landscape with features both conducive and preventive to tree growth and sustainability. Neighborhoods across the city can see significant changes in income, traffic, infrastructure, natural resources such water or sunlight, and attitudes towards the environment.^[Neckerman, K., Lovasi, G., Davies, S. et al. Disparities in Urban Neighborhood Conditions: Evidence from GIS Measures and Field Observation in New York City. *Public Health Pol 30* (Suppl 1), S264–S285 (2009). https://doi.org/10.1057/jphp.2008.47] Tree size can relay information to researchers on the development and sustainability of their goal to promote urban greening.^[Colleen E. Reid, Laura D. Kubzansky, Jiayue Li, Jessie L. Shmool, Jane E. Clougherty. It's not easy assessing greenness: A comparison of NDVI datasets and neighborhood types and their associations with self-rated health in New York City. *Health & Place 54*, 92-101 (2018).https://doi.org/10.1016/j.healthplace.2018.09.005.] This information can then be used to inform and direct spending and efforts. While there exists a study by Jian Lin, Qiang Wang, and Xiaojiang Li regarding tree characteristics related to socioeconomic and spatial inequalities, my research aims to focus in tree characteristics to specific named neighborhoods.^[Jian Lin, Qiang Wang, Xiaojiang Li. Socioeconomic and spatial inequalities of street tree abundance, species diversity, and size structure in New York City. Landscape and Urban Planning, Volume 206. 2021. 103992. https://doi.org/10.1016/j.landurbplan.2020.103992.] Rather than using socioeconomic elements, my research question is focused on smaller geographic areas where individuals can identify themselves as members on a more personal level than a borough.

With these factors, I can test the hypothesis:  


H1: There is a significant difference in tree diameter based on which New York City borough the tree is located.  
H2: There is a significant difference in tree health based on which New York City borough the tree is located.  

# Descriptive Statistics  

```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(formattable)
library(dbplyr)
library(summarytools)
library(stargazer)
```

```{r}
library(readxl)
treecensus <- read_excel("C:/Users/aep05/Documents/UMASS_GRAD_SCHOOL/DACSS 603/603_Spring_2023/posts/finalproject/2015_Street_Tree_Census.xls")
print(treecensus)
```
This data set is obtained from the 2015 tree census through NYC Open Data provided by the New York City Department of Parks and Recreation.^[https://data.cityofnewyork.us/Environment/2015-Street-Tree-Census-Tree-Data/uvpi-gqnh]

The information was collected by predominantly volunteers in addition to the New York City Department of Parks and Recreation staff.  


There are 45 variables and 16,383 observations. 

```{r}
summary(treecensus)
```


Full list of variables:

```{r}
colnames(treecensus)
```
```{r}
dfSummary(treecensus)
```


At a glance, the columns I am interested in are:

tree_id	- Unique identification number for each tree point.  
tree_dbh	- Diameter of the tree, measured at approximately 54" / 137cm above the ground. Data was collected for both living and dead trees; for stumps, use stump_diam  
curb_loc - Location of tree bed in relationship to the curb; trees are either along the curb (OnCurb) or offset from the curb (OffsetFromCurb)   
status	- Indicates whether the tree is alive, standing dead, or a stump.  
health	- Indicates the user's perception of tree health.  
spc_latin	- Scientific name for species, e.g. "Acer rubrum"  
spc_common	- Common name for species, e.g. "red maple"  
steward - Indicates the number of unique signs of stewardship observed for this tree. Not recorded for stumps or dead trees.  
guards - Indicates whether a guard is present, and if the user felt it was a helpful or harmful guard. Not recorded for dead trees and stumps.  
sidewalk - 	Indicates whether one of the sidewalk flags immediately adjacent to the tree was damaged, cracked, or lifted. Not recorded for dead trees and stumps.  
root_stone - Indicates the presence of a root problem caused by paving stones in tree bed  
root_grate - Indicates the presence of a root problem caused by metal grates in tree bed  
root_other - Indicates the presence of other root problems   
trunk_wire - Indicates the presence of a trunk problem caused by wires or rope wrapped around the trunk  
trnk_light - Indicates the presence of a trunk problem caused by lighting installed on the tree 
trnk_other - Indicates the presence of other trunk problems  
brch_light - Indicates the presence of a branch problem caused by lights (usually string lights) or wires in the branches  
brch_shoe - Indicates the presence of a branch problem caused by sneakers in the branches  
brch_other - Indicates the presence of other branch problems  
postcode	- Five-digit zipcode in which tree is located  
borocode	- Code for borough in which tree point is located: 1 (Manhattan), 2 (Bronx), 3 (Brooklyn), 4 (Queens), 5 (Staten Island)  
borough	- Name of borough in which tree point is located  
nta	- This is the NTA Code corresponding to the neighborhood tabulation area from the 2010 US Census that the tree point falls into.  
nta_name	- This is the NTA name corresponding to the neighborhood tabulation area from the 2010 US Census that the tree point falls into.  
boro_ct	- This is the boro_ct identifyer for the census tract that the tree point falls into.  



```{r}
group_by(treecensus, status, health)%>%
 summarize()
```
Because I am only interested in trees with the status "Alive", I'll also need to clean the data set to exclude trees labeled as "Dead" or "Stump". 


```{r}
treecensus_clean <- treecensus[treecensus$status != "Dead" & treecensus$status != "Stump",]
group_by(treecensus_clean, status, health)%>%
 summarize()
```

The independent variable of interest is the categorization of geographic divisions within New York City into 5 boroughs. This variable is:  
borocode  

The dependent variables of interest are different tree characteristics observed during the census. These variables are:  
tree_dbh  
health    

Additional variables of interest for potential interactions are:  
curb_loc  
steward  
root_stone  
root_grate  
root_other  
trunk_wire  
trnk_light  
brch_light  
trnk_other  
brch_shoe  
brch_other  


```{r} 
#recoding character variables

treecensus_clean <- treecensus_clean %>% 
  mutate(curb_loc = recode(curb_loc,
                            "OnCurb" = 1,
                            "OffsetFromCurb" = 0)) %>%
  mutate(health = recode(health,
                            "Poor" = 1,
                            "Fair" = 2,
                            "Good" = 3)) %>%
  mutate(steward = recode(steward,
                            "None" = 0,
                            "1or2" = 1,
                            "3or4" = 3,
                            "4orMore" = 4)) %>%
  mutate(sidewalk = recode(sidewalk,
                            "NoDamage" = 0,
                            "Damage" = 1)) %>%
  mutate(root_stone = recode(root_stone,
                            "No" = 0,
                            "Yes" = 1)) %>%
  mutate(root_grate = recode(root_grate,
                            "No" = 0,
                            "Yes" = 1))%>%
  mutate(root_other = recode(root_other,
                            "No" = 0,
                            "Yes" = 1))%>%
  mutate(trunk_wire = recode(trunk_wire,
                            "No" = 0,
                            "Yes" = 1))%>%
  mutate(trnk_light = recode(trnk_light,
                            "No" = 0,
                            "Yes" = 1))%>%
  mutate(trnk_other = recode(trnk_other,
                            "No" = 0,
                            "Yes" = 1))%>%
  mutate(brch_light = recode(brch_light,
                            "No" = 0,
                            "Yes" = 1))%>%
  mutate(brch_shoe = recode(brch_shoe,
                            "No" = 0,
                            "Yes" = 1))%>%
  mutate(brch_other = recode(brch_other,
                            "No" = 0,
                            "Yes" = 1))
```





# Analysis

Expand the description of the data with visualizations, e.g. scatter plots among key
variables, box and whisker plots, histograms etc.

The following histograms show the frequency of key variables.  

```{r}
table(treecensus$health)%>%
  barplot(xlab = "Health", ylab = "Frequency")

table(treecensus$tree_dbh)%>%
  barplot(xlab = "Tree Diameter", ylab = "Frequency")

table(treecensus$steward)%>%
  barplot(xlab = "Steward", ylab = "Frequency")

table(treecensus$borocode)%>%
  barplot(xlab = "Borocode", ylab = "Frequency")

table(treecensus$curb_loc)%>%
  barplot(xlab = "Curb Location", ylab = "Frequency")
```


The first graph below displays the counts of observed problems trees may possess, then segmented by borocode. After the initial project proposal I have included these variables as another characteristic to test differences among the New York City neighborhoods. Each variable on the x-axis corresponds to a problem or characteristic a tree might have, codes for these are listed in the variable description section. It's immediately apparent that the answer "No" to these problems represents the majority of answers. It's also apparent that borough 3 & 4 have both the largest tree counts, and largest count of problem "root_stone".

The second graph below displays the distribution of tree diameter across those problems, then segmented by borough. Interestingly, the distribution shows similar values across all boroughs.


```{r}
treecensus %>% 
  pivot_longer(
    cols=starts_with(c('root', 'trunk', 'trnk', 'brch')),
    names_to = "Problem",
    values_to = "Response")%>%
  ggplot(aes(x=Problem))+
  geom_bar(aes(fill=Response))+
  facet_grid(borocode~.)


treecensus %>% 
  pivot_longer(
    cols=starts_with(c('root', 'trunk', 'trnk', 'brch')),
    names_to = "Problem",
    values_to = "Response")%>%
  ggplot(aes(x=Problem, y=tree_dbh))+
  geom_boxplot(aes(fill=Response))+
  ylim(0,50)+
  facet_grid(borocode~.)


```


```{r}
treecensus_clean$health <- as.factor(treecensus_clean$health)
```


The graphs below demonstrate how tree diameter is observed across "borocodes" (borough codes). It's then segmented based on the variable health. The graph shows that generally the median diameter falls within 5-15 inches among the boroughs but there is a different range for each borough. It also demonstrates the same general trend for health across borough that ranks health the lowest as poor, then fair, then good. There are 4 outliers on this map from boroughs 2, 3, & 5. These values are 425, 169, 122, 72. 

```{r}
ggplot(treecensus_clean, aes(x = factor(borocode), y = tree_dbh)) +
    geom_boxplot()+
    ylab("Diameter")+ 
    xlab("Borough") + 
    ylim(0,50)


ggplot(treecensus_clean, aes(x = factor(borocode), y = tree_dbh)) +
    geom_boxplot(aes(color = health))+
    scale_color_discrete(name = "Health", labels = c("Poor", "Fair", "Good"))+
    ylab("Diameter")+ 
    xlab("Borough") + 
    ylim(0,50)


treecensus_clean %>%
  arrange(desc(tree_dbh))%>%
  select(tree_dbh, borocode)%>%
  slice(1:4)
```

The graph below displays tree diameter based on borocode, then segmented by levels of steward. 0 = no signs, 1 = 1 to 2 signs, 3 = 3 to 4 signs, 4 = 4 or more. The distribution pattern of steward level 1 and 3 are similar across boroughs, where 0 and 4 appear to be different. The median across tree diameter does not appear to be consistent.  


```{r}
ggplot(treecensus_clean, aes(x = factor(borocode), y = tree_dbh)) +
    geom_boxplot()+
    ylab("Diameter")+ 
    xlab("Borocode")+
    ylim(0,75)+
  facet_grid(. ~steward)
    
```




This graph below demonstrates another visualization of the variables tree diameter, health, and borocode. It demonstrated the same trend among the three health groups that boroughs follow a similar trend of median with the highest borough 4 then 3. The lowest median across health falls both in borough 1 and 5. 

```{r}
ggplot(treecensus_clean, aes(x= health, y = tree_dbh, fill = factor(borocode)))+
  geom_boxplot()+
  ylim(0,75)
```

This graph demonstrates tree diameter based on borocode segmented by signs of stewardship. While the medians lay at relatively similar values, the distributions of diameter do differ noticeably. 

```{r}
ggplot(treecensus_clean, aes(x = factor(borocode), y = tree_dbh)) +
    geom_boxplot(aes(color = factor(steward)))+
    ylab("Diameter")+ 
    xlab("Borocode") + 
    ylim(0,50)
```


This graph below demonstrates tree diameter based on borocode segmented by curb location of the tree. While the graph shows generally the median measured at similar values, the range of diameter differs across borough. It does follow the same trend that a tree located off the curb has a larger or similar diameter in it's borough than a tree located on the curb.


```{r}
ggplot(treecensus_clean, aes(x = factor(borocode), y = tree_dbh)) +
    geom_boxplot(aes(color = factor(curb_loc)))+
    ylab("Diameter")+ 
    xlab("Borocode") + 
    ylim(0,50)
```



## Hypothesis Testing  

First I'll run an ANOVA test to review the population mean differences in key variables. 

H0 = There is no significant difference in means  
Ha = The means are different in at least one of the groups  


The first ANOVA is testing the independent variable borocode with dependent variable tree diameter.  

```{r}
anova_treedbh <- aov(tree_dbh ~ factor(borocode), data = treecensus_clean)
summary(anova_treedbh)

ggplot(treecensus_clean, aes(x = factor(borocode), y = tree_dbh)) +
    geom_boxplot(aes(color=factor(borocode)))+
    ylab("Diameter")+ 
    xlab("Borough") + 
    ylim(0,50)+
  scale_color_discrete(name = "Borocode")
```
This ANOVA between tree_dbh ~ borocode shows a very small p-value, indicating the null hypothesis can be rejected at 0.001.   



Next I'll test the ANOVA with the dependent variable health and independent variable borocde.    

```{r}
treecensus_clean$health <- as.numeric(treecensus_clean$health)
anova_treehealth <- aov(health ~ factor(borocode), data = treecensus_clean)
summary(anova_treehealth)

ggplot(treecensus_clean, aes(x = factor(borocode), y = health)) +
    geom_boxplot(aes(color=factor(borocode)))+
    ylab("Health")+ 
    xlab("Borough") + 
  scale_color_discrete(name = "Borocode")
```
This ANOVA between health ~ borocode shows a very small p-value, indicating the null hypothesis can be rejected at 0.001. From the graph, however, it seems there are two distinct patterns and not much variation.



## Model Comparisons  

Next we'll test different linear regression models. 

```{r, results='hide'}
slrtree <- lm(tree_dbh ~ borocode, data = treecensus_clean)
summary(slrtree)

slrtree2 <- lm(tree_dbh ~ borocode + steward, data = treecensus_clean)
summary(slrtree2)

slrtree3 <- lm(tree_dbh ~ borocode + steward + curb_loc, data = treecensus_clean)
summary(slrtree3)

slrtreehealth <- lm(health ~ borocode, data = treecensus_clean)
summary(slrtreehealth)

slrtreehealth2 <- lm(health ~ borocode + steward, data = treecensus_clean)
summary(slrtreehealth2)

slrtreehealth3 <- lm(health ~ borocode + steward + curb_loc, data = treecensus_clean)
summary(slrtreehealth3)

```



```{r}
stargazer(slrtree, slrtree2, slrtree3, type= 'text')
```
 
We'll first exam the models with DV as tree diameter.  

In the first model, Adjusted R2 only represents 1.2% of the variation in response variable tree diameter can be explained by the predictor variable borocode. Adjusted R2 does increase with each model to model 4's 3.8%, however, this is still very low. 

```{r}
stargazer(slrtree, slrtree2, slrtree3, type= 'text')
stargazer(slrtreehealth, slrtreehealth2, slrtreehealth3, type= 'text')
```


Next we'll exam the models with DV as health.  

In the first model, Adjusted R2 only represents 0.1% of the variation in response variable health can be explained by the predictor variable borocode. Adjusted R2 does increase with to 0.2% in Model 2 & 3, however, this is still very low. In Model 3, curb_loc is not statistically significant.   


## Diagnostics

#### Tree Diameter

We'll first run diagnostics on the linear regression for lm(tree_dbh ~ borocode + steward + curb_loc, data = treecensus_clean)


```{r}
plot(slrtree3, which = 1)
```
In this Residual vs. Fitted values plot, it does seem like it violates the constant variance assumption but does appear linear. 


```{r}
plot(slrtree3, which = 2)
```

This next plot is a QQ-Plot. These points do generally seem to normally fall along the line. 

```{r}
plot(slrtree3, which = 3)
```

The scale location plot also seems to violate constant variance as there are more plots between 10-14 range. 

```{r}
plot(slrtree3, which = 4)
```

The Cook's Distance plot also shows a violation of the influential observation assumption. Several observations are greater than 4/15,442=0.00026  


#### Tree Health  

Next we'll run dianogstics for the second dependent variable health, lm(health ~ borocode + steward, data = treecensus_clean)  

```{r}
plot(slrtreehealth2, which = 1)
```

This Residuals vs. fitted plot appears to violate the assumption of constant variance. 

```{r}
plot(slrtreehealth2, which = 2)
```

This QQ-plot shows the violation of normality as the points are not falling along the line. 


```{r}
plot(slrtreehealth2, which = 3)
```

The Scale-Location plot also shows a violation of the constant variance assumption.

```{r}
plot(slrtreehealth2, which = 4)
```


The Cook's Distance plot also shows a violation of the influential observation assumption. Many observations are greater than 4/15,442=0.00026  


