---
title: "Homework 1"
author: "Alexa Potter"
description: "DACSS 603"
date: "02/25/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Load Libraries

```{r}
library(tidyverse)
library(ggplot2)
library(formattable)
```


# Question 1

## a

First, let's read in the data from the Excel file:

```{r, echo=T}
library(readxl)
df <- read_excel("_data/LungCapData.xls")
```


The distribution of LungCap looks as follows:

```{r, echo=T}
hist(df$LungCap, xlab= 'LungCap', main = 'Histogram of LungCap Distribution')
```

The histogram suggests that the distribution is close to a normal distribution. Most of the observations are close to the mean. Very few observations are close to the margins (0 and 15).


##b

Compare the probability distribution of the LungCap with respect to Males and Females:

```{r}

boxplot(df$LungCap~df$Gender, xlab = 'Gender', ylab = 'LungCap')

```


##c

Compare the mean lung capacities for smokers and non-smokers. Does it make sense?

```{r}
LungCap_mean <- df %>% 
  group_by(Smoke) %>%
  summarise(Lungcap_Mean = mean(LungCap))

LungCap_mean

```

These results show the lung capacity for non-smokers is lower than the lung capacity for smokers. With no background knowledge of lung capacity, I would think these are not standard results. 


##d) 

Examine the relationship between Smoking and Lung Capacity within age groups: “less than or
equal to 13”, “14 to 15”, “16 to 17”, and “greater than or equal to 18”.

```{r}
df_Agegroup <- df %>% 
  mutate(
    Age_group = dplyr::case_when(
      Age <= 13            ~ "0-13",
      Age > 13 & Age <= 15 ~ "14-15",
      Age > 15 & Age <= 17 ~ "16-17",
      Age >= 18             ~ "18+"))


ggplot(data = df_Agegroup, aes(x=Age_group, y=LungCap)) + 
  geom_boxplot(aes(fill=Gender))
```


##e) 

Compare the lung capacities for smokers and non-smokers within each age group. Is your answer
different from the one in part c. What could possibly be going on here?

```{r}
ggplot(data = df_Agegroup, aes(x=Age_group, y=LungCap)) + 
  geom_boxplot(aes(fill=Smoke))
```

This data is different from what was seen in part C. The only age group displaying greater lung capacity for smokers is the group 0-13. This would suggest this sample of the population is influencing the overall lung capacity mean. This by far is the largest sample of age groups with 428 participants

```{r}
count(df_Agegroup, Age_group, Smoke)
```


# Question 2


First create the dataframe:

```{r}
stateprison <- data.frame (X  = c("0", "1", "2", "3", "4"),
                  Frequency = c("128", "434", "160", "64", "24")
                  )
stateprison$Frequency <- as.integer(stateprison$Frequency)
stateprison$X <- as.integer(stateprison$X)
stateprison

```



##a 

What is the probability that a randomly selected inmate has exactly 2 prior convictions?

```{r}
formattable::percent(160/summarise(stateprison, sum(Frequency)))
```


##b

What is the probability that a randomly selected inmate has fewer than 2 prior convictions?
```{r}
formattable::percent((128+434)/summarise(stateprison, sum(Frequency)))
```

##c

What is the probability that a randomly selected inmate has 2 or fewer prior convictions?
```{r}
formattable::percent((128+434+160)/summarise(stateprison, sum(Frequency)))
```

##d

What is the probability that a randomly selected inmate has more than 2 prior convictions?
```{r}
formattable::percent((64+24)/summarise(stateprison, sum(Frequency)))
```

##e

What is the expected value for the number of prior convictions?
```{r}
stateprison_PRB <- transform(stateprison, PROB = (Frequency/810))

df_prison_ev <- transform(stateprison_PRB, XP = X*PROB )


EV <- sum(df_prison_ev$XP)
print(EV)

```


##f

Calculate the variance and the standard deviation for the Prior Convictions.

Variance:
```{r}

varX <-sum(((df_prison_ev$X-EV)^2)*df_prison_ev$PROB)
print(varX)

```
Standard deviation:
```{r}
sdX <- sqrt(varX)
print(sdX)
```

