---
title: "Homework-2"
editor: visual
---

## Libraries

```{r}
library(tidyverse)

```

## Q.1)

```{r}
surgical_procedures <- c("bypass","angiography")
sample_size <- c(539, 847)
mean_wait_time <- c(19, 18)
standard_deviation <- c(10,9)

df <- data.frame(surgical_procedures, sample_size, mean_wait_time, standard_deviation)

```

CI for bypass

```{r}
# confidence level for bypass and angiography
conf_level <- 0.9

# standard error for bypass
bypass_se <- 10 / sqrt(539)

# confidence interval for bypass
bypassCI <- 19 + qt(c(0.05, 0.95), 539-1) * bypass_se
bypassCI
```

Ci for angio

```{r}
# standard error for angiography
angio_se <- 9 / sqrt(847)

# confidence interval for angiography
angioCI <- 18 + qt(c(0.05, 0.95), 847-1) * angio_se
angioCI
```

Size of confidence interval twice the margin of error

```{r}
2*qt(c(0.05, 0.95), 539-1) * bypass_se
```

```{r}
2*qt(c(0.05, 0.95), 847-1) * angio_se
```

-   The confidence interval for angiography is narrower.

## Q.2)

```{r}
# out of 1031 Americans surveyed
p <- 567 / 1031 
# 54% of Americans believe college education is essential for success

# 95% significant level
conf<- 0.95 

# standard error
college_se <- sqrt(p*(1-p)/1031) 

# confidence interval
collegeCI <- p + qnorm(c(0.025, 0.975)) * college_se
collegeCI 
```

## Q.3)

```{r}
range = 200-30
population_sd = range/4
z = qnorm(.975)
s = population_sd
n = ((z *s) / 5)^2
print(n)
```

## Q.4)

-   Setting up our t-test formula

    ```{r}
    t_test <- function(x_bar, mu, sd, n){
      return((x_bar - mu) / (sd / sqrt(n)))
    }

    # getting t-test score
    t_statistic <- t_test(x_bar = 410, mu = 500, sd = 90, n = 9)
    ```

## A)

```{r}
n = 9
pval_two_tail = 2*pt(t_statistic, df = n-1)
pval_two_tail
```

Rejecting null Hypothesis that population mean in 500

## B)

```{r}
pval_left_side_tail = pt(t_statistic, df = n-1)
pval_left_side_tail
```

-   Rejecting null hypothesis that population mean is greater than 500

## C)

```{r}
pval_right_side_tail = pt(t_statistic, df = n-1, lower.tail=FALSE)
pval_right_side_tail
```

-   Fail to reject that population mean is less than 500.

    ## Q.5)

    ```{r}
    #  T values
    t_jones <- (519.5 - 500) / 10 # sample mean = 519.5 - 500 for population mean / sample error of 10.0
    t_jones
    ```

```{r}
t_smith <- (519.7 - 500) / 10 # sample mean = 519.7 - 500 for population mean / sample error of 10.0
t_smith
```

```{r}
# p values
p_jones <- 2 * pt(-abs(t_jones), df = 999)
p_jones
```

```{r}
p_smith <- 2 * pt(-abs(t_smith), df = 999)
p_smith
```

-   Jones is not significant at 0.05 level but smith's result is significant

    ## Q.6)

```{r}
# dataframe creation
grade <- c(rep("6th grade", 100), rep("7th grade", 100), rep("8th grade", 100))
snack <- c(rep("healthy snack", 31), rep("unhealthy snack", 69), rep("healthy snack", 43),
           rep("unhealthy snack", 57), rep("healthy snack", 51), rep("unhealthy snack", 49))
snack_df <- data.frame(grade, snack)
```

Using chi square test as we are testing association between two categorical variables.

```{r}
table(snack_df$snack,snack_df$grade)
```

```{r}
chisq.test(snack_df$snack,snack_df$grade)
```

-   There is a relation between grades and choice of snack

## Q.7)

```{r}
Area <- c(rep("Area1", 6), rep("Area2", 6), rep("Area3", 6))
cost <- c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5, 7.5, 8.2, 8.5, 8.2, 7.0, 9.3,
          5.8, 6.4, 5.6, 7.1, 3.0, 3.5)
Area_cost <- data.frame(Area,cost)
```

using anova as comparing means of more than two groups

```{r}
one.way <- aov(cost ~ Area, data = Area_cost)
summary(one.way) 
```
