---
title: "Homework_5"
author: "Paritosh G"
desription: "HW_5"
date: "05/28/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - challenge1
  - my name
  - dataset
  - ggplot2
---

Loding the libraries

```{r}
library(tidyverse)
library(ggplot2)
library(smss)
library(alr4)
library(magrittr)
library(broom)
```

## Question 1

```{r}

data("house.selling.price.2")

df <- house.selling.price.2
```

```{r}
house_price <- lm(P ~ S + Be + Ba + New, data = df)
summary(house_price)
```

## a)

-   Beds would be deleted first as its p value of 0.4867 not statistically significant

## b)

-   Size(s) has the smallest p value and hence in forward method when we start with no explanatory variable Size(s) should be added first.

## c)

House with more bedrooms have most likely more number of bathrooms hence larger in size. If size, bedrooms, and bathrooms all are measuring price the high p value of bed is due to multi-collinearity.

## d)

```{r}
fit_No_Bed <- lm(P ~ S + Ba + New, data = df)
fit_No_Bed_and_Bath <- lm(P ~ S + New, data = df)
fit_new <- lm(P ~ New, data = df)

```

***R2***

```{r}
summary(fit_No_Bed)$r.squared
summary(fit_No_Bed_and_Bath)$r.squared
summary(fit_new)$r.squared
```

-   fit_No_Bed model is the best with highest R2 and our best model.

***Adjusted R2***

```{r}
summary(fit_No_Bed)$adj.r.squared
summary(fit_No_Bed_and_Bath)$adj.r.squared
summary(fit_new)$adj.r.squared





```

-   fit_No_Bed model has the highest adjusted R squared and our best model.

***Press***

**function for calculating press stat.**

```{r}
press_stat <- function(model) {
  # Calculate PRESS residuals
  pr <- resid(model) / (1 - lm.influence(model)$hat)
  
  # Compute the PRESS statistic
  press <- sum(pr^2)
  
  return(press)
}
```

calculating Press statistic for all models

```{r}
press_stat(fit_No_Bed)
press_stat(fit_No_Bed_and_Bath)
press_stat(fit_new)

```

-   fit_no_Bed_and_Bath has the lowest Press of 31066 and it is our best model.

***AIC***

```{r}
AIC(fit_No_Bed)
AIC(fit_No_Bed_and_Bath)
AIC(fit_new)

```

-   fit_No_bed has the lowest AIC score and is our best model.

***BIC***

```{r}
BIC(fit_No_Bed)
BIC(fit_No_Bed_and_Bath)
BIC(fit_new)
```

-   fit_No_Bed has the lowest BIC score and is our best model

## e)

hence model fit_no_Bed with variables size,bathroom and new is our best model.

## Q.2)

```{r}
data(trees)
```

## a)

```{r}
model <- lm(Volume ~ Girth + Height, data = trees)
summary(tree_model)
```

## b)

```{r}
par(mfrow = c(2,3))
plot(model, which = 1:6)
```

**Residuals vs Fitted**

The curved shape of the line suggests this model violates our assumption of linearity that the relationship between the independent and dependent variables is linear.

**Scale-Location**

The curved shape of the line suggests this model violates our assumption of constant variance, or homoscedasticity, which affects the statistical significance of the model.

**Cooks Distance, Residuals vs Leverage, Cooks dist vs Leverage**

All three charts show there is single, potentially influence point, which can impact whether our model meets the assumptions of linear regression and can disproportionately affect our regression coefficients.

## Q.3)

## a)

```{r}
data("florida")
model_2 <- lm(Buchanan ~ Bush, data = florida)
par(mfrow = c(2,3))
plot(model_2, which = 1:6)
```

Yes, based on the diagnostic plots, Palm Beach County is an outlier.

On the Residuals vs. Fitted Values plot, the residuals largely "bounce randomly" around the zero line and form a roughly horizontal band around the zero line - with Dade and Palm Beach as significant outliers.

The Normal Q-Q plot follows a straight line and the points do not deviate significantly from the line - again, with two major exceptions: Dade, and Palm Beach. The Scale-Location plot is roughly horizonal, but again has a major exception: Palm Beach.

Finally, the Resisuals vs. Leverage plot shows outliers outside of Cook's distance - Dade, and once again, Palm Beach.

## b)

```{r}
LogBuchanan <- lm(log(Buchanan) ~ log(Bush), data = florida)
summary(LogBuchanan)
```

```{r}
# Produce the regression diagnostic plots
par(mfrow = c(2, 2))
plot(LogBuchanan)
```

The p-value has gotten smaller and the adjusted r squared has gone from 0.3795 to 0.8485, suggesting that the log-log model is a better-fitting model. When it comes to the diagnostic plots, however, Palm Beach is still a notable outlier on the Residuals vs. Fitted, Normal Q-Q, and Scale-Location plots. It is now within Cook's distance on the Residuals vs. Leverage plot.
