---
title: "Homework_3"
editor: visual
Author : Paritosh_G
---

## Library

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(alr4)
library(smss)
library(stargazer)
```

## Q.1)

## A

-   Predictor variable :- ppgdp

-   Response variable :- fertility, birth rate

## B

```{r}
ggplot(UN11, aes(ppgdp, fertility, color = "red")) +
  geom_point()
```

## C

```{r}
ggplot(data = UN11, aes(x = log(ppgdp), y = log(fertility))) +
  geom_point()
```

we can imagine negatively sloped straight line and a plausible linear regression.

## Q.2)

## A)

the conversion from USD to British pound the numerical value will be divided by 1.23 to mitigate the effect slope will be divided by 1.23 as well.

## B)

correlation will not change as it does not affect by measuring unit.

## Q.3)

```{r}
data(water)
pairs(water)
```

## Q.4)

```{r}
#data(Rateprof)
pairs(Rateprof[,c('quality', 'clarity', 'helpfulness',
                  'easiness', 'raterInterest')])
```

There is a strong correlation among quality, clarity, and helpfulness . easiness is correlated with the other three. raterInterest is also fairly correlated, but raters almost always say they are at least somewhat interested in the subject. Overall, the results might show that people do not differentiate all these dimensions very well in their minds---or that professors that do one in one dimension tend to do well on the others too.

## Q.5)

## A)

```{r}
data("student.survey")
ggplot(student.survey, aes(x = re, fill = pi)) +
    geom_bar()
```

```{r}
ggplot(data = student.survey, aes(x = tv, y = hi)) +
  geom_point() +
  theme_bw()
```

## B)

```{r}
data("student.survey")

model_1 <- lm(as.numeric(pi) ~ as.numeric(re), 
         data = student.survey)
model_2 <- lm(hi ~ tv, data = student.survey)
summary(model_1)
summary(model_2)

```

There is a Positive and statistically significant relation between Religiosity and conservatism.

Watching TV for one more hour will decline the GPA by 0.018
