---
title: "Homework 1"
author: "Xiaoyan"
description: "Template of course blog qmd file"
date: "02/27/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---
```{r}
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
data<-read_excel("/Users/cassie199/Desktop/23spring/603_Spring_2023-1/posts/_data/LungCapData.xls")
head(data)
```
# Question 1
1. Use the LungCapData to answer the following questions. (Hint: Using dplyr, especially
group_by() and summarize() can help you answer the following questions relatively efficiently.)
a) What does the distribution of LungCap look like? (Hint: Plot a histogram with probability density on the y axis)

-the distribution of LungCap looks like a normal distribution
```{r}
ggplot(data, aes(x=LungCap)) + geom_histogram()
```
b) Compare the probability distribution of the LungCap with respect to Males and Females? (Hint:make boxplots separated by gender using the boxplot() function)
```{r}
boxplot(data$LungCap~data$Gender)
```
c) Compare the mean lung capacities for smokers and non-smokers. Does it make sense?
```{r}
boxplot(data$LungCap~data$Smoke)
```
d) Examine the relationship between Smoking and Lung Capacity within age groups: “less than or
equal to 13”, “14 to 15”, “16 to 17”, and “greater than or equal to 18”.
```{r}
age_ranges <- c("<=13", "14-15", "16-17",  ">18")
data$age_ranges <- cut(data$Age, breaks = c(13, 14,15, 16,17, 18),
                     include.lowest = TRUE)
ggplot(data, aes(x = age_ranges, y = LungCap)) +
  geom_boxplot() +
  labs(x = "Age Range", y = "LungCap")
```
e) Compare the lung capacities for smokers and non-smokers within each age group. Is your answer different from the one in part c. What could possibly be going on here?

#Question 2

2. Let X = number of prior convictions for prisoners at a state prison at which there are 810
prisoners.

X 0 1 2 3 4
Frequency 128 434 160 64 24

a) What is the probability that a randomly selected inmate has exactly 2 prior convictions?
```{r, echo=T}
n<-810
X<-tibble(x=0:4,
          F=c(128,434,160,64,24))
X
pa<-160/n
pa

```
b) What is the probability that a randomly selected inmate has fewer than 2 prior convictions?
```{r, echo=T}
pb<-(128+434)/n
pb
```
c) What is the probability that a randomly selected inmate has 2 or fewer prior convictions?
```{r, echo=T}
pc<-(128+434+160)/n
pc
```
d) What is the probability that a randomly selected inmate has more than 2 prior convictions?
```{r, echo=T}
pd<-(64+24)/n
pd
```
e) What is the expected value1 for the number of prior convictions?
```{r, echo=T}
prior<-c(434,160,64,24)
mean(prior)
```
f) Calculate the variance and the standard deviation for the Prior Convictions.
```{r, echo=T}
var(prior)
sd(prior)
```


