---
title: "Homework1"
author: "Rahul Somu"
description: "HW1 course blog qmd file"
date: "02/27/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Question 1

## a

First, let's read in the data from the Excel file:

```{r, echo=T}
library(readxl)
library(dplyr)
library(ggplot2)
getwd()
df <- read_excel("_data/LungCapData.xls")
```


The distribution of LungCap looks as follows:

1a) The histogram suggests that the distribution is close to a normal distribution. Most of the observations are close to the mean. Very few observations are close to the margins (0 and 15).


```{r, echo=T}
hist(df$LungCap)
```

1b) Median lung capacity of male is greater than that of female.

```{r, echo=T}
boxplot(LungCap ~ Gender, data = df, xlab = "Gender", ylab = "Lung Capacity",
        main = "Distribution of Lung Capacity by Gender")
```

1c) Logically mean lung capacity of non-smokers should be more than of smokers but with the data, it's other way round

```{r, echo=T}
# Calculate the mean lung capacity for smokers and non-smokers
mean_lungcap_smokers <- mean(df$LungCap[df$Smoke == "yes"])
mean_lungcap_non_smokers <- mean(df$LungCap[df$Smoke == "no"])

# Print the mean lung capacities
cat("Mean lung capacity for smokers:", round(mean_lungcap_smokers, 2), "\n")
cat("Mean lung capacity for non-smokers:", round(mean_lungcap_non_smokers, 2), "\n")
```

1d) The average lung capacity growth for the non-smokers is more that of the smokers.
The lung capacity has been gradually increasing with age
1e) The discrepancy in 1c is due the data for the 13 or younger age where the average lung capacity for the non-smokers is less that of the smokers. Also there have been more data points for 13 or younger age group non-smokers which is affecting the mean of entire distribution.

```{r, echo=T}
# Define the age groups
df <- df %>%
  mutate(age_groups = cut(Age, c(0, 13, 15, 17, Inf), labels = c("<= 13", "14-15", "16-17", ">= 18")))

# Compare the probability distribution of lung capacity by gender
df %>%
  ggplot(aes(x = Gender, y = LungCap)) +
  geom_boxplot() +
  labs(x = "Gender", y = "Lung Capacity", 
       title = "Lung Capacity by Gender")

# Compare the mean lung capacities for smokers and non-smokers
df %>%
  group_by(Smoke) %>%
  summarize(mean_lungcap = mean(LungCap)) %>%
  print()

# Examine the relationship between smoking and lung capacity within age groups
df %>%
  filter(Smoke %in% c("yes", "no")) %>%
  group_by(age_groups, Smoke) %>%
  summarize(mean_lungcap = mean(LungCap)) %>%
  print()

# Compare the lung capacities for smokers and non-smokers within each age group
df %>%
  filter(Smoke %in% c("yes", "no")) %>%
  ggplot(aes(x = age_groups, y = LungCap, fill = Smoke)) +
  geom_boxplot() +
  labs(x = "Age Group", y = "Lung Capacity", 
       title = "Lung Capacity by Smoking Status and Age Group")


```

#Challange2

# 2a) probability of having exactly 2 prior convictions: 0.1975

```{r, echo=T}
# Define the values
x <- c(0, 1, 2, 3, 4)

p2 <- 160/810
p2
```
# 2b) probability of having fewer than 2 prior convictions: 0.6938272

```{r, echo=T}
p_less2 <- (128 + 434) / 810
p_less2
```

# 2c) probability of having 2 or fewer prior convictions: 0.891358

```{r, echo=T}
p_2less <- (128 + 434 + 160) / 810
p_2less
```

# 2d)probability of having more than 2 prior convictions: 0.108642

```{r, echo=T}
p_more2 <- (64 + 24) / 810
p_more2
```

# 2e)expected value for the number of prior convictions: 1.28642

```{r, echo=T}
ex <- sum(c(0, 1, 2, 3, 4) * c(128, 434, 160, 64, 24) / 810)
ex
```

#2f) Variance: 0.898864 & Standard deviation: 0.9480844

```{r, echo=T}
p_x <- c(0.158, 0.536, 0.198, 0.079, 0.029)
mu <- 1.502

# Calculate variance and standard deviation
variance <- sum((x - mu)^2 * p_x)
sd <- sqrt(variance)

# Print results
cat("Variance:", variance, "\n")
cat("Standard deviation:", sd)
```
