---
title: "Homework 1"
author: "Guanhua Tan"
description: "Homework 1"
date: "02/05/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Question 1

## a)

First, let's read in the data from the Excel file:

```{r, echo=T}
library(readxl)
library(tidyverse)
library(ggplot2)
df <- read_excel("_data/LungCapData.xls")
head(df)
```

The distribution of LungCap looks as follows:

```{r, echo=T}
hist(df$LungCap)
```

The histogram suggests that the distribution is close to a normal distribution. Most of the observations are close to the mean. Very few observations are close to the margins (0 and 15).

## b)  Compare the probability distribution of the LungCap with respect to Males and Females?

```{r, echo=T}
df %>%
  ggplot(aes(x=Gender,y=LungCap))%+%
  stat_boxplot(geom = "errorbar", # Error bars
               width = 0.2)+
  geom_boxplot()


```

The box graphic suggests that the median of male lung capacities are slightly larger than the one of female ones.

## c)  Compare the mean lung capacities for smokers and non-smokers. Does it make sense?

```{r, echo=T}
df_c <- df %>%
  group_by(Smoke) %>%
  mutate(mean_lungcap=mean(LungCap))%>%
  distinct(mean_lungcap)
df_c 
```

The data indicates smokers' lung capacities are larger than no-smokers' ones. It runs counter to the intuition.

## d)  Examine the relationship between Smoking and Lung Capacity within age groups: "less than or equal to 13", "14 to 15", "16 to 17", and "greater than or equal to 18".

```{r, echo=T}
# less than or equal to 13
df_d_13<-df %>%
  filter(Smoke == "yes" & Age <= 13) %>%
  mutate(mean_lungcap=mean(LungCap)) %>%
  distinct(mean_lungcap)


df_d_14_15 <-df %>%
   filter(Smoke == "yes" & Age <= 15 | Age >= 14) %>%
  mutate(mean_lungcap=mean(LungCap)) %>%
  distinct(mean_lungcap)


df_d_16_17 <-df %>%
   filter(Smoke == "yes" & Age <= 17 | Age >= 16) %>%
  mutate(mean_lungcap=mean(LungCap)) %>%
  distinct(mean_lungcap)


df_d_18 <-df %>%
   filter(Smoke == "yes" & Age >= 18) %>%
  mutate(mean_lungcap=mean(LungCap)) %>%
  distinct(mean_lungcap)
result <-c(df_d_13, df_d_14_15, df_d_16_17, df_d_18)
print(result)
```

The data indicates that with the increase of the age, the lung capacities grows larger.

## e)  Compare the lung capacities for smokers and non-smokers within each age group. Is your answer different from the one in part c. What could possibly be going on here?

```{r}
df_e_13<-df %>%
  filter(Age <= 13) %>%
  group_by(Smoke) %>%
  mutate(mean_lungcap=mean(LungCap)) %>%
  distinct(mean_lungcap)

df_e_14<-df %>%
  filter(Age == 15 | Age == 14) %>%
  group_by(Smoke) %>%
  mutate(mean_lungcap=mean(LungCap)) %>%
  distinct(mean_lungcap)


df_e_16<-df %>%
  filter(Age == 17 | Age == 16) %>%
  group_by(Smoke) %>%
  mutate(mean_lungcap=mean(LungCap)) %>%
  distinct(mean_lungcap)


df_e_18<-df %>%
  filter( Age >= 18) %>%
  group_by(Smoke) %>%
  mutate(mean_lungcap=mean(LungCap)) %>%
  distinct(mean_lungcap)

df_e_13
df_e_14
df_e_16
df_e_18


```

The shows a big difference from the part C. Only in age group under 13, smokers have larger lung capacities than non-smokers. In other age groups, unlike what the part C suggests, non-smokers have large lung capacities that smokers.

# Question2
2.  Let X = number of prior convictions for prisoners at a state prison at which there are 810 prisoners.

<!-- -->

## a)  What is the probability that a randomly selected inmate has exactly 2 prior convictions?

```{r, echo=T}
c <- 160/810
c
```

The probability is 19.8%.

## b)  What is the probability that a randomly selected inmate has fewer than 2 prior convictions?

```{r, echo=T}
c<-(128+434)/810
c
```

The probability is 69.4%.

## c)  What is the probability that a randomly selected inmate has 2 or fewer prior convictions?

```{r, echo+T}
c <- (434+160+128)/810
c
```

The probability is 89.1%.

## d)  What is the probability that a randomly selected inmate has more than 2 prior convictions?

```{r,echo=T}
c<-(64+24)/810
c
```

The probability is 10.9%.

## e)  What is the expected value1 for the number of prior convictions?

```{r, echo=T}
vals<-c(0,1,2,3,4)
probs<-c(128/810, 434/810, 160/810, 64/801, 24/810)
exv<-weighted.mean(vals, probs)
exv
```

The expected value is 1.29.

f)  Calculate the variance and the standard deviation for the Prior Convictions.

```{r, echo=T}
var <- sum((vals-exv)^2*probs)
var
sd <- sqrt(var)
sd
```

The variance is 0.8588. The standard deviation is 0.9267.
