---
title: "Homework - 1"
author: "Thrishul"
description: "Homework 1"
date: "02/05/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Question 1

## a

First, let's read in the data from the Excel file:

```{r, echo=T}
library(readxl)
library(tidyverse)
df <- read_excel("_data/LungCapData.xls")
head(df)
```


The distribution of LungCap looks as follows:

```{r, echo=T}
hist(df$LungCap)
```

The histogram suggests that the distribution is close to a normal distribution. Most of the observations are close to the mean. Very few observations are close to the margins (0 and 15).

update and comit check
```{r, echo=T}
# Subset the data frame by gender
male_df <- df[df$Gender == "male", ]
female_df <- df[df$Gender == "female", ]

# Create separate boxplots for males and females
boxplot(male_df$LungCap, female_df$LungCap, 
        names = c("Male", "Female"),
        xlab = "Gender", ylab = "Lung Capacity",
        main = "Lung Capacity by Gender")

```
```{r}
no_smoke <- df[df$Smoke == "no",]
yes_smoke <- df[df$Smoke == "yes",]
mean(no_smoke$LungCap)
mean(yes_smoke$LungCap)
```


```{r}
```


```{r}
age_group_1 <- df[df$Age <= 13, ]
age_group_2 <- df[df$Age >= 14 & df$Age <= 15, ]
age_group_3 <- df[df$Age >= 16 & df$Age <= 17, ]
age_group_4 <- df[df$Age >= 18, ]
par(mfrow=c(2,2)) # Set up a 2x2 grid of plots

boxplot(LungCap ~ Smoke, data = age_group_1, 
        names = c("Non-Smoker", "Smoker"),
        main = "Age Group <= 13")

boxplot(LungCap ~ Smoke, data = age_group_2, 
        names = c("Non-Smoker", "Smoker"),
        main = "Age Group 14-15")

boxplot(LungCap ~ Smoke, data = age_group_3, 
        names = c("Non-Smoker", "Smoker"),
        main = "Age Group 16-17")

boxplot(LungCap ~ Smoke, data = age_group_4, 
        names = c("Non-Smoker", "Smoker"),
        main = "Age Group >= 18")
```


```{r}
# Calculate the mean and standard deviation of Lung Capacity for each age group and smoking status
agg_data <- aggregate(LungCap ~ Age + Smoke, data = df, 
                      FUN = function(x) c(mean = mean(x), sd = sd(x)))

agg_data

```
#Question 2
# Let X = number of prior convictions for prisoners at a state prison at which there are 810 prisoners.

# X 0 1 2 3 4
# Frequency 128 434 160 64 24

# a) What is the probability that a randomly selected inmate has exactly 2 prior convictions?
```{r}
# P(x=2)= 160/810
# P(x) = 0.1975
# p(x) = 19.75%
```
# b) What is the probability that a randomly selected inmate has fewer than 2 prior convictions?
```{r}
# Frequency of X=0 + Frequency of X=1 = 128 + 434 = 562
# p(x) = 562/810
# P(x) = 0.6938 or 69.38%
# Probability = 69.38%
```
# c) What is the probability that a randomly selected inmate has 2 or fewer prior convictions?
```{r}
# Frequency of X=0 + Frequency of X=1 + Frequency of X=2 = 128 + 434 + 160 = 722
# P(X ≤ 2) = frequency of X≤2 / total number of inmates
#  722/810 = 0.8914
# probability = 0.8914 or 89.14%
```
# d) What is the probability that a randomly selected inmate has more than 2 prior convictions?
```{r}
# Frequency of X=3 + Frequency of X=4 = 64 + 24 = 88
# P(X > 2) = frequency of X>2 / total number of inmates

# 88/810 = 0.1086
# Probability = 0.1086 or 10.86%
```
# e) What is the expected value1 for the number of prior convictions?
```{r}
# E(X) = Σ(xi * pi)
# (0 * 128/810) + (1 * 434/810) + (2 * 160/810) + (3 * 64/810) + (4 * 24/810)
# 1.283

# Expected value for the number of prior convictions is 1.283.
```
# f) Calculate the variance and the standard deviation for the Prior Convictions
```{r}
# Var(X) = E(X^2) - [E(X)]^2
# E(X^2) = Σ(xi^2 * pi)
# P(X=x) 128/810 434/810 160/810 64/810 24/810
# E(X^2) = Σ(xi^2 * pi)
# = (0^2 * 128/810) + (1^2 * 434/810) + (2^2 * 160/810) + (3^2 * 64/810) + (4^2 * 24/810)
# = 2.105

# VARIANCE
# Var(X) = E(X^2) - [E(X)]^2
# = 2.105 - (1.283)^2
# variance for the number of prior convictions is = 0.921

# STANDARD DEVIATION
# SD(X) = sqrt(Var(X))
# = sqrt(0.921)
# = 0.960
# 


```






























