---
title: "Homework 2"
author: "Miguel Curiel"
description: "Confidence intervals and hypothesis testing homework for DACSS 603."
date: "03/28/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - confidence intervals 
  - hypothesis testing
  - chi-square
  - anova
editor: 
  markdown: 
    wrap: 72
---

# Question 1

The time between the date a patient was recommended for heart surgery
and the surgery date for cardiac patients in Ontario was collected by
the Cardiac Care Network ("Wait Times Data Guide," Ministry of Health
and Long-Term Care, Ontario, Canada, 2006). The sample mean and sample
standard deviation for wait times (in days) of patients for two cardiac
procedures are given in the accompanying table. Assume that the sample
is representative of the Ontario population.

| Surgical Procedure | Sample Size | Mean Wait Time | Standard Deviation |
|--------------------|-------------|----------------|--------------------|
| Bypass             | 539         | 19             | 10                 |
| Angiography        | 847         | 18             | 9                  |

Construct the 90% confidence interval to estimate the actual mean wait
time for each of the two procedures. Is the confidence interval narrower
for angiography or bypass surgery?

```{r, eval=TRUE, echo=TRUE}
# setting seed for reproducible results
set.seed(0)

# 90% confidence interval for bypass
bypass_ci <- t.test(x=c(rnorm(539, mean=19, sd=10)), conf.level=0.90)$conf.int

# 90% confidence interval for bypass
angiography_ci <- t.test(x=c(rnorm(847, mean=18, sd=9)), conf.level=0.90)$conf.int

# print results
cat("Bypass 90% confidence interval:", bypass_ci, "\n"
    , "Angiography 90% confidence interval:", angiography_ci, "\n")
```

As we can see from the results, bypass' 90% confidence interval is
18.6-20.1, while angiography's is 17.5-18.5. Therefore, angiography's
confidence interval is narrower.

# Question 2

A survey of 1031 adult Americans was carried out by the National Center
for Public Policy. Assume that the sample is representative of adult
Americans. Among those surveyed, 567 believed that college education is
essential for success. Find the point estimate, p, of the proportion of
all adult Americans who believe that a college education is essential
for success. Construct and interpret a 95% confidence interval for p.

```{r, eval=TRUE, echo=TRUE}
# setting seedd for reproducible results
set.seed(0)

# point estimate
point_estimate <- 567/1031

# 95% confidence interval
proportion_test <- prop.test(567, 1031, conf.level = 0.95)$conf.int

# print results
cat("Point estimate:", point_estimate, "\n"
    , "95% confidence interval:", proportion_test, "\n")
```

The results indicate that we are 95% confident that the true proportion
of all adult Americans who believe that college education is essential
for success lies between .52-.58. In other words, we can say with 95%
certainty that 52%-58% of adult Americans believe that college is
essential for success.

# Question 3

Suppose that the financial aid office of UMass Amherst seeks to estimate
the mean cost of textbooks per semester for students. The estimate will
be useful if it is within \$5 of the true population mean (i.e. they
want the confidence interval to have a length of \$10 or less). The
financial aid office is pretty sure that the amount spent on books
varies widely, with most values between \$30 and \$200. They think that
the population standard deviation is about a quarter of this range (in
other words, you can assume they know the population standard
deviation). Assuming the significance level to be 5%, what should be the
size of the sample?

```{r, eval=TRUE, echo=TRUE}
# margin of error formula
## ME = z * (sigma / sqrt(n))
### solving for n
#### n = (z * sigma/ME)^2
##### replacing values
###### z = significance level 5% = critical z value for 95% conf int = 1.96
###### sigma = 1/4 of range of textbook costs = (200-30)/4 = 42.5
###### ME = estimate within 5 of the true population mean = 5

# calculate result
n <- round((1.96*42.5/5)^2)

# print result
cat("Ideal sample size:", n, "\n")
```

Using the margin of error formula for confidence intervals and solving
for n (sample size), we see that the ideal sample size (rounded to the
nearest integer) is 278.

# Question 4

According to a union agreement, the mean income for all senior-level
workers in a large service company equals \$500 per week. A
representative of a women's group decides to analyze whether the mean
income μ for female employees matches this norm. For a random sample of
nine female employees, ȳ = \$410 and s = 90.

A.  Test whether the mean income of female employees differs from \$500
    per week. Include assumptions, hypotheses, test statistic, and
    P-value. Interpret the result.

    1.  Assumptions:

        a.  The data is normally distributed.

        b.  The sample is a simple random sample.

        c.  The standard deviation of the population is unknown.

    2.  Hypotheses:

        a.  H0: μ = 500

        b.  H1: Ha: μ ≠ 500

    3.  Test statistic:

        a.  t = (ȳ - μ) / (s / sqrt(n)) = (410 - 500) / (90 / sqrt(9)) =
            -3

    4.  P-value:

        a.  The P-value is 0.01707168, which is less than the level of
            significance of 0.05. Therefore, we reject the null
            hypothesis and conclude that there is sufficient evidence to
            suggest that the mean income of female employees **differs**
            from \$500 per week.

B.  Report the P-value for Ha: μ \< 500. Interpret.

    1.  The P-value is 0.008535841, which is less than the level of
        significance of 0.05. Therefore, we reject the null hypothesis
        and conclude that there is sufficient evidence to suggest that
        the mean income of female employees **is less than** \$500 per
        week.

C.  Report and interpret the P-value for Ha: μ \> 500.

    1.  The P-value is 0.9914642, which is more than the level of
        significance of 0.05. Therefore, we fail to reject the null
        hypothesis and conclude that there is sufficient evidence to
        suggest that the mean income of female employees **is greater
        than** \$500 per week.

```{r, eval=TRUE, echo=TRUE, results='hide'}
# A.3.a. Test statistic
(410 - 500) / (90 / sqrt(9))

# A.4.a. P-value two-sided
pt(-3, 8) * 2

# B.1. P-value μ < 500
pt(-3, 8)

# C.1. P-value μ > 500
pt(-3, 8, lower.tail=FALSE)
```

# Question 5

Jones and Smith separately conduct studies to test H0: μ = 500 against
Ha: μ ≠ 500, each with n = 1000. Jones gets ȳ = 519.5, with se = 10.0.
Smith gets ȳ = 519.7, with se = 10.0.

A.  Show that t = 1.95 and P-value = 0.051 for Jones. Show that t = 1.97
    and P-value = 0.049 for Smith.

    1.  Jones:

        a.  t = (ȳ - μ) / se = (519.5 - 500) / 10 = 19.5 / 10 = 1.95

        b.  P-value = round(2 \* pt(q=1.95, df=999,
            lower.tail=FALSE), 3) = 0.051

    2.  Smith:

        a.  t = (ȳ - μ) / se = (519.7 - 500) / 10 = 19.7 / 10 = 1.97

        b.  P-value = round(2 \* pt(q=1.97, df=999,
            lower.tail=FALSE), 3) = 0.049

B.  Using α = 0.05, for each study indicate whether the result is
    "statistically significant."

    1.  Since Jones' P-value is greater than 0.05, we fail to reject the
        null hypothesis, indicating that his results are **not
        statistically significant**. In contrast, Smith's P-value is
        less than 0.05, therefore we reject the null hypothesis and his
        results **are statistically significant**.

C.  Using this example, explain the misleading aspects of reporting the
    result of a test as "P ≤ 0.05" versus "P \> 0.05," or as "reject H0"
    versus "Do not reject H0," without reporting the actual P-value.

    1.   "P ≤ 0.05" or "reject H0" without reporting the actual P-value
        can be misleading because it doesn't provide information on how
        strong the evidence is against the null hypothesis. For example,
        both Jones and Smith barely pass the 0.05 threshold, having
        0.049 and 0.051, respectively. Reporting this would help readers
        and analysts to take the strength of the evidence in
        consideration (in this case, "rejecting H0" or "failing to
        reject H0" should be taken with a grain of salt).

```{r, eval=TRUE, echo=TRUE, results='hide'}
# A.1.b. Proving Jones' p-value
round(2 * pt(q=1.95, df=999, lower.tail=FALSE), 3)

# A.2.b. Proving Smith's p-value
round(2 * pt(q=1.97, df=999, lower.tail=FALSE), 3)
```

# Question 6

A school nurse wants to determine whether age is a factor in whether
children choose a healthy snack after school. She conducts a survey of
300 middle school students, with the results below. Test at α = 0.05 the
claim that the proportion who choose a healthy snack differs by grade
level. What is the null hypothesis? Which test should we use? What is
the conclusion?

| Grade level     | 6th grade | 7th grade | 8th grade |
|-----------------|-----------|-----------|-----------|
| Healthy snack   | 31        | 43        | 51        |
| Unhealthy snack | 69        | 57        | 49        |

-   Null hypothesis: There is no difference in the proportion of
    students who choose a healthy snack based on grade level.

-   Test: Chi-squared test because we are assessing whether proportions
    of outcomes (choosing healthy versus unhealthy snacks) in each grade
    are equal or different.

-   Conclusion: Since the p-value is 0.01547, we reject the null
    hypothesis at the 0.05 level of significance and conclude that there
    is a significant difference in the proportion of healthy snack
    choices among the different grade levels.

```{r, eval=TRUE, echo=TRUE}
# create a matrix of the observed values
observed <- matrix(c(31, 43, 51, 69, 57, 49), nrow = 2, byrow = TRUE)

# perform the chi-squared test
result <- chisq.test(observed)

# print the results
print(observed)
print(result)
```

# Question 7

Per-pupil costs (in thousands of dollars) for cyber charter school
tuition for school districts in three areas are shown. Test the claim
that there is a difference in means for the three areas, using an
appropriate test. What is the null hypothesis? Which test should we use?
What is the conclusion?

|        |     |     |     |     |     |     |
|--------|-----|-----|-----|-----|-----|-----|
| Area 1 | 6.2 | 9.3 | 6.8 | 6.1 | 6.7 | 7.5 |
| Area 2 | 7.5 | 8.2 | 8.5 | 8.2 | 7.0 | 9.3 |
| Area 3 | 5.8 | 6.4 | 5.6 | 7.1 | 3.0 | 3.5 |

-   Null hypothesis: There is no difference in means for the three
    areas.

-   Test: Analysis of Variance (ANOVA) because we are computing the
    difference between the means of three or more groups.

-   Conclusion: Given that the P-value associated to the F-statistic is
    0.00397, we reject the null hypothesis and conclude that there is a
    significant difference in means for the three areas.

```{r, eval=TRUE, echo=TRUE}
area1 <- c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5)
area2 <- c(7.5, 8.2, 8.5, 8.2, 7.0, 9.3)
area3 <- c(5.8, 6.4, 5.6, 7.1, 3.0, 3.5)

anova_result <- aov(c(area1, area2, area3) ~ rep(c("Area 1", "Area 2", "Area 3")
                                                 , c(6, 6, 6)))
print(summary(anova_result))
```
