---
title: "Homework - 2"
author: "Thrishul"
description: "Homework 2"
date: "02/05/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - desriptive statistics 
  - probability
---

## Question 1
1. The time between the date a patient was recommended for heart surgery and the surgery date
for cardiac patients in Ontario was collected by the Cardiac Care Network (“Wait Times Data
Guide,” Ministry of Health and Long-Term Care, Ontario, Canada, 2006). The sample mean
and sample standard deviation for wait times (in days) of patients for two cardiac procedures
are given in the accompanying table. Assume that the sample is representative of the Ontario
population

        Surgical Procedure Sample Size Mean Wait Time Standard Deviation
Bypass     539            19                10
Angiography 847           18                9

```{r}
# To construct a confidence interval for the mean wait time for each procedure, we will use the following formula:

# CI = x̄ ± zα/2 * (σ/√n)
# For bypass surgery:
# x̄ = 19
# σ = 10
# n = 539
# zα/2 = 1.645

# CI = 19 ± 1.645 * (10/√539) = (17.8, 20.2)

# For angiography:
# x̄ = 18
# σ = 9
# n = 847
# zα/2 = 1.645

# CI = 18 ± 1.645 * (9/√847) = (17.2, 18.8)

# The confidence interval for bypass surgery is (17.8, 20.2) and the confidence interval for angiography is (17.2, 18.8). We can see that the confidence interval for angiography is narrower, which means that we are more certain about the mean wait time for angiography than for bypass surgery.


```

## Question 2
```{r}
# The point estimate, p, of the proportion of all adult Americans who believe that a college education is essential for success is:

# p = 567/1031 = 0.5498

# To construct a 95% confidence interval for p, we will use the following formula:

# CI = p ± zα/2 * √(p(1-p)/n)

# 95% corresponds to a z-score of 1.96

# CI = 0.5498 ± 1.96 * √(0.5498(1-0.5498)/1031) = (0.517, 0.582)
# The 95% confidence interval for the proportion of all adult Americans who believe that a college education is essential for success is (0.517, 0.582). This means that we can be 95% confident that the true proportion of all adult Americans who believe that a college education is essential for success falls within this interval. 

# We can interpret this as saying that, based on the sample data, we estimate that between 51.7% and 58.2% of all adult Americans believe that a college education is essential for success.
```
## Question 3

```{r}



```































