---
title: "Homework 2"
author: "Rosemary Pang"
description: "The second homework on hypothesis testing and confidence intervals"
date: "03/29/2023"
format:
  html:
    toc: true
    code-fold: false 
    code-copy: true
    code-tools: true
categories:
  - hw2
  - hypothesis testing
  - confidence intervals
---

Please check your answers against the solutions.

## Question 1

```{r, echo=TRUE}
bypass_n = 539
angio_n = 847

bypass_sample_mean = 19
angio_sample_mean = 18

bypass_sample_sd = 10
angio_sample_sd = 9

bypass_se = bypass_sample_sd/sqrt(bypass_n)
angio_se = angio_sample_sd/sqrt(angio_n)

bypass_me = qt(0.95, df = bypass_n - 1)*bypass_se
angio_me = qt(0.95, df = angio_n - 1)*angio_se
```

The confidence intervals:

```{r, echo=TRUE}
print(bypass_sample_mean + c(-bypass_me, bypass_me))
print(angio_sample_mean + c(-angio_me, angio_me))
```

The size of the confidence intervals, which is twice the margin of error:

```{r, echo=TRUE}
2 * bypass_me
2 * angio_me
```

The confidence interval for angiography is narrower.

## Question 2

one-step solution:

```{r, echo=TRUE}
n = 1031
k = 567
prop.test(k, n)
```

Alternatively:

```{r, echo=TRUE}
p_hat <- k/n # point estimate
se = sqrt((p_hat*(1-p_hat))/n) # standard error
e = qnorm(0.975)*se # margin of error
p_hat + c(-e, e) # confidence interval 
```

Alternatively, we can use the exact binomial test. In large samples like the one we have, the results should essentially be the same as prop.test().

```{r, echo=TRUE}
binom.test(k, n)
```

## Question 3

```{r, echo=TRUE}
range = 200-30
population_sd = range/4
```

Remember:

$$CI_{95} = \bar x \pm z \frac{s}{\sqrt n}$$
(We can use $z$ because we assume population standard deviation is known.)

We want the number $n$ that ensures:

$$ z \frac{s}{\sqrt n} = 5 $$
$$ zs = 5 \sqrt n$$
$$ \frac{zs}{5} = \sqrt n$$
$$  (\frac{zs}{5})^2 = n$$

In our case:

```{r, echo = T}
z = qnorm(.975)
s = population_sd
n = ((z *s) / 5)^2
print(n)
```

Rounding up, we need a sample of 278.

## Question 4

We can write a function to find the t-statistic, and then do all the tests in a, b, and c using that.

$$t = \frac{\bar x - \mu}{s / \sqrt n}$$

where $\bar x$ is them sample mean, $\mu$ is the hypothesizes population mean, $s$ is the sample standard deviation, and $n$ is the sample size.

Writing this in R:

```{r, echo=TRUE}
get_t_stat <- function(x_bar, mu, sd, n){
  return((x_bar - mu) / (sd / sqrt(n)))
}

```


Find the t-statistic:

```{r, echo=TRUE}
t_stat <- get_t_stat(x_bar = 410, mu = 500, sd = 90, n = 9)
```

### A

Two-tailed test

```{r, echo=TRUE}
n = 9
pval_two_tail = 2*pt(t_stat, df = n-1)
pval_two_tail
```

We can reject the hypothesis that population mean is 500.

### B

```{r, echo=TRUE}
pval_lower_tail = pt(t_stat, df = n-1)
pval_lower_tail
```

We can reject the hypothesis that population mean is greater than 500.

### C

```{r, echo=TRUE}
pval_upper_tail = pt(t_stat, df = n-1, lower.tail=FALSE)
pval_upper_tail
```

We fail to reject the hypothesis that population mean is less than 500.

Alternatively for C, we could just subtract the answer in B from 1:

```{r, echo=TRUE}
1 - pval_lower_tail
```


## Question 5

```{r, echo=TRUE}
t_jones = ((519.5 - 500)/ 10)
t_smith = ((519.7 - 500)/ 10)
cat("t value for Jones:", t_jones, '\n')
cat("t value for Smith:", t_smith, '\n')
cat('p value for Jones:', round(2*pt(t_jones, df = 999, lower.tail=FALSE), 4), '\n')
cat('p value for Smith:', round(2*pt(t_smith, df = 999, lower.tail=FALSE), 4), '\n')
```

At 0.05 level Smith's result is statistically significant but Jones's is not.
The result show the arbitrariness of the 0.05 demarcation line and the importance of reporting actual p-values to better make sense of results.

## Question 6:

```{r, echo=TRUE}
# Creating the dataframe
grade_level <- c(rep("6th grade", 100), rep("7th grade", 100), rep("8th grade", 100))
snack <- c(rep("healthy snack", 31), rep("unhealthy snack", 69), rep("healthy snack", 43),
           rep("unhealthy snack", 57), rep("healthy snack", 51), rep("unhealthy snack", 49))
snack_data <- data.frame(grade_level, snack)
```

We are conducting a Chi-square test in this question since we are testing the association between two categorical variables.  

```{r,echo=TRUE}
table(snack_data$snack,snack_data$grade_level)
chisq.test(snack_data$snack,snack_data$grade_level,correct = FALSE)
```

A p-value smaller than 0.05 indicates that there is a relationship between grade level and the choice of snack. 

## Question 7:

```{r, echo=TRUE}
# Creating the dataframe
Area <- c(rep("Area1", 6), rep("Area2", 6), rep("Area3", 6))
cost <- c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5, 7.5, 8.2, 8.5, 8.2, 7.0, 9.3,
          5.8, 6.4, 5.6, 7.1, 3.0, 3.5)
Area_cost <- data.frame(Area,cost)
```

Since we are comparing the means of more than two groups, we are using the ANOVA test in this question. 

```{r, echo=TRUE}
one.way <- aov(cost ~ Area, data = Area_cost)
summary(one.way)
```

The small p-value tells us that the three areas have a difference in means. 