---
title: "Homework 2"
author: "Justine Shakespeare"
description: "Homework 2 for DACSS 603"
date: "03/27/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - hypothesis testing
  - confidence intervals
---

## Question 1

*The time between the date a patient was recommended for heart surgery
and the surgery date for cardiac patients in Ontario was collected by
the Cardiac Care Network ("Wait Times Data Guide," Ministry of Health
and Long-Term Care, Ontario, Canada, 2006). The sample mean and sample
standard deviation for wait times (in days) of patients for two cardiac
procedures are given in the accompanying table. Assume that the sample
is representative of the Ontario population.*

*Construct the 90% confidence interval to estimate the actual mean wait
time for each of the two procedures. Is the confidence interval narrower
for angiography or bypass surgery?*

First we'll calculate the confidence interval manually with the formula:
CI = (X bar) ± (t × s/sqrt(n))

We'll create objects in R to plug into this formula. We'll start with
the Bypass data.

```{r, echo=T}
library(tidyverse)

# Bypass data
BypassSample <- 539
BypassMean <- 19
BypassSD <- 10

# standard error
BypassSEM <- BypassSD/sqrt(BypassSample)

# t-value
ConfidenceLevel <- .90
BYPtail <- (1-ConfidenceLevel)/2
Bypass_t_score <- qt(1-BYPtail, df=BypassSample-1)

# plug everything back into the CI formula:
CIBypass <- c(BypassMean-Bypass_t_score*BypassSEM, 
              BypassMean+Bypass_t_score*BypassSEM)

print(CIBypass)
```

Now let's do the same for the Angiography data.

```{r}
AngioSample <- 847
AngioMean <- 18
AngioSD <- 9

# standard error
AngioSEM <- AngioSD/sqrt(AngioSample)

# t-value
ConfidenceLevel <- .90
Angtail <- (1-ConfidenceLevel)/2
Angio_t_score <- qt(1-Angtail, df=AngioSample-1)

# plug everything back in

CIAngio <- c(AngioMean-Angio_t_score*AngioSEM, 
             AngioMean+Angio_t_score*AngioSEM)

print(CIAngio)
```

Now we can compare the confidence intervals.

```{r}
abs(diff(CIAngio)) 

abs(diff(CIBypass))
```

The confidence interval for Angiography is narrower than for Bypass.

## Question 2

*A survey of 1031 adult Americans was carried out by the National Center
for Public Policy. Assume that the sample is representative of adult
Americans. Among those surveyed, 567 believed that college education is
essential for success. Find the point estimate, p, of the proportion of
all adult Americans who believe that a college education is essential
for success. Construct and interpret a 95% confidence interval for p.*

We can use the prop.test() function to find the point estimate and the
confidence interval:

```{r}
prop.test(x=567, n=1031, conf.level = .95)

```

## Question 3

*Suppose that the financial aid office of UMass Amherst seeks to
estimate the mean cost of textbooks per semester for students. The
estimate will be useful if it is within \$5 of the true population mean
(i.e. they want the confidence interval to have a length of \$10 or
less). The financial aid office is pretty sure that the amount spent on
books varies widely, with most values between \$30 and \$200. They think
that the population standard deviation is about a quarter of this range
(in other words, you can assume they know the population standard
deviation). Assuming the significance level to be 5%, what should be the
size of the sample.*

We have to solve for n here. Because we have the standard deviation,
we'll use the CI formula that uses the z-score: CI = (X bar) ± (z ×
s/sqrt(n))

This can be rewritten: n=((s\*z)/5)\^2

Using the data we have:

```{r}
Q3_s = (200-30)/4
Q3_z = qnorm(.975)

((Q3_s*Q3_z)/5)^2
```

## Question 4

*According to a union agreement, the mean income for all senior-level
workers in a large service company equals 500 per week. A representative
of a women's group decides to analyze whether the mean income μ for
female employees matches this norm. For a random sample of nine female
employees, ȳ = \$410 and s = 90*

### 4.A

*Test whether the mean income of female employees differs from \$500 per
week. Include assumptions, hypotheses, test statistic, and P-value.
Interpret the result.*

To solve this problem I followed the steps of hypothesis testing:

1\. Specify the null hypothesis and alternative hypothesis

2\. Set the significance level (alpha)

3\. Calculate the test statistic

4\. Compare test statistics to the critical value determined by alpha,
or check if p-value is smaller than alpha

5\. Draw a conclusion

1.  Our null hypothesis is that the population mean is 500 dollars, our
    alternative hypothesis is that the population mean is NOT \$500

2.  The significance level (alpha) is .05, or 5%

3.  Calculate the test statistic:

```{r}
Q4_SD <- 90
Q4_ymean <- 410
Q4_umean <- 500
Q4_s_size <- 9
Q4_SEM <- Q4_SD/sqrt(Q4_s_size)

Q4_T_Statistic <- (Q4_ymean - Q4_umean)/(Q4_SEM)
Q4_T_Statistic
```

4.  Calculate the critical value so that we can compare it with the test
    statistic.

```{r}
Q4_Crit_Value <- qt(0.025, df=(Q4_s_size-1))
Q4_Crit_Value
```

The critical value (on the lower tail end) is -2.3060. Because the test
statistic is smaller than the critical value (and larger in absolute
terms), we can conclude that we should reject the null hypothesis.

4.  (continued) calculate the p-value to check to see if it is smaller
    than alpha.

```{r}
Q4_p_value <- pt(Q4_T_Statistic, df = (Q4_s_size-1), lower.tail = TRUE)*2
Q4_p_value
```

The p-value is 0.01707, which is smaller than our alpha (.05). 0.01707
\< 0.05 This again means that we can reject the null hypothesis.

5.  Draw a conclusion

Because the test statistic (-3) is in absolute terms larger than the
critical values (2.30) AND because the p-value (0.01707) is less than
the alpha levelof .05, we can reject the null hypothesis

Note that when we calculate the confidence interval of these values we
can also see that the estimated population mean here is outside of a 95%
CI.

To calculate the confidence level we'll first create some objects we'll
need in our formulas:

```{r}
ConfLevel <- 0.95
Q4_tail_area <- (1-ConfLevel)/2
Q4_tscore <- qt(p=1-Q4_tail_area, df = Q4_s_size-1)

# Then, using this and some of the objects we created earlier we'll plug everything in: 
Q4_CI <- c(Q4_ymean - (Q4_tscore * Q4_SEM), Q4_ymean + (Q4_tscore * Q4_SEM))

print(Q4_CI)
```

We can see here that \$500 is not within the 95% confidence interval,
which is another reason we can reject the null hypothesis.

### 4.B

*Report the P-value for Ha: μ \< 500. Interpret.*

To calculate a one sided P-value you do not need to multiply the formula
by 2 and you need to specify that we're only looking at one side. An
alternative hypothesis of μ \< 500 means the null hypothesis is μ \>=
500. We need to specify that we are looking at the lower tail.

```{r}
Q4_p_value1 <- pt(Q4_T_Statistic, df = (Q4_s_size-1), lower.tail = TRUE)
Q4_p_value1
```

The p-value here is even lower than it was for the two sided test, it is
0.0085. We can definitely reject the null hypothesis that the true
population mean is over \$500.

### 4.C

*Report and interpret the P-value for Ha: μ \> 500.*

To calculate the P-value for Ha: μ \> 500 we run the same formula as
before but we specify that we're *not* interested in the lower tail.

```{r}
Q4_p_value2 <- pt(Q4_T_Statistic, df = (Q4_s_size-1), lower.tail = FALSE)
Q4_p_value2
```

The p-value for this one sided test is quite large. It means that we
should not reject the null hypothesis that the true population mean is
less than \$500.

To test these p-values we can add them up.

```{r}
Q4_p_value1 + Q4_p_value2
```

We find that they add up to 1.

## Question 5

*Jones and Smith separately conduct studies to test H0: μ = 500 against
Ha: μ ≠ 500, each with n = 1000. Jones gets ȳ = 519.5, with se = 10.0.
Smith gets ȳ = 519.7, with se = 10.0.*

### 5.A

*Show that t = 1.95 and P-value = 0.051 for Jones. Show that t = 1.97
and P-value = 0.049 for Smith.*

First we can create some objects to represent the values for this
question. Then we can use those to get the test statistic and the
p-value for Jones and Smith.

```{r}
Q5_SEM <- 10 # they have the same standard error
Jones_ymean <- 519.5
Smith_ymean <- 519.7
Q5_s_size <- 1000 # they have the same sample size

Q5_umean <- 500

# calculate the test statistic and p-value for Jones:
Jones_T_Statistic <- (Jones_ymean - Q5_umean)/(Q5_SEM)
Jones_T_Statistic

# we find that Jones' test statistic is 1.95. We can use that to find the p-value:
Jones_p_value <- pt(Jones_T_Statistic, df = (Q5_s_size-1), lower.tail = FALSE)*2
Jones_p_value

# Calculate test statistic and p-value for Smith:
Smith_T_Statistic <- (Smith_ymean - Q5_umean)/(Q5_SEM)
Smith_T_Statistic

# we find that Smith's test statistic is 1.97. We can use that to find the p-value:
Smith_p_value <- pt(Smith_T_Statistic, df = (Q5_s_size-1), lower.tail = FALSE)*2
Smith_p_value
```

### 5.B

*Using α = 0.05, for each study indicate whether the result is
"statistically significant."*

Above we found the p-value for both Jones and Smith, we can use that to
show how those values compare to alpha set at 0.05.

```{r}
alpha <- 0.05

Jones_p_value < alpha

Smith_p_value < alpha
```

This shows that technically Jones' p-value is not statistically
significant because it is more than the alpha of 0.05, but Smith's
p-value is technically statistically significant because it is less than
the alpha.

### 5.C

*Using this example, explain the misleading aspects of reporting the
result of a test as "P ≤ 0.05" versus "P \> 0.05," or as "reject H0"
versus "Do not reject H0," without reporting the actual P-value.*

This question demonstrates how close two results can be and still
technically not have the same significance level. In these cases it is
important to include the p-value so that readers can put the results
into context.

## Question 6

*A school nurse wants to determine whether age is a factor in whether
children choose a healthy snack after school. She conducts a survey of
300 middle school students, with the results below. Test at α = 0.05 the
claim that the proportion who choose a healthy snack differs by grade
level. What is the null hypothesis? Which test should we use? What is
the conclusion?*

Since we're dealing with categorical variables (grade and the choice of
a healthy or unhealthy snack) we should use a chi-square test. First
we'll create the dataframe we'll need to run the test in r.

```{r}
obs <- matrix(c(31, 69, 43, 57, 51, 49), nrow=3, byrow = TRUE)
dimnames(obs) <- list(Grade = c("Grade_6", "Grade_7", "Grade_8"),
                      Snack = c("healthy", "unhealthy"))

obs
```

Then we can run the chi-square test and print the result:

```{r}
ChiResult <- chisq.test(obs)

print(ChiResult)
```

Given the p-value \< 0.05, we can reject the null hypothesis that grade
level does not have an effect on the choice of healthy snack. The
relatively high X-squared value also indicates that there is a
meaningful difference in the choices made by kids in different grade
levels.

## Question 7

*Per-pupil costs (in thousands of dollars) for cyber charter school
tuition for school districts in three areas are shown. Test the claim
that there is a difference in means for the three areas, using an
appropriate test. What is the null hypothesis? Which test should we use?
What is the conclusion?*

The null hypothesis is that all three districts have equal average
tuition. We can use the ANOVA test since we are comparing three means.

First we'll create the dataframe in r:

```{r}
Area1 <- c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5)  
Area2 <- c(7.5, 8.2, 8.5, 8.2, 7.0, 9.3)
Area3 <- c(5.8, 6.4, 5.6, 7.1, 3.0, 3.5)

Q7df <- data.frame(Area1, Area2, Area3)

Q7df_PL <- pivot_longer(Q7df, cols = c(Area1, Area2, Area3), names_to = "district", values_to = "tuition")
Q7df_PL
```

Now that we have a dataframe with one continuous variable and one
categorical variable, we can use the anova test function in r `aov()` to
test our hypothesis. This type of hypothesis test is appropriate for
this scenario because we want to compare the mean of three groups.

```{r}
Q7o <- aov(tuition ~ district, data = Q7df_PL)
summary(Q7o)
```

Given that the p value here is less than 0.01, we can reject the null
hypothesis that the mean tuition is the same at all three schools.
