---
title: "Homework 2"
author: "Xiaoyan"
description: "Template of course blog qmd file"
date: "03/24/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - desriptive statistics 
  - probability
---
```{r}
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
```
# Question 1
1. The time between the date a patient was recommended for heart surgery and the surgery date for cardiac patients in Ontario was collected by the Cardiac Care Network (“Wait Times Data
Guide,” Ministry of Health and Long-Term Care, Ontario, Canada, 2006). The sample mean and sample standard deviation for wait times (in days) of patients for two cardiac procedures are given in the accompanying table. Assume that the sample is representative of the Ontario population 

Construct the 90% confidence interval to estimate the actual mean wait time for each of the two procedures. Is the confidence interval narrower for angiography or bypass surgery?
```{r}

# mean mean()
Mean1<-19
Mean2<-18
# standard deviation sd()
sd1<-10
sd2<-9
# sample size
n1<-539
n2<-847
# standard error sd/sqrt(n)
se1<-sd1/sqrt(n1)
se2<-sd2/sqrt(n2)
se1
se2
# t-value

# tail_area<-(1-confidence_level)/2
TA<-(1-0.9)/2
#t score <-qt(p=1-tail_area,df=s_size-1 )
tscore1<-qt(p=1-TA, df = n1-1)
tscore2<-qt(p=1-TA, df = n2-1)
tscore1
tscore2
#CI <-c(s_mean-t_score*SE,s_mean+t_score*SE )
CI1<-c(Mean1-tscore1*se1, Mean1+tscore1*se1)
CI2<-c(Mean2-tscore2*se2, Mean2+tscore2*se2)
CI1
CI2
DiffCI1<-18.29029- 19.70971
DiffCI2<-17.49078-18.50922
DiffCI1
DiffCI2
```


# Question 2

A survey of 1031 adult Americans was carried out by the National Center for Public Policy. Assume that the sample is representative of adult Americans. Among those surveyed, 567 believed that college education is essential for success. Find the point estimate, p, of the proportion of all adult Americans who believe that a college education is essential for success.
Construct and interpret a 95% confidence interval for p.
```{r, echo=T}
#p 
p1<-567/1031
#sample size 
n<-1031
#alpha = 0.05
se<-sqrt((p1*(1-p1))/n)
#z score1.96
CI<-c(p1-1.96*se,p1+1.96*se)
CI

```

# Question 3
Suppose that the financial aid office of UMass Amherst seeks to estimate the mean cost of textbooks per semester for students. The estimate will be useful if it is within $5 of the true population mean (i.e. they want the confidence interval to have a length of $10 or less). The financial aid office is pretty sure that the amount spent on books varies widely, with most values between $30 and $200. They think that the population standard deviation is about a quarter of this range (in other words, you can assume they know the population standard deviation).Assuming the significance level to be 5%, what should be the size of the sample
```{r, echo=T}
#z*sd/sqrt(n)=5
#population sd = (200-30)/4 = 42.5
#alpha = 0.05
#n=(z*sd/5)^2
#SD = 5
n<-(1.96*42.5/5)^2

n

qnorm(0.975)
```
# Question 4
According to a union agreement, the mean income for all senior-level workers in a large service company equals $500 per week. A representative of a women’s group decides to analyze whether the mean income μ for female employees matches this norm. For a random sample of nine female employees, ȳ = $410 and s = 90
A. Test whether the mean income of female employees differs from $500 per week. Include assumptions, hypotheses, test statistic, and P-value. Interpret the result.
```{r}
#assumptions
#hypothesis
  #h0:u=500
  #ha:u≠500
#t-test
u4<-500
y4<-410
sd4<-90
n4<-9

t4<-(y4-u4)/(sd4/sqrt(n4))
t4
#pvalue
p4= 2*pt(t4, df = n4-1)
p4
#P3<- P-value = P(|t| > |t3|) = P(t < -t3) + P(t > t3)

```
reject the null hypothesis

B. Report the P-value for Ha: μ < 500. Interpret.
B.
```{r}
#hypothesis
  #h0:u>=500
  #ha:u<500
p4.2<-pt(t4, df = n4-1)
p4.2
```
reject ha
C. Report and interpret the P-value for Ha: μ > 500.
(Hint: The P-values for the two possible one-sided tests must sum to 1.

```{r}
#hypothesis
  #h0:u<=500
  #ha:u>500
p4.3<-pt(t4, df = n4-1)
1-p4.3
```



# Question5.
Jones and Smith separately conduct studies to test H0: μ = 500 against Ha: μ ≠ 500, each
with n = 1000. Jones gets ȳ = 519.5, with se = 10.0. Smith gets ȳ = 519.7,
with se = 10.0.
A. Show that t = 1.95 and P-value = 0.051 for Jones. Show that t = 1.97 and P-value = 0.049
for Smith.
```{r}
#H0: μ = 500, Ha: μ ≠ 500
u5<-500
n5<-1000
n5<-1000
y51<-519.5
se51<-10
y52<-519.7
se52<-10

t51<-(y51-u5)/se51
t51
t52<-(y52-u5)/se52
t52
p51<-2*pt(t51, df=n5-1,lower.tail = F)
p52<-2*pt(t52, df=n5-1,lower.tail = F)
p51
p52
```
why lower tail is false here?

B. Using α = 0.05, for each study indicate whether the result is “statistically significant.”

Jones is not significant but Smith's is

C. Using this example, explain the misleading aspects of reporting the result of a test as “P ≤
0.05” versus “P > 0.05,” or as “reject H0” versus “Do not reject H0,” without reporting the actual
P-value.

It is important to have a significant level at 0.05 and report the accurate and true p-value to evaluate the siginifance. 

# Question6.
A school nurse wants to determine whether age is a factor in whether children choose a
healthy snack after school. She conducts a survey of 300 middle school students, with the results
below. Test at α = 0.05 the claim that the proportion who choose a healthy snack differs by grade
level. What is the null hypothesis? Which test should we use? What is the conclusion?
Grade level 6th grade 7th grade 8th grade
Healthy snack 31 43 51
Unhealthy snack 69 57 49
hypothesis:there is a relationship between chosing snacks and grade
```{r}
grade_level <- c(rep("6th grade", 100), rep("7th grade", 100), rep("8th grade", 100))
snack <- c(rep("healthy snack", 31), rep("unhealthy snack", 69), rep("healthy snack", 43),
           rep("unhealthy snack", 57), rep("healthy snack", 51), rep("unhealthy snack", 49))
snack_data <- data.frame(grade_level, snack)
snack_data
chisq.test(snack_data$snack,snack_data$grade_level,correct = FALSE)
```
p value smaller than 0.05 means there is a relationship between grade and snack choice.

# Question7.
Per-pupil costs (in thousands of dollars) for cyber charter school tuition for school
districts in three areas are shown. Test the claim that there is a difference in means for the three
areas, using an appropriate test. What is the null hypothesis? Which test should we use? What is
the conclusion?
Area 1 6.2 9.3 6.8 6.1 6.7 7.5
Area 2 7.5 8.2 8.5 8.2 7.0 9.3
Area 3 5.8 6.4 5.6 7.1 3.0 3.5
```{r}
Area <- c(rep("Area1", 6), rep("Area2", 6), rep("Area3", 6))
cost <- c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5, 7.5, 8.2, 8.5, 8.2, 7.0, 9.3,
          5.8, 6.4, 5.6, 7.1, 3.0, 3.5)
Area_cost <- data.frame(Area,cost)
Area_cost
summary(aov(cost~Area, data=Area_cost))
```