---
title: "Homework 2"
author: "Emma Narkewicz"
description: "Template of course blog qmd file"
date: "02/28/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - confidence_interval
  - emma_narkewicz
editor: 
  markdown: 
    wrap: 72
---

```{r}
library(tidyverse)
```


# Question 1

I constructed the 90% confidence interval of each of the two heart
procedures manually using the equation: *CI = (X bar) ± (t × s/sqrt(n))*
and the mean wait time, sample standard deviation, and sample size
provided for the Bypass and Angiography heart procedures.

For both procedures I calculated the t-score based on 2 tails, as there
was no information indicating we should only focus on the upper or lower
tail.

```{r}
# Calculating 90% CI  manually of Bypass heart surgery
## sample mean wait time
s_mean_bypass <- 19
## sample sd
s_sd_bypass <- 10
## sample size (n)
s_size_bypass <- 539
## standard error  (sd/swrt(n))
s_se_bypass <- s_sd_bypass / sqrt (s_size_bypass)
s_se_bypass
## Specify confidence level 
confidence_level_bypass <- 0.9
##Calculate tail area of 2-tail t-test
tail_area_bypass <- (1 - confidence_level_bypass)/2
tail_area_bypass
#calculation t-score using qt
t_score_bypass <-  qt(p = 1 - tail_area_bypass, df = s_size_bypass - 1)
t_score_bypass
##Calculating upper & lower CI bypass
CI_bypass <- c(s_mean_bypass - t_score_bypass * s_se_bypass, s_mean_bypass + t_score_bypass * s_se_bypass)
print(CI_bypass)
```

-   *The 90% confidence interval for the Bypass wait time is 18.29
    days - 19.71 days*

```{r}
#Calculating the 90% CI manually of angiography heart surgery
##sample mean wait time
s_mean_angio <- 18
##sample sd
s_sd_angio <- 9
## sample size (n)
s_size_angio <- 847
## standard error  (sd/swrt(n)
s_se_angio <- s_sd_angio / sqrt (s_size_angio)
s_se_angio
## Specify confidence level 
confidence_level_angio <- 0.9
##Calculate tail area of 2-tail t-test
tail_area_angio <- (1 - confidence_level_angio)/2
tail_area_angio
#calculation t-score using qt
t_score_angio <-  qt(p = 1 - tail_area_angio, df = s_size_angio - 1)
t_score_angio
##Calculating upper & lower CI bypass
CI_angio <- c(s_mean_angio - t_score_angio * s_se_angio, s_mean_angio + t_score_angio * s_se_angio)
print(CI_angio)
```

-   *The 90% confidence interval for the Angiography wait time is 17.49
    days - 18.51 days*

```{r}
#narrower CI
delta_CI_bypass = 19.71 - 18.29
delta_CI_bypass
delta_CI_angio = 18.51 - 17.49
delta_CI_angio
```

-   *The confidence interval is narrower for the Angiography (CI delta
    of 1.02 days) than the confidence interval for the Bypass (CI delta
    of 1.42 days)*

# Question 2

To calculate & interpret a 95% CI for the proportion of all adult
Americans who believe that a college education is essential for success
I used prop.test() because this involves estimating a proportion, not a
probability. In this test, a "success" is defined as an adult American
reporting a college education is essential for success, with the
alternative being a response of an adult American reporting a college
education is not essential for success.

The point estimate for the proportion of adult Americans who believe
that college is essential is 0.55.

```{r}
#Calculate point estimate college is essential

sample_size_survey <- 1031
point_estimate_college_essential <- 567/sample_size_survey
point_estimate_college_essential

```

Specific information plugged in was:

-   sample proportion p = point estimate (0.5499515),
-   number sample "successes" x = 567
-   sample size n = 1031

```{r}
#Performing prop.test
prop.test(x = 567, n = 1031, p = point_estimate_college_essential)
```

The 95% confidence interval is \[0.52, 0.58\]

This should be interpreted as 95% of confidence intervals calculated
with this procedure would contain the true proportion of adult Americans
who believe that a college education is essential for success.

# Question 3

Suppose that the financial aid office of UMass Amherst seeks to estimate
the mean cost of textbooks per semester for students. The estimate will
be useful if it is within \$5 of the true population mean (i.e. they
want the confidence interval to have a length of \$10 or less). The
financial aid office is pretty sure that the amount spent on books
varies widely, with most values between \$30 and \$200. They think that
the population standard deviation is about a quarter of this range (in
other words, you can assume they know the population standard
deviation).

Assuming the significance level to be 5%, what should be the size of the
sample?

Because we are assuming we know the population sd, this signals we
should a 2-sided z-test to calculate the sample size. I can solve for
sample size using the z-score through the Margin of Error (MOE) equation
that:

MOE = zscore @ confidence level * (sqrt(sd\^2/n))

rearranging the equation to solve for n by squaring both sides,
multiplying by n, and then dividing by MOE\^2 results in the equation:

n = (z\^2 \* sd\^2)/ (MOE\^2)

-   MOE = +/- 5
-   pop sd = (200-30)/4 = 42.5
-   at an alpha level 0.05, 2-sided z-score = 1.959964

```{r}
#solving for sample size

n = ((1.959964^2) * (42.5^2)) / (5^2) 
n

```

*The ideal sample size is 278*

# Question 4

According to a union agreement, the mean income for all senior-level
workers in a large service company equals \$500 per week. A
representative of a women's group decides to analyze whether the mean
income μ for female employees matches this norm. For a random sample of
nine female employees, ȳ = \$410 and s = 90

## A

Test whether the mean income of female employees differs from \$500 per
week. Include assumptions, hypotheses, test statistic, and P-value.
Interpret the result.

Assumptions of a 1-sample t-test:

-   population is normally distributed

-   observations in our sample are generated independently of one
    another

Information we are provided is that:

-   Population mean (μ) = 500

-   Sample mean (y bar) = 410

-   Sample sd = 90 s

-   Sample n = 9

Null Hypothesis: (H0) female mean income μ = 500 Alternative Hypothesis:
(Ha) female mean income μ ≠ 500

Calculate the t statistic using the equation:

*t = (y bar - μ ) / (sd estimate / sqrt(N))*

```{r}
## Calculating tscore

t =  (410 - 500) / (90/sqrt(9))
t
```

The t-score is -3

To solve for the p-value from the t-score I will use the pt() function,
with a q = the t-statistic of -3, df = 8 (9-1).

It it is important to note that an assumption of the pt() function is
you are looking for the probability of the lower tail, not both tails.
Because the t score is negative looking at the lower tail is
appropriate, but to get the probability for 2-tailed test we need to
multiply the lower tail probability by 2.

```{r}
#Calculating p-value

p_lower_tail = pt(q = t, df =8, lower.tail = TRUE)
p_lower_tail


```

This gives us the probability that μ is less than 500, but that is only
1 tail, so we multiply this problity by 2 to get the p value for our
2-sided t-test

```{r}
#Two-side t
p_twotail =  p_lower_tail * 2
p_twotail

```

With a 2-tail p value of 0.017, we can confidently reject the null
hypothesis that female salary μ = 500 as p\< 0.05. This suggests that
the true mean salary of female workers is not equal to \$500.

## B
 Report the P-value for Ha: μ \< 500. Interpret.

```{r}
#P-value Ha: μ < 500

p_lower_tail = pt(q = t, df =8, lower.tail = TRUE)

p_lower_tail

```

With a small p-value of 0.00854 we reject the null hypothesis that μ =
500, suggesting that the true mean salary of female workers is less than
$500. 

## C
 Report and interpret the P-value for Ha: μ \> 500.

```{r}
#P-value Ha: μ > 500

p_upper_tail = pt(q = t, df =8, lower.tail = FALSE)

p_upper_tail
```

With a large p-value of 0.99 we fail to reject the null hypothesis that
μ = 500, suggesting that the true means salary of female workers is not
greater than \$500.

# Question 5

Jones and Smith separately conduct studies to test H0: μ = 500
against Ha: μ ≠ 500, each with n = 1000. Jones gets ȳ = 519.5, with
se = 10.0. Smith gets ȳ = 519.7,with se = 10.0.

## A

Show that Show that t = 1.95 and P-value = 0.051 for Jones. Show that t
= 1.97 and P-value = 0.049 for Smith.

Because we are provided with standard error not standard deviation, I
used the equation:

t = (ȳ - μ) / se

```{r}
#Jones T score

t_Jones = (519.5 - 500) / 10.0
t_Jones

```

*For Jones the t score is indeed 1.95*

To calculate the p-value, I will use pt(), q = 1.95, df = (1000-1) =
999). Because Jone's t-score is positive, we look at the upper tail.

```{r}
#Calculate 1-tail p-value Jones
p_upper_tail_Jones = pt(q = t_Jones, df = 999, lower.tail = FALSE)
p_upper_tail_Jones

```

The p value of the upper tail for Jones is 0.02572777, but because the
alternative hypothesis is 2-tailed, it needs to be multiplied by 2

```{r}
#Calculating p_value Jones
p_Jones = p_upper_tail_Jones * 2
p_Jones

```

*This shows the p-value of Jones is indeed 0.51*

```{r}
# Smith T score 
t_Smith = (519.7 - 500) / 10.0
t_Smith
```

*The t-score for Smith is indeed 1.97.* The p value is calculated the
same way as for Jones, using pt() & the upper tail:

```{r}
#Calculate 1-tail p-value Smith
p_upper_tail_Smith = pt(q = t_Smith, df = 999, lower.tail = FALSE)
p_upper_tail_Smith
```

```{r}
#Smith p=value
p_Smith = p_upper_tail_Smith * 2
p_Smith
```

*Smith's p-value does = 0.049*

## B

Using an alpha level = 0.05 for both

-   Jones' p-value = 0.51 means a non-statistically significant result,
    where we fail to reject the null hypothesis that μ = 500\
-   Smith's p=value = 0.49 means a statistically significant results,
    where we reject the null hypothesis that μ = 500

## C

This example showcases how reporting the actual p-value is important, as
in this case the p-values are very close between Jones & Smith, which
are both right on the edge of significance (0.050 +/- 0.01). A p-value
of 0.049 & a p-value of 0.0000000049 are both less than 0.05 & would
both lead us to reject the null hypothesis, but the second p-value is
highly significant while the first p-value is only barely significant at
an alpha level of 0.05. Reporting the p-value itself gives readers &
other researchers information to much better understand your statistical
findings than just reporting if the p value is \> or \< 0.05 or if you
do or do not reject H0.

# Question 6

A school nurse wants to determine whether age is a factor in whether
children choose a healthy snack after school. She conducts a survey of
300 middle school students, with the results below. Test at α = 0.05 the
claim that the proportion who choose a healthy snack differs by grade
level. What is the null hypothesis? Which test should we use? What is
the conclusion?

-   The null hypothesis is that there is no association between the
    grade of the student and if they choose a healthy snack (they are independent). In this
    question independence would result in the proportion of students in each
    grade who choose a healthy snack is the same: H0: p6th = p7th = p8th

-   Because I want to test if there is an association between 2
    categorical variables, snack preference & grade level, I used a
    chi-squared test.

First I recreated the HW2 snacks data table in R:

```{r}
#Recreating nurse data

Health_tab <- matrix(c(31, 43, 51, 69, 57, 49), nrow=2, byrow=TRUE)
colnames(Health_tab) <- c('6th','7th','8th')
rownames(Health_tab) <- c('Healthy','Unhealthy')
Health_tab <- as.table(Health_tab)

Health_tab

```

Then a chi-squared was run on the data set:

```{r}
#Chi Squared test

chisq.test(Health_tab)

```

-   Conclusion : The p-value of 0.015 from the chi-squared test is less
    than alpha=0.05, meaning we reject the null hypothesis that there is
    no association between student grade level and preference for
    healthy snacks.

-   At a significance level of 0.05 and df = 2, the critical value =
    5.991. The X-squared = 8.3383 is greater than the critical value,
    once again supporting rejecting the null hypothesis. This suggests
    snack preference is not independent of grade level.

# Question 7

Per-pupil costs (in thousands of dollars) for cyber charter school
tuition for school districts in three areas are shown. Test the claim
that there is a difference in means for the three areas, using an
appropriate test. What is the null hypothesis? Which test should we use?
What is the conclusion?

-   The null hypothesis is that there is no difference between the mean
    per-pupil cost for cyber charter tuition in 3 different areas, which
    can be written out as: 

H0: μ1 = μ2 = μ3

-   Because we are testing if there is a difference between the means of
    3 or more groups that differ by the categorical variable of area we
    should use a ANOVA test

-   I entered in the data below manually, and then pivoted longer to get
    2 columns, Area and Cost.

```{r}
Area_1 <- c( 6.2, 9.3, 6.8, 6.1, 6.7, 7.5)
Area_2 <- c( 7.5, 8.2, 8.5, 8.2, 7.0, 9.3)
Area_3 <- c( 5.8, 6.4, 5.6, 7.1, 3.0, 3.5)

charter_data <- data.frame (Area_1, Area_2, Area_3)

```

```{r}
#Pivoted Longer

pivot_charter <- charter_data %>%
  pivot_longer(c(Area_1, Area_2, Area_3 ), names_to= "Area", values_to="Cost")

pivot_charter

```

Lastly, I ran the ANOVA, specifying Area as the independent variable &
Cost (per pupil) as the dependent variable.

```{r}
#Anova 
anova_charter <- aov( Cost ~ Area, data = pivot_charter)

summary(anova_charter)


```

The resulting p-value of 0.00397 is less than an alpha = 0.05 and
therefor significant, leading us to reject out null hypothesis that
there is no difference in the mean per-pupil charter school cost between
the 3 areas. There does appear to be a difference in per-pupil charter
school cost based on which of the 3 areas the school is in.
