---
title: "HW2"
author: "Liam Tucksmith"
desription: "HW2"
date: "03/12/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - Liam Tucksmith
  - LungCapData
  - tidyverse
  - readxl
  - ggplot2
  - dplyr
  - tidyr
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=T}
#| label: setup
#| warning: false

library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=T}
df <- read_excel("_data/LungCapData.xls")
```

1.  The time between the date a patient was recommended for heart
    surgery and the surgery date for cardiac patients in Ontario was
    collected by the Cardiac Care Network ("Wait Times Data Guide,"
    Ministry of Health and Long-Term Care, Ontario, Canada, 2006). The
    sample mean and sample standard deviation for wait times (in days)
    of patients for two cardiac procedures are given in the accompanying
    table. Assume that the sample is representative of the Ontario
    population

Construct the 90% confidence interval to estimate the actual mean wait
time for each of the two procedures. Is the confidence interval narrower
for angiography or bypass surgery?

```{r, echo=T}
bp_num <- 539
bp_mean <- 19
bp_sd <- 10
bp_se <- sqrt(bp_sd)
ag_num <- 847
ag_mean <- 18
ag_sd <- 9
ag_se <- sqrt(ag_sd)

alpha = 0.1
degrees.freedom = bp_num - 1
t.score = qt(p=alpha/2, df=degrees.freedom,lower.tail=F)

margin.error <- t.score * bp_se

lower.bound <- bp_mean - margin.error
upper.bound <- bp_mean + margin.error
bypass_confidenceInterval <- print(c(lower.bound,upper.bound))

degrees.freedom = ag_num - 1
t.score = qt(p=alpha/2, df=degrees.freedom,lower.tail=F)

margin.error <- t.score * ag_se

lower.bound <- ag_mean - margin.error
upper.bound <- ag_mean + margin.error
angiography_confidenceInterval <- print(c(lower.bound,upper.bound))
```

The confidence interval is narrower for angiography surgery than it is
for bypass surgery. The bypass surgery confidence interval spans from
13.78954 to 24.21046 days while the angiography surgery spans from
13.06003 to 22.93997 days. Since the bypass surgery confidence interval
spans 10.42 days and the angiography surgery spans 9.88 days, the
angiography surgery has a narrower confidence interval than the bypass
surgery.

2.  A survey of 1031 adult Americans was carried out by the National
    Center for Public Policy. Assume that the sample is representative
    of adult Americans. Among those surveyed, 567 believed that college
    education is essential for success. Find the point estimate, p, of
    the proportion of all adult Americans who believe that a college
    education is essential for success. Construct and interpret a 95%
    confidence interval for p.

```{r}
n = 1031 
college_y = 567
p = college_y/n
print(p)

margin_error <- qnorm(0.95)*sqrt(p*(1-p)/n)
lowerbound <- p - margin_error
upperbound <- p + margin_error

p_confidenceInterval <- print(c(lowerbound,upperbound))
```

3.  Suppose that the financial aid office of UMass Amherst seeks to
    estimate the mean cost of textbooks per semester for students. The
    estimate will be useful if it is within \$5 of the true population
    mean (i.e. they want the confidence interval to have a length of
    \$10 or less). The financial aid office is pretty sure that the
    amount spent on books varies widely, with most values between \$30
    and \$200. They think that the population standard deviation is
    about a quarter of this range (in other words, you can assume they
    know the population standard deviation). Assuming the significance
    level to be 5%, what should be the size of the sample

```{r}
margin_error_max = 5
confidence_interval_length = 10
lower_bound_avg = 30
upper_bound_avg = 200
range = upper_bound_avg - lower_bound_avg

z = .95 +(1-.95)/2
n = ((z - (range/4))/confidence_interval_length)^2

print(n)
```

4.  According to a union agreement, the mean income for all senior-level
    workers in a large service company equals \$500 per week. A
    representative of a women's group decides to analyze whether the
    mean income μ for female employees matches this norm. For a random
    sample of nine female employees, ȳ = \$410 and s = 90

```{r}
income_n <- c(rnorm(9, mean = 410, sd = 90))
```

A. Test whether the mean income of female employees differs from \$500
per week. Include assumptions, hypotheses, test statistic, and P-value.
Interpret the result.

assumption: The mean income of female employees is \$500 per week. 
null hypothesis: The mean income of female employees is \$500 per week. alternative hypothesis: The mean income of female employees differs from \$500 per week.

```{r}
income_female <- t.test(income_n, mu=500)
print(income_female)
```

The result is that with 95% certainty the mean female income falls
between \$322.6557 and \$441.1327. Since the range falls below \$500, we reject the null hypothesis that the mean female income is \$500, and accept the alternative hypothesis that the true mean female income is not equal to \$500.

B. Report the P-value for Ha: μ \< 500. Interpret.

```{r}
income_female <- t.test(income_n, mu=500, alternative = 'less')
print(income_female)
```

The p_value is 0.0081. It is less than the 0.5 alpha so we reject the null hypothesis that that the mean female income is \$500, and accept the alternative hypothesis that the true mean female income is less than $500.

C. Report and interpret the P-value for Ha: μ \> 500.

```{r}
income_female <- t.test(income_n, mu=500, alternative = 'greater')
print(income_female)
```

The p_value is 0.9919. It is not less than the 0.5 alpha so we fail to reject the null hypothesis that that the mean female income is $500.

5.  Jones and Smith separately conduct studies to test H0: μ = 500 against Ha: μ ≠ 500, each with n = 1000. Jones gets ȳ = 519.5, with se = 10.0. Smith gets ȳ = 519.7, with se = 10.0.

```{r}
sd.js = 10*(sqrt(1000))

income_jones <- c(rnorm(1000, mean = 519.5, sd = sd.js))
incomes_smith <- c(rnorm(1000, mean = 519.7, sd = sd.js))
```

A. Show that t = 1.95 and P-value = 0.051 for Jones. Show that t = 1.97 and P-value = 0.049 for Smith.

```{r}
jones_t <- (519.5 - 500)/(sd.js/sqrt(1000))
jones_p <- 2*(pt(-jones_t, 999))

print(jones_t)
print(jones_p)


smith_t <- (519.7 - 500)/(sd.js/sqrt(1000))
smith_p <- 2*(pt(-smith_t, 999))

print(smith_t)
print(smith_p)
```

B. Using α = 0.05, for each study indicate whether the result is "statistically significant."

The Jones study result is not statistically significant because their p-value is above the alpha. The Smith study result is statistically significant because the p-value is below the alpha.

C. Using this example, explain the misleading aspects of reporting the result of a test as "P ≤ 0.05" versus "P \> 0.05," or as "reject H0" versus "Do not reject H0," without reporting the actual P-value.

Because in this case the p-value is so close to the alpha, it's misleading to not report the actual p-value. If someone only read that a null hypothesis was not rejected and not that statistical significance criteria was almost met, they may falsely attribute the strength of confidence of which the null hypothesis was not rejected with. It's important to report the p-value so that the strength of confidence is
known.

6.  A school nurse wants to determine whether age is a factor in whether children choose a healthy snack after school. She conducts a survey of 300 middle school students, with the results below. Test at α = 0.05 the claim that the proportion who choose a healthy snack differs by grade level. What is the null hypothesis? Which test should we use? What is the conclusion?

x = proportion of students who choose a healthy snack
null hypothesis: X is not affected by grade level. 
alternative hypothesis: X is affected by grade level. 
```{r}
grade <- c(6, 7, 8) 
healthy_prop <- c(31, 43, 51)
unhealthy_prop <- c(69, 57, 49)
student_n <- c(100, 100, 100)

student_data <- data.frame(grade, healthy_prop, unhealthy_prop)

student_t <- prop.test(student_data$healthy_prop, n=student_n)
print(student_t)
```
The conclusion is that since the p-value of 0.01547 is below our 0.05 value we reject the null hypothesis that the proportion of students who choose a healthy snack is not affected by grade level in favor of the alternative hypothesis that the proportion of students who choose a healthy snack is affected by grade level. 

7. Per-pupil costs (in thousands of dollars) for cyber charter school tuition for school districts in three areas are shown. Test the claim that there is a difference in means for the three areas, using an appropriate test. What is the null hypothesis? Which test should we use? What is the conclusion?

null hypothesis: There is not a difference in mean per-pupil cost (in thousands of dollars) for cyber charter school tuition for school districts for the three areas.
alternative hypothesis: There is a difference in mean per-pupil cost (in thousands of dollars) for cyber charter school tuition for school districts for the three areas. 
```{r}
values <- c(6.2,9.3,6.8,6.1,6.7,7.5,
           7.5,8.2,8.5,8.2,7.0,9.3,
           5.8,6.4,5.6,7.1,3.0,3.5)
area <- c(1,1,1,1,1,1,
          2,2,2,2,2,2,
          3,3,3,3,3,3)

cost_data <- data.frame(values, area)

anova <- aov(values~area, data = cost_data)
summary(anova)
```
The conclusion is that since the p-value of 0.0542 is above our 0.05 value we fail reject the null hypothesis that there is not a difference in mean per-pupil cost (in thousands of dollars) for cyber charter school tuition for school districts for the three areas in favor of the alternative hypothesis that there is a difference in mean per-pupil cost (in thousands of dollars) for cyber charter school tuition for school districts for the three areas. 