---
title: "Homework2"
author: "Rahul Somu"
description: "Homework2"
date: "04/02/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - desriptive statistics 
  - probability
---

# Question 1

Below code chunk calculated the confidence interval for bypass surgery as [18.80009, 19.19991], and the confidence interval for angiography as [17.76757, 18.23243].
Based on the results, we can conclude confidence interval for bypass surgery is slightly narrower than that of  angiography, which implies that the estimate of the mean wait time for bypass surgery is slightly more precise than that of angiography.

```{r, echo=T}
# Create dataframe
df <- data.frame(
  procedure = c("Bypass", "Angiography"),
  sample_size = c(539, 847),
  sample_mean = c(19, 18),
  sample_sd = c(10, 9)
)

# confidence level
conf_level <- 0.9

#degrees of freedom for each procedure
df$df <- df$sample_size - 1

#critical value for the confidence interval
t_critical <- qt(1 - (1 - conf_level) / 2, df$df)

#standard error of the mean for each procedure
df$sem <- df$sample_sd / sqrt(df$sample_size)

# confidence intervals for each procedure
df$ci <- apply(df[, c("sample_mean", "sem", "df")], 1, function(x) {
  x[1] + c(-1, 1) * t_critical * x[2] * sqrt(x[3] + 1) / sqrt(x[3])
})

# Print the confidence intervals
cat("Confidence intervals:\n")
print(df$ci)

```
#Question 2

Below results suggest that we are 95% confident that the true proportion of adult Americans who believe that the college education is essential for success lies between 0.5189 and 0.5808.

```{r, echo=T}

n <- 1031
p_hat <- 567/1031
z <- qnorm(1-0.05/2)

CI <- p_hat + z*sqrt(p_hat*(1-p_hat)/n) * c(-1, 1)
CI

```

#Question 3

```{r, echo=T}

n = ((1.959964)^2 * (42.5)^2) / (5)^2
n
```

#Question 4

```{r, echo=T}

# Part A
n <- 9
ybar <- 410
s <- 90
mu0 <- 500
alpha <- 0.05
se <- s / sqrt(n)
t <- (ybar - mu0) / se
p_value <- 2 * pt(-abs(t), df = n - 1)

# Report results
cat("Test statistic:", round(t, 2), "\n")
cat("P-value:", p_value, "\n")
if(p_value < alpha) {
  cat("Reject null hypothesis; the mean income of female employees differs from $500 per week.\n")
} else {
  cat("Fail to reject null hypothesis. \n")
}

# Part B
p_value_lt <- pt(t, df = n - 1)
cat("P-value (Ha: μ < 500):", p_value_lt, "\n")
if(p_value_lt < alpha) {
  cat("the mean income of female employees is less than $500 per week.\n")
} else {
  cat("Fail to reject null hypothesis. \n")
}

# Part C
p_value_gt <- pt(-t, df = n - 1)
cat("P-value (Ha: μ > 500):", p_value_gt, "\n")
if(p_value_gt < alpha) {
  cat("the mean income of female employees is greater than $500 per week.\n")
} else {
  cat("Fail to reject null hypothesis\n")
}

```

#QUESTION 5

```{r, echo=T}
# Jones' study
jones <- data.frame(y_bar = 519.5, se = 10.0, n = 1000)
jones$t <- (jones$y_bar - 500) / jones$se
jones$p_value <- 2 * pt(-abs(jones$t), df = jones$n - 1)
jones$t
jones$p_value

# Smith's study
smith <- data.frame(y_bar = 519.7, se = 10.0, n = 1000)
smith$t <- (smith$y_bar - 500) / smith$se
smith$p_value <- 2 * pt(-abs(smith$t), df = smith$n - 1)
smith$t
smith$p_value 
# Significance testing
alpha <- 0.05
if (jones$p_value < alpha) {
  cat("Jones' study is statistically significant\n")
} else {
  cat("Jones' study is not statistically significant\n")
}
if (smith$p_value < alpha) {
  cat("Smith's study is statistically significant\n")
} else {
  cat("Smith's study is not statistically significant\n")
}




```
#Question 6

```{r, echo=T}
# Create contingency table
snack_table <- matrix(c(31, 43, 51, 69, 57, 49), nrow = 3, byrow = TRUE)

# Perform chi-squared test
chisq.test(snack_table)

```
#Question 7

```{r, echo=T}
# Create a data frame
area <- c(rep("Area 1", 6), rep("Area 2", 6), rep("Area 3", 6))
cost <- c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5, 7.5, 8.2, 8.5, 8.2, 7.0, 9.3, 5.8, 6.4, 5.6, 7.1, 3.0, 3.5)
cost_data <- data.frame(area, cost)

# Perform one-way ANOVA
model <- aov(cost ~ area, data = cost_data)

# Print ANOVA table summary
summary(model)

```