---
title: "Homework - 3"
author: "Hannah Rosenbaum"
description: "Homework 3"
date: "05/04/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw3
  - desriptive statistics 
  - probability
---

# Question 1

```{r}
library(alr4)
library(smss)
```

```{r}
data(UN11)
```

### a

```{r}
head(UN11)
summary(UN11)
pairs(UN11)
```

Predictor: ppgdp, Response: fertility

## b

```{r}
plot(UN11$ppgdp, UN11$fertility)
```

No, a linear model will not fit this curve.

### c

```{r}
plot(log(UN11$ppgdp), log(UN11$fertility))
```

Yes, a simple linear regression will fit this dataset much better.

# Question 2

### a

The slope will not change, but the y-intercept will increase.

### b

The residuals will change.

# Question 3

```{r}
data(water)
```

```{r}
summary(water)
pairs(water)
```

Pairs of higher correlation: APMAM & APSAB, APSAB & APSLAKE, OPBPC & OPRC, OPBPC & OPSLAKE, OPBPC & BSAAM, OPRC & OPSLAKE, OPRC & BSAAM, OPSLAKE & BSAAM

# Question 4

```{r}
data(Rateprof)
```

```{r}
pairs(Rateprof[, c('quality', 'helpfulness', 'clarity', 'easiness', 'raterInterest')])
```

Quality, Helpfulness, and Clarity have a strong correlation. Easiness compared to the prior three has a weak relation. raterInterest has little correlation to the other four.

# Question 5

```{r}
data("student.survey")
summary(student.survey)
```

```{r}
fita <- lm(hi ~ tv, data=student.survey)
#fitb <- lm(pi ~ re, data=student.survey)
```

```{r}
summary(fita)
AIC(fita)
BIC(fita)
```
