---
title: "Homework 3"
author: "Miguel Curiel"
description: "Regression analyses and transformations for DACSS 603."
date: "04/11/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw3
  - linear regression
  - distribution transformation
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, eval=TRUE}
# load necessary packages
library(tidyverse)
library(alr4)
library(smss)
```

# Question 1

**United Nations** (Data file: UN11in alr4) The data in the file UN11
contains several variables, including ppgdp, the gross national product
per person in U.S. dollars, and fertility, the birth rate per 1000
females, both from the year 2009. The data are for 199 localities,
mostly UN member countries, but also other areas such as Hong Kong that
are not independent countries. The data were collected from the United
Nations (2011). We will study the dependence of fertility on ppgdp.

a.  Identify the predictor and the response.

    a.  Predictor = ppgdp

    b.  Response = fertility

b.  Draw the scatterplot of fertility on the vertical axis versus ppgdp
    on the horizontal axis and summarize the information in this graph.
    Does a straight-line mean function seem to be plausible for a
    summary of this graph?

    a.  No, a straight-line mean function does not seem plausible
        because the distribution seems to be somewhat curvilinear.

c.  Draw the scatterplot of log(fertility) versus log(ppgdp) using
    natural logarithms. Does the simple linear regression model seem
    plausible for a summary of this graph? If you use a different base
    of logarithms, the shape of the graph won't change, but the values
    on the axes will change.

    a.  Yes, a simple linear regression model with a logarithmic
        transformation seems to be plausible as it is better at
        capturing the curvilinear relationship between variables.

```{r, eval=TRUE, echo=TRUE}
ggplot(data = UN11, aes(x = ppgdp, y = fertility)) +
  geom_point() +
  geom_smooth(method = 'lm', se=F) +
  labs(title='1.b. Scatterplot of fertility versus ppgdp')
```

```{r, eval=TRUE, echo=TRUE}
ggplot(data = UN11, aes(x = log(ppgdp), y = log(fertility))) +
  geom_point() +
  geom_smooth(method = 'lm', se=F) +
  labs(title='1.c. Scatterplot of log(fertility) versus log(ppgdp)')
```

# Question 2

Annual income, in dollars, is an explanatory variable in a regression
analysis. For a British version of the report on the analysis, all
responses are converted to British pounds sterling (1 pound equals about
1.33 dollars, as of 2016).

a.  How, if at all, does the slope of the prediction equation change?

    a.  Using the standard slope-intercept equation - `y = mx + b` - the
        slope (m) will not change as this this relationship is
        independent of the unit of measurement. What will change is the
        intercept (b), meaning that the starting point will be 1.33
        instead of 1.

b.  How, if at all, does the correlation change?

    a.  Similarly to the slope, the correlation does not change as this
        is affected by the relationship between two variables (y = mx)
        rather than the starting point or, in this case, the unit of
        measurement (b).

# Question 3

**Water runoff in the Sierras** (Data file: water in alr4) Can Southern
California's water supply in future years be predicted from past data?
One factor affecting water availability is stream runoff. If runoff
could be predicted, engineers, planners, and policy makers could do
their jobs more efficiently. The data file contains 43 years' worth of
precipitation measurements taken at six sites in the Sierra Nevada
mountains (labeled APMAM, APSAB, APSLAKE, OPBPC, OPRC, and OPSLAKE) and
stream runoff volume at a site near Bishop, California, labeled BSAAM.
Draw the scatterplot matrix for these data and summarize the information
available from these plots.

```{r, eval=TRUE, echo=TRUE}

pairs(water[,-1])
```

As we can see from the scatterplot matrix, the sites that seem to
correlate the most to BSAAM are OPSLAKE, OPRC, and OPBPC, while APSLAKE,
APSAB, and APMAM do not seem to be correlated at all. Therefore, if a
model were to be created to predict runoff, a good idea would be to
start by including the precipitation of sites that begin with an "O" and
exclude those that begin "A".

# Question 4

**Professor ratings** (Data file: Rateprof in alr4) In the website and
online forum RateMyProfessors.com, students rate and comment on their
instructors. Launched in 1999, the site includes millions of ratings on
thousands of instructors. The data file includes the summaries of the
ratings of 364 instructors at a large campus in the Midwest
(Bleske-Rechek and Fritsch, 2011). Each instructor included in the data
had at least 10 ratings over a several year period. Students provided
ratings of 1--5 on quality, helpfulness, clarity, easiness of
instructor's courses, and raterInterest in the subject matter covered in
the instructor's courses. The data file provides the averages of these
five ratings. Create a scatterplot matrix of these five variables.
Provide a brief description of the relationships between the five
ratings.

```{r, eval=TRUE, echo=TRUE}

pairs(~ quality + helpfulness + clarity + easiness + raterInterest, data=Rateprof)
```

Just by looking at the scatterplot matrix, there are a couple of
variables that are seemingly correlated. In particular, quality seems to
have a nearly perfect, positive correlation with helpfulness and
clarity. Similarly, helpfulness seems to have a moderate-to-high
positive correlation. In contrast, easiness of the instructor's courses
and raterInterest do not seem to be highly correlated with any of the
variables.

# Question 5

For the student.survey data file in the smss package, conduct regression
analyses relating (i) y = political ideology and x = religiosity, and
(ii) y = high school GPA and x = hours of TV watching.

a.  Graphically portray how the explanatory variable relates to the
    outcome variable in each of the two cases.

    1.  Plots can be found below.

b.  Summarize and interpret results of inferential analyses.

    1.  From the numerical analysis, both instances are statistically
        significant, however scenario i (political ideology against
        religiosity) is much more significant than scenario ii (high
        school GPA against hours spent watching TV). Further, from the
        graphics we can see that only scenario i (political ideology
        against religiosity) is clearly linear, whereas scenario ii
        (high school GPA against hours spent watching TV) could benefit
        from preprocessing techniques such as a logarithmic
        transformation or treating outliers/missing data.

```{r, eval=TRUE, echo=TRUE}
data("student.survey", package = "smss")
df <- student.survey

df$pi_numeric <- factor(df$pi, levels = c(
  "very conservative"
  ,"conservative"
  ,"slightly conservative"
  ,"moderate"
  ,"slightly liberal"
  ,"liberal"
  ,"very liberal"
), labels = c(1,2,3,4,5,6,7)
)

df$pi_numeric <- as.numeric(as.character(df$pi_numeric))

df$re_numeric <- factor(df$re, levels = c(
  "never"
  ,"occasionally"
  ,"most weeks"
  ,"every week"
), labels = c(1,2,3,4)
)

df$re_numeric <- as.numeric(as.character(df$re_numeric))

summary(lm(pi_numeric ~ re_numeric, data=df))
```

```{r, eval=TRUE, echo=TRUE}
ggplot(data = df, aes(x = re_numeric, y = pi_numeric)) +
  geom_point() +
  geom_smooth(method = 'lm', se=F) +
  labs(title='5.a.i. Scatterplot of political ideology versus religiosity'
       ,x="religiosity"
       ,y='political ideology') +
  scale_y_continuous(labels = c(
  "very conservative"
  ,"conservative"
  ,"slightly conservative"
  ,"moderate"
  ,"slightly liberal"
  ,"liberal"
  ,"very liberal"
    ), breaks = c(1,2,3,4,5,6,7)) +
  scale_x_continuous(labels = c(
  "never"
  ,"occasionally"
  ,"most weeks"
  ,"every week"
    ), breaks = c(1,2,3,4))
```

```{r, eval=TRUE, echo=TRUE}
ggplot(data = df, aes(x = tv, y = hi)) +
  geom_point() +
  geom_smooth(method = 'lm', se=F) +
  labs(title='5.a.ii. Scatterplot of high school GPA against hours spent watching TV'
       ,x='hours spent watching TV'
       ,y='high school GPA')
```
