---
title: "Homework 3"
author: "Xiaoyan"
description: "Template of course blog qmd file"
date: "04/05/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw3
  - desriptive statistics 
  - probability
---
```{r}
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(alr4)
library(smss)
library(stargazer)
```
# Question 1
1. United Nations (Data file: UN11in alr4) The data in the file UN11 contains several variables,
including ppgdp, the gross national product per person in U.S. dollars, and fertility, the birth
rate per 1000 females, both from the year 2009. The data are for 199 localities, mostly UN
member countries, but also other areas such as Hong Kong that are not independent countries.
The data were collected from the United Nations (2011). We will study the dependence of
fertility on ppgdp.
(a) Identify the predictor and the response.
```{r}
head(UN11)
```
predictor is ppgdp and response is fertility and birth rate. 

(b) Draw the scatterplot of fertility on the vertical axis versus ppgdp on the horizontal axis
and summarize the information in this graph. Does a straight-line mean function seem to
be plausible for a summary of this graph?
```{r}
ggplot(UN11, aes(ppgdp, fertility)) +
  geom_point()
```
straight line doesn't fit for summary this graph.

(c) Draw the scatterplot of log(fertility) versus log(ppgdp) using natural logarithms. Does
the simple linear regression model seem plausible for a summary of this graph? If you use
a different base of logarithms, the shape of the graph won’t change, but the values on the
axes will change.
```{r}
ggplot(UN11, aes(ppgdp, fertility)) +
  geom_point()+scale_y_log10()+scale_x_log10()
```
# Question 2
Annual income, in dollars, is an explanatory variable in a regression analysis. For a British
version of the report on the analysis, all responses are converted to British pounds sterling (1 pound
equals about 1.33 dollars, as of 2016).
(a) How, if at all, does the slope of the prediction equation change?
```{r}
```
(b) How, if at all, does the correlation change?
```{r}
```
# Question 3
Water runoff in the Sierras (Data file: water in alr4) Can Southern California’s water
supply in future years be predicted from past data? One factor affecting water availability is stream
runoff. If runoff could be predicted, engineers, planners, and policy makers could do their jobs

more efficiently. The data file contains 43 years’ worth of precipitation measurements taken at six
sites in the Sierra Nevada mountains (labeled APMAM, APSAB, APSLAKE, OPBPC, OPRC, and
OPSLAKE) and stream runoff volume at a site near Bishop, California, labeled BSAAM. Draw
the scatterplot matrix for these data and summarize the information available from these
plots. (Hint: Use the pairs() function.)
```{r}
```
# Question 4
Professor ratings (Data file: Rateprof in alr4) In the website and online forum
RateMyProfessors.com, students rate and comment on their instructors. Launched in 1999, the site
includes millions of ratings on thousands of instructors. The data file includes the summaries of
the ratings of 364 instructors at a large campus in the Midwest (Bleske-Rechek and Fritsch, 2011).
Each instructor included in the data had at least 10 ratings over a several year period. Students
provided ratings of 1–5 on quality, helpfulness, clarity, easiness of instructor’s courses, and
raterInterest in the subject matter covered in the instructor’s courses. The data file provides the
averages of these five ratings. Create a scatterplot matrix of these five variables. Provide a
brief description of the relationships between the five ratings.
```{r}
```
# Question 5
For the student.survey data file in the smss package, conduct regression analyses relating
(by convention, y denotes the outcome variable, x denotes the explanatory variable)
(i) y = political ideology and x = religiosity,
```{r}
```
(ii) y = high school GPA and x = hours of TV watching.
(You can use ?student.survey in the R console, after loading the package, to see what each variable
means.)
```{r}
```
(a) Graphically portray how the explanatory variable relates to the outcome variable in
each of the two cases
```{r}
```
(b) Summarize and interpret results of inferential analyses.
```{r}
```