---
title: "Homework - 4"
author: "Hannah Rosenbaum"
description: "Homework 4"
date: "05/04/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw4
  - desriptive statistics 
  - probability
---

```{r}
library(alr4)
library(smss)
```

```{r}
data(salary)
data("house.selling.price")
```

# Question 1

### a

ŷ = −10536 + 53.8x1 + 2.84x2

ŷ = -10536 + 53.8(1240) + 2.84(18000)

ŷ = 107296

residual = 145000 - 107296 = 37704

### b

ŷ = −10536 + 53.8x1 + 2.84x2

ŷ = -10536 + 53.8(1240) + 2.84(18000)

ŷ = 107296

ŷ2 = −10536 + 53.8x1 + 2.84x2

ŷ2 = -10536 + 53.8(1241) + 2.84(18000)

ŷ2 = 107349.8

A 1 point increase in square foot home size results in a \$53.8 home value increase.

### c

107349.8 = -10536 + 53.8(1240) + 2.84(x)

2.84(x) = 51173.8

x = 18018.94

# Question 2

### a

```{r}
boxplot(salary ~ sex, data=salary)
```

### b

```{r}
fit <- lm(salary ~ year + ysdeg + degree + rank + sex, data=salary)
summary(fit)
confint(fit)
```

### c

year and ysdeg have the narrowest interval. rankAssoc has the widest interval.

### d

```{r}
fit <- lm(salary ~ rank, data=salary)
summary(fit)
confint(fit)
```

### e

```{r}
fit2 <- lm(salary ~ year + ysdeg + degree + sex, data=salary)
summary(fit2)
confint(fit2)
```

Rsquared decreased by 0.2359, residuals decreased by 130

### f

```{r}
new <- salary[salary$year <= 15,]
pairs(salary)
pairs(new)
```

```{r}
summary(salary)
summary(new)
```

```{r}
fit3 <- lm(salary ~ year + ysdeg + degree + sex, data=new)
summary(fit3)
confint(fit3)
```

# Question 3

### a

```{r}
fit <- lm(Price ~ New + Size, data=house.selling.price)
summary(fit)
```

### b

```{r}
fit2 <- lm(Price ~ New + Size, data=house.selling.price[house.selling.price$New==0,])
summary(fit2)

fit3 <- lm(Price ~ New + Size, data=house.selling.price[house.selling.price$New==1,])
summary(fit3)
```

### c

```{r}
predict(fit, data.frame(New=0, Size = 3000))
predict(fit, data.frame(New=1, Size = 3000))
```

### d

```{r}
fit4 <- lm(Price ~ New + Size + New:Size, data=house.selling.price)
summary(fit4)
```

### e

```{r}
plot(fit2)
plot(fit3)
```

### f

```{r}
predict(fit4, data.frame(New=0, Size = 3000))
predict(fit4, data.frame(New=1, Size = 3000))
```

### g

```{r}
predict(fit4, data.frame(New=0, Size = 1500))
predict(fit4, data.frame(New=1, Size = 1500))
```

### h

The model with interaction appears to estimate lower prices than the original regression.
