---
title: "Homework4"
author: "Rahul Somu"
description: "Homework4"
date: "05/16/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw4
editor_options: 
  chunk_output_type: inline
---

# Question 1


```{r, echo=T}
calculate_prediction <- function(home_size, lot_size) {
  intercept <- -10536
  coefficient_home_size <- 53.8
  coefficient_lot_size <- 2.84
  predicted_price <- intercept + coefficient_home_size * home_size + coefficient_lot_size * lot_size
  
  actual_price <- 145000
  residual <- actual_price - predicted_price
  
  result <- list(
    predicted_price = predicted_price,
    residual = residual
  )
  return(result)
}

result <- calculate_prediction(home_size = 1240, lot_size = 18000)

predicted_price <- result$predicted_price
residual <- result$residual

predicted_price
residual

increase_price <- 53.8

increase_lot_size <- 53.8*1 / 2.84

increase_price
increase_lot_size

```
#Question 2

#A
we can reject the null hypothesis that the mean salary for men and women is the same. There is a significant difference in salary between men and women, with male faculty earning more on average.

#C
Based on the coefficients, we can interpret that degreePhD,sexFemale,ysdeg are Not statistically significant and rankAssoc,rankProf,year are Statistically significant.

#E
The coefficient for sex show that females make $1286 less than males.



```{r, echo=T}
library(alr4)

data(salary)

summary(lm(salary ~ sex, data = salary))

#B
model <- lm(salary ~ ., data = salary)
summary(model)

summary(model)$coefficients

#D
salary$rank <- relevel(salary$rank, ref = "Prof")
model <- lm(salary ~ ., data = salary)
summary(model)$coefficients

model <- lm(salary ~ . - rank, data = salary)
summary(model)

#F
salary$new_dean <- ifelse(salary$ysdeg <= 15, 1, 0)
cor.test(salary$new_dean, salary$ysdeg)
summary(lm(salary ~ . -ysdeg, data = salary))

```

#Question 3


```{r, echo=T}
library(smss)

data(house.selling.price)

model1 <- lm(Price ~ Size + New, data = house.selling.price)

summary(model1)

summary(lm(Price ~ Size * New, data = house.selling.price))
```


#B
selling_price_new = -40230.867 + 116.132 * Size + 57736.283
selling_price_not_new = -40230.867 + 116.132 * Size



#C

selling_price_new =  -40230.867 + 116.132 * 3000 + 57736.283 

selling_price_not_new = -40230.867 + 116.132 * 3000 = 25693.6



```{r, echo=T}
#D
new_interaction_term <- lm(Price ~ Size + New + Size * New, data = house.selling.price)
summary(new_interaction_term)

```


#E

new_sp_pred = -22228 + 104*Size - 78528 + 62*Size = -100756 + 168*Size

not_new_pred = -22228 + 104*Size


```{r, echo=T}
#F
new_pred <- -100756 + 168*(3000)
not_new_pred <- 22228 + 104*(3000)

new_pred
not_new_pred


```


```{r, echo=T}

#G
new_pred <- -100756 + 168*(1500)
not_new_pred <- 22228 + 104*(1500)

new_pred
not_new_pred
```

H. I prefer the model with interaction as it provides a better fit to the data. The R-squared value for the model with interaction is 0.72, which is higher than the R-squared value for the model without interaction (0.67). This indicates that the model with interaction is able to explain more of the variation in the data.
