---
title: "Homework 4"
author: "Guanhua Tan"
description: "Homework 4"
date: "04/01/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw4
  - regression analysis
---

```{r}
library(tidyverse)
library(smss)
library(alr4)
library(ggplot2)
data(UN11)
```

# Question 1

```{r}
y<--10536 + 53.8 *1240 +2.84 *18000
y
redidual_1 <- 145000-y
redidual_1

x2 <-53.8/2.84
x2
```

A the predicted selling price is 107,292; the residual is 37,704; interpret is -10536.

B For fixed lot size, if the house size increase one square feet, the house price will increase 53.8.

C if the lot size need to increase 18.94, that will have the same impact as a one-square-foot increase in home size.


# Question 2
```{r}
data(salary)
head(salary)
summary(lm(salary~sex, salary))
```

A. According the Chisq test, sex has the impact on salary.

```{r}
lm(salary ~ ., data = salary) |>
  confint()
```
The 95% confidence intervals for the female salary is (-697.81, 3030.56), other variables controlled.

```{r}

sex_salary <- lm(salary~degree+rank+sex+year+ysdeg, data=salary)
summary(sex_salary)
```
C. degreePhD, sexFemale, and ysdeg doesn't show statistical significance. For other fixed terms, if one person hold a associate professor ranking, the salary increases 5292.36, compared to a assistant professor. In the same condition, if one person holds a full professor, the salary increases 11118.76, compared to a assistant professor. People's age has the positive impact on their salary. When the age increases 1 year, the salary increases 476.31.

```{r}
salary$rank <-relevel(salary$rank, ref="Prof")
salary$sex <-relevel(salary$sex, ref="Female")

sex_salary <- lm(salary~degree+rank+sex+year+ysdeg, data=salary)
summary(sex_salary)
```

D. For other fixed terms, if one person holds a assistant professor ranking, the salary decreases 11118.76, compared to a full professor. In the same condition, if one person hold a associate professor, the salary decreases 5826.40, compared to a full professor.

```{r}
sex_salary <- lm(salary~degree+sex+year+ysdeg, data=salary)
summary(sex_salary)
```
E if the model drops the variable rank, the variable years after the degree presents the strongly statistical significance.

```{r}
salary_F <- salary %>%
  mutate(fifteen_years= case_when(ysdeg > 15 ~0,ysdeg <= 15 ~ 1))
summary(lm(salary~degree + sex + rank+ year+fifteen_years,salary_F))

```

# Question 3

```{r}
data("house.selling.price")
head(house.selling.price)
house_model_1 <- lm(Price~Size+New, data=house.selling.price)
summary(house_model_1)
```
A. The variables Size and New show the statistical significance. the price will increase 116 when the house size increases 1 square feet. If the house is new, the price will increase 57736.283, compared with the size old house.

B. Price= 116.132*Size+57736.283*New-40230.86
For new homes
Price=115.132*Size + 57736.283-40230.86
For not new homes
Price=115.132*Size -40239.86.

```{r}
y_new <- 116.32*3000+57736.283-40230.86
y_old <- 116.32*3000-40230.86
y_new
y_old
```

C. the new home price is 366465.4; the old home price is 308729.1

```{r}
house_model_2 <- lm(Price~Size+New*Size, data=house.selling.price)
summary(house_model_2)

```
D. The regression result shows that the interaction term is statistical significance and r squared value is larger than the previous model.

E. for new homes, y=-22227.808+104.438*Size-78527.502+61.916*Size
for old homes, y=104.438*Size-100755.3

```{r}
Size_3000 <-3000
Price_new <- 22227.808+104.438*Size_3000-78527.502+61.916*Size_3000
Price_old <-104.438*Size_3000-100755.3
Price_new
Price_old
```
F. For 3000 square feet, the new hous is 442,762.3 which the old one is 212,558.7

```{r}
Size_1500 <-1500
Price_new_1500 <- 22227.808+104.438*Size_1500-78527.502+61.916*Size_1500
Price_old_1500 <-104.438*Size_1500-100755.3
Price_new_1500
Price_old_1500
```

G. According to the predicted lines, when the size increases, the new house price will increase more than the old one's.

H. I think the model with the interaction one represents the relationship between size and new because it has a larger value of R square.








