---
title: "HW4"
author: "Liam Tucksmith"
desription: "HW4"
date: "05/12/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw4
  - Liam Tucksmith
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(smss)
library(alr4)
```

1.  For recent data in Jacksonville, Florida, on y = selling price of
    home (in dollars), x1 = size of home(in square feet), and x2 = lot
    size (in square feet), the prediction equation is y = ???10,536 +
    53.8x1 + 2.84x2.

A. A particular home of 1240 square feet on a lot of 18,000 square feet
sold for \$145,000. Find the predicted selling price and the residual,
and interpret.

```{r}
x1 <- 1240
x2 <- 18000
y <- 145000

predicted <- (-10536 + 53.8*(x1) + 2.84*(x2))
residual <- y - predicted
print(predicted)
print(residual)
```

The predicted value is lower than the actual value, indicating that the
equation is undervaluing the homes.

B. For fixed lot size, how much is the house selling price predicted to
increase for each square- foot increase in home size? Why?

```{r}
x1 <- 1240
x2 <- 18000
y <- 145000

fixed_price <- (-10536 + 53.8*(x1) + 2.84*(x2))

x1 <- 1241
fixed_price_sqft_inc <- (-10536 + 53.8*(x1) + 2.84*(x2))

print(fixed_price_sqft_inc - fixed_price)
```

Using the numbers from part A, we can see that a 1 sqft increase results
in a \$53.80 predicted sell price increase when the lot size is fixed.

C. According to this prediction equation, for fixed home size, how much
would lot size need to increase to have the same impact as a
one-square-foot increase in home size?

```{r}
home_size_inc <- fixed_price_sqft_inc - fixed_price

x1 <- 1240
x2 <- 18000
y <- 145000

fixed_price_c <- (-10536 + 53.8*(x1) + 2.84*(x2))

lot_inc_pred <- fixed_price_c + home_size_inc
part1 <- (-10536 + 53.8*(x1))
```

part1 + 2.84*((x2)+z) = lot_inc_pred (lot_inc_pred - part1)=
(((x2)+z)*2.84)/2.84 (lot_inc_pred - part1)/2.84 = (x2)+z

```{r}
z <- (lot_inc_pred - part1)/2.84 - x2
print(z)
```

To have an impact of a 1 sqft increase in home size, lot size would have
to increase by 18.9 sqft.

2 (Data file: salary in alr4 R package). The data file concerns salary
and other characteristics of all faculty in a small Midwestern college
collected in the early 1980s for presentation in legal proceedings for
which discrimination against women in salary was at issue. All persons
in the data hold tenured or tenure track positions; temporary faculty
are not included. The variables include degree, a factor with levels PhD
and MS; rank, a factor with levels Asst, Assoc, and Prof; sex, a factor
with levels Male and Female; Year, years in current rank; ysdeg, years
since highest degree, and salary, academic year salary in dollars.

A. Test the hypothesis that the mean salary for men and women is the
same, without regard to any other variable but sex. Explain your
findings.

```{r}
summary(lm(salary ~ sex, data = salary))
```

On average, the women make \$3340 less than the men. The p-value of 0.07
indicates that this relationship is not statistically significant at the
.05 level.

B. Run a multiple linear regression with salary as the outcome variable
and everything else as predictors, including sex. Assuming no
interactions between sex and the other predictors, obtain a 95%
confidence interval for the difference in salary between males and
females.

```{r}
summary(lm(salary ~ ., data = salary))
confint(lm(salary ~ ., data = salary))
```

C. Interpret your finding for each predictor variable; discuss (a)
statistical significance, (b) interpretation of the coefficient / slope
in relation to the outcome variable and other variables

degreePhD - Not statistically significant. The coefficient of \$1388
suggests that the faculty with phDs have, on avergage, a salary \$1388
higher than faculty with masters.

rankAssoc - Statistically significant. The coefficient of \$5292
suggests that the faculty with associate rank have, on avergage, a
salary \$5292 higher than faculty with assistant rank.

rankProf - Statistically significant. The coefficient of \$11118
suggests that the faculty with phDs have, on avergage, a salary \$11118
higher than faculty with assistant rank.

sexFemale - Not statistically significant. The coefficient of \$1166
suggests that the female faculty have, on avergage, a salary \$1166
higher than male faculty.

year - Statistically significant. The coefficient of \$476 suggests that
the faculty gain \$476 in salary each additional year worked.

ysdeg - Not statistically significant. The coefficient of -\$124
suggests that the faculty with phDs have, on avergage, a salary -\$124
lower with each passing year since high school graduation.

D. Change the baseline category for the rank variable. Interpret the
coefficients related to rank again.

```{r}
salary$rank <- relevel(salary$rank, ref = 'Prof')
summary(lm(salary ~ ., data = salary))
confint(lm(salary ~ ., data = salary))
```

Nothing changed, the model displays the same information, just with
different variables. Here, the coefficients tell us that assistant
faculty members make \$11118 less than professors, and associate faculty
members make \$7866 less than professors.

E. Finkelstein (1980), in a discussion of the use of regression in
discrimination cases, wrote, "\[a\] variable may reflect a position or
status bestowed by the employer, in which case if there is
discrimination in the award of the position or status, the variable may
be 'tainted.' " Thus, for example, if discrimination is at work in
promotion of faculty to higher ranks, using rank to adjust salaries
before comparing the sexes may not be acceptable to the courts. Exclude
the variable rank, refit, and summarize how your findings changed, if
they did.

```{r}
summary(lm(salary ~ . -rank, data = salary))
```

The sex coefficient flipped to show that females make -\$1286 less than
males, but it's still not statistically signigicant.

F. Everyone in this dataset was hired the year they earned their highest
degree. It is also known that a new Dean was appointed 15 years ago, and
everyone in the dataset who earned their highest degree 15 years ago or
less than that has been hired by the new Dean. Some people have argued
that the new Dean has been making offers that are a lot more generous to
newly hired faculty than the previous one and that this might explain
some of the variation in Salary.

Create a new variable that would allow you to test this hypothesis and
run another multiple regression model to test this. Select variables
carefully to make sure there is no multicollinearity. Explain why
multicollinearity would be a concern in this case and how you avoided
it. Do you find support for the hypothesis that the people hired by the
new Dean are making higher than those that were not?

```{r}
salary$yrs_since_dean <- ifelse(salary$ysdeg <= 15, 1, 0)
cor.test(salary$yrs_since_dean, salary$ysdeg)

summary(lm(salary ~ . -ysdeg, data = salary))
```

3.  Looking at the yrs_since_dean variable, the p-value is statistically
    significant and the positive coefficient, both supporting the
    hypothesis that the dean has been offering higher salaries to newer
    faculty.

(Data file: house.selling.price in smss R package)

A. Using the house.selling.price data, run and report regression results
modeling y = selling price (in dollars) in terms of size of home (in
square feet) and whether the home is new (1 = yes; 0 = no). In
particular, for each variable; discuss statistical significance and
interpret the meaning of the coefficient.

```{r}
data("house.selling.price")
summary(lm(Price ~ Size + New, data = house.selling.price))
```

Both Size and New are statistically significant in this model, and both
have positive coefficients. For New, this indicates that new houses are,
on average, \$57736 more expensive than non-new. And for Size, a 1 sqft
increase in size indicates a \$116 increase in sell price.

B. Report and interpret the prediction equation, and form separate
equations relating selling price to size for new and for not new homes.

selling price prediction = -40230 + 116(size) + 57736(new)

The prediction equation indicates that while controlling for size and
whether or not the size of the car is new, the prediction on average is
\$40230 lower than the actual selling price.

New: selling price prediction = -40230 + 116(size) + 57736

Not new: selling price prediction = -40230 + 116(size)

C. Find the predicted selling price for a home of 3000 square feet that
is (i) new, (ii) not new.

```{r}
new_prediction <- -40230 + 116*(3000) + 57736
not_new_prediction <- -40230 + 116*(3000)

print(new_prediction)
print(not_new_prediction)
```

D. Fit another model, this time with an interaction term allowing
interaction between size and new, and report the regression results

```{r}
summary(lm(Price ~ Size + New + (Size * New), data = house.selling.price))
```

E. Report the lines relating the predicted selling price to the size for
homes that are (i) new, (ii) not new.

New: selling price prediction = -22228 + 104(size) - 78528 + 62(size) =
-100756 + 168(size)

Not new: selling price prediction = -22228 + 104(size)

F. Find the predicted selling price for a home of 3000 square feet that
is (i) new, (ii) not new.

```{r}
new_prediction <- -100756 + 168*(3000)
not_new_prediction <- 22228 + 104*(3000)

print(new_prediction)
print(not_new_prediction)
```

G. Find the predicted selling price for a home of 1500 square feet that
is (i) new, (ii) not new.

```{r}
new_prediction <- -100756 + 168*(1500)
not_new_prediction <- 22228 + 104*(1500)

print(new_prediction)
print(not_new_prediction)
```

Comparing to (F), explain how the difference in predicted selling prices
changes as the size of home increases.

As the size of the home increase, the not new homes increase in price at
a lower rate than the new homes. This is shown by the not new homes
being predicted as more selling for more on average than the new homes
for the 1500 sqft home, and the opposite being true for the 3000 sqft
home predictions.

H. Do you think the model with interaction or the one without it
represents the relationship of size and new to the outcome price? What
makes you prefer one model over another?

I prefer the model without the interaction term, because Size and New
are both statistically significant and therefore all the variables in
the model are statistically significant, which isn't the case for the
model with the interaction term.