---
title: "Homework - 5"
author: "Hannah Rosenbaum"
description: "Homework 5"
date: "05/04/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw5
  - desriptive statistics 
  - probability
---

```{r}
library(alr4)
library(smss)
```

# Question 1

```{r}
data("house.selling.price.2")
```

### a

Using backward elimination, Beds would be removed first.

### b

Using forward elimination, New would be selected first.

### c

Because Beds is correlated to Size, which is correlated to Price, so inherently, more Beds has an impact on Price but still have a low p-value.

### d

```{r}
fit <- lm(Price ~ . -Size -Beds -Baths -New, data=house.selling.price)
summary(fit)
fit <- lm(Price ~ . -Size -Baths -New, data=house.selling.price)
summary(fit)
fit <- lm(Price ~ . -Size -New, data=house.selling.price)
summary(fit)
```

```{r}
fit2 <- lm(Price ~ . -New, data=house.selling.price)
summary(fit)
fit2 <- lm(Price ~ . -New -Size, data=house.selling.price)
summary(fit)
fit2 <- lm(Price ~ . -New -Size -Beds, data=house.selling.price)
summary(fit)
```

```{r}
summary(fit)
AIC(fit)
BIC(fit)

summary(fit2)
AIC(fit2)
BIC(fit2)
```

### e

I prefer the model generated by backward elimination, since the AIC, BIC, and multiple R-squared statistics are higher.

# Question 2

```{r}
data("trees")
```

### a

```{r}
fit <- lm(Volume ~ Girth + Height, data=trees)
summary(fit)
```

### b

```{r}
plot(fit)
```

Residuals vs Fitted violation: curvature

Scale-Location violation: positive trend

# Question 3

```{r}
data(florida)
```

```{r}
fit <- lm(Gore ~ Bush, data=florida)
summary(fit)
```

```{r}
plot(fit)
```

Palm Beach County is an outlier based on falling out of the group in the residuals vs. fitted and Q-Q plots. The Residuals vs. Leverage is just on the line, but it still supports Palm Beach being an outlier.

### b

```{r}
fit <- lm(log(Gore) ~ log(Bush), data=florida)
summary(fit)
```

```{r}
plot(fit)
```

Palm Beach County does not appear to be an outlier.
