---
title: "Final Project 603"
author: "Jerin Jacob"
description: "Final project qmd file for 603"
date: "03/10/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart1
---

Now a days, movies are a well marketed entertainment product. Just like any other products in the market, movies are also having an allocated a marketing budget and promotional activities are done in scale. This often result in the opening weekend's gross ticketing volume to rise. But are the pre-release promotional activities helping the movie to collect more or is it just creating a hype initially? Or does the movie's gross collection is not at all dependant on pre release promotions? This dataset has 200 highest grossing movies of 2022. It has both the opening week's gross as well as the total gross collection of the movies, along with other variables. Assuming that opening week's collection is depending on the pre-release promotion, by looking on the relationship between opening week's gross and total gross, I am trying to see how the pre-release activities help the producers earn more in boxoffice.

**Research Question**: Is the total collection of a movie depending on the pre-release promotional activities?

**Null hypothesis**: The total gross collection of a movie is not dependant on the opening week's collection

**Alternative hypotheses**: Total gross collection is positively dependant on the opening week's collection

## Reading the data

```{r, echo=T}
library(readxl)
library(tidyverse)
library(lubridate)
library(dplyr)
```

```{r}
df <- read.csv("_data/project_data.csv") 
df$Open <- ymd(df$Open)
df$Close <- ymd(df$Close)
df<- df|>
  mutate(open_day = yday(Open), close_day = yday(Close), days_of_show = case_when(close_day < open_day ~ (365-open_day)+close_day, close_day >= 0 ~ close_day-open_day)) 
df

```

There are 10 variables with 200 rows.

### **COLUMN DESCRIPTION**

-   'Rank': rank of the movie

-   'Release': release date of the movie

-   'Gross': domestic gross of the movie

-   'max_th': maximum number of theaters the movie was released in

-   'Opening': gross on opening weekend

-   'perc_tot_gr': domestic percentage of the total gross

-   'open_th': number of theaters the movie opened in

-   'Open': opening date

-   'Close': closing date

-   'Distributor': name of the distributor

-   'int_gross': international gross

-   'world_gross': worldwide gross

## Summary of each variables

```{r, echo=T}
summary(df)
```
```{r}
unique(df$Distributor)
movies_per_distributor <- df |>
  group_by(Distributor) |>
  count(Distributor)
movies_per_distributor  
ggplot(movies_per_distributor, aes(x=Distributor, y=n)) +
  geom_bar(position = 'dodge', stat = 'identity')

```
```{r}


```


Using the glimpse() function, let's have a look at how our data would look like!

```{r}
glimpse(df)
```
Code sample
```{r}
# Load necessary libraries
library(dplyr) # for data manipulation
library(ggplot2) # for data visualization

# Load data into a data frame
movie_data <- read.csv("_data/project_data.csv") 

# Rename columns for clarity
colnames(movie_data) <- c("rank", "release_date", "domestic_gross", "max_theaters", "opening_weekend_gross", "domestic_percentage", "opening_theaters", "opening_date", "closing_date", "distributor", "international_gross", "worldwide_gross")

# Subset data to include only relevant columns
movie_data <- select(movie_data, opening_weekend_gross, worldwide_gross)

# Check data for any missing values or outliers
summary(movie_data)

# Calculate descriptive statistics
initial_week_mean <- mean(movie_data$opening_weekend_gross)
initial_week_sd <- sd(movie_data$opening_weekend_gross)
total_mean <- mean(movie_data$worldwide_gross)
total_sd <- sd(movie_data$worldwide_gross)

# Create scatterplot to visualize the relationship between the variables
ggplot(data = movie_data, aes(x = opening_weekend_gross, y = worldwide_gross)) +
  geom_point() +
  labs(x = "Opening Weekend Gross", y = "Worldwide Gross")

# Calculate correlation coefficient between the two variables
correlation <- cor(movie_data$opening_weekend_gross, movie_data$worldwide_gross)

# Create linear regression model
lm_model <- lm(worldwide_gross ~ opening_weekend_gross, data = movie_data)

# View summary of the linear regression model
summary(lm_model)

# Perform hypothesis test
# Null hypothesis: The total gross collection of a movie is not dependent on the opening week's collection
# Alternative hypothesis: The total gross collection of a movie is dependent on the opening week's collection
#test <- t.test(movie_data$worldwide_gross ~ movie_data$opening_weekend_gross, var.equal = TRUE)

# View results of hypothesis test
#test


```



## References:

1.  Nasir, Suphan & Öcal, Figen. (2016). Film Marketing: The Impact of Publicity Activities on Demand Generation. 10.4018/978-1-5225-0143-5.ch019.

2.  *Elizabeth Cooper-Martin (1991) ,"Consumers and Movies: Some Findings on Experiential Products", in NA - Advances in Consumer Research Volume 18, eds. Rebecca H. Holman and Michael R. Solomon, Provo, UT : Association for Consumer Research, Pages: 372-378.*
