---
title: "Homework 1"
author: "Jerin Jacob"
desription: "Homework 1- 603 Spring 2023"
date: "03/20/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - challenge1
  - Jerin Jacob
  - Lung Capacity Data
  - ggplot2
  - tidyverse
  - 
---

```{r}
#| label: setup
#| warning: false

library(readxl)
library(dplyr)
library(magrittr)
knitr::opts_chunk$set(echo = TRUE)
```
# Question 1

## Loading the Data

```{r}
df <- read_excel("_data/LungCapData.xls")
df
```

## A
```{r}
summary(df)
hist(df$LungCap, xlab = "Lung Capacity", main = "", freq = F)

```
The histogram shows that the distribution is almost a normal distribution with most of the values close to the mean. 
## B
```{r}
boxplot(LungCap ~ Gender, df)

```
The minimum, first quartile, median, third quartile and maximum, all of them appear to be slightly higher for males than females.

## C
```{r}
dff<- df |>
  group_by(Smoke) |>
  summarise(LungCap = mean(LungCap))
```

Lung capacity of non-smokers is higher than that of smokers which is against the expectation

## D
### Age less than or equal to 13
```{r}
df1 <- df |>
  filter(Age <= 13) |>
  group_by(Smoke)|>
  summarise(LungCap = mean(LungCap)) |>
  mutate(Age = "<=13")
df1
```

### Age 14 & 15
```{r}
df2 <- df |>
  filter(Age >= 14 & Age <= 15) |>
  group_by(Smoke)|>
  summarise(LungCap = mean(LungCap))|>
  mutate(Age = "14&15")
class(df2)
```

### Age 16 to 17

```{r}
df3 <- df |>
  filter(Age >= 16 & Age <= 17) |>
  group_by(Smoke)|>
  summarise(LungCap = mean(LungCap))|>
  mutate(Age = "16&17")
```

### Age greater than or equal to 18

```{r}
df4 <- df |>
  filter(Age >= 18) |>
  group_by(Smoke)|>
  summarise(LungCap = mean(LungCap))|>
  mutate(Age = ">=18")
```

## E
```{r}
new_df <- rbind(df1,df2, df3,df4)
new_df
ggplot(new_df, aes(fill=Smoke, y=LungCap, x=Age)) +
  geom_bar(position='dodge', stat='identity')

ggplot(dff, aes(y=LungCap, x=Smoke, fill = Smoke)) +
  geom_bar(position='dodge', stat='identity')  

```
Only age group that showed similar result that of step C is <=13. The analysis could go wrong if the lung capacity of smokers and non smokers are studied without considering the age.

# Question 2

```{r}
df_prison <- tibble(X = c(0,1,2,3,4), Frequency = c(128,434,160,64,24))

df_prison

n <- sum(df_prison$Frequency)
n

```

```{r}
df_prison |>
  filter(X == 2) |>
  pull(Frequency) |>
  divide_by(n)

```

```{r}
#log(write) = 3.135 + 0.115 female + 0.0066 read + 0.0077 math



#write = - 99.164 + 5.389 female + 20.941 log(math) + 16.852 log(read)



```


