---
title: "Maternal Mortality and Women's Empowerment: Exploring Socio-Political Factors Associated with Better Outcomes for Mothers Across the Globe"
author: "Justine Shakespeare"
desription: "Final project proposal for DACSS 603 - Updated 3/25"
date: "03/21/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart1
  - Justine Shakespeare
  - maternal mortality
  - women's empowerment
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(readxl)
library(summarytools)

knitr::opts_chunk$set(echo = TRUE)
```

## Research Question

Despite advancements in health and technology, maternal mortality remains a significant public health challenge around the world. According to the World Health Organization, approximately 287,000 women died from complications related to pregnancy or childbirth in 2020.[^1] Recently it was reported that global declines in maternal mortality have stalled,[^2] even increased in the United States,[^3] and data from 2020 has shown the pandemic brought a sharp uptick in maternal deaths.[^4] The World Health Organization has stated that the majority of maternal deaths are preventable.[^5]

[^1]: Maternal mortality, World Health Organization (2023) https://www.who.int/news-room/fact-sheets/detail/maternal-mortality

[^2]: Global Declines in Maternal Mortality Have Stalled, New York Times (2023) https://www.nytimes.com/2023/02/22/health/pregnancy-complications-death-who.html

[^3]: Singh GK. Trends and Social Inequalities in Maternal Mortality in the United States, 1969-2018. Int J MCH AIDS. 2021;10(1):29-42. doi: 10.21106/ijma.444. Epub 2020 Dec 30. PMID: 33442490; PMCID: PMC7792749. https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7792749/

[^4]: Maternal Deaths Rose During the First Year of the Pandemic, New York Times (2022) https://www.nytimes.com/2022/02/23/health/maternal-deaths-pandemic.html

[^5]: Maternal mortality, World Health Organization (2023) https://www.who.int/news-room/fact-sheets/detail/maternal-mortality

Past research has investigated variables associated with higher rates of maternal mortality, both at the level of the individual and country. Many studies have found that factors associated with healthcare, such as skilled birth attendance or antenatal visits, are strongly associated with maternal mortality.[^6] [^7] [^8] Other studies considering social determinants of health have found that factors such as adult literacy, education, income levels, and access to water and sanitation are all associated with maternal mortality.[^9] [^10] The majority of maternal deaths occur in low-income countries.[^11]

[^6]: Diana S, Wahyuni CU, Prasetyo B. Maternal complications and risk factors for mortality. J Public Health Res. 2020 Jul 3;9(2):1842. doi: 10.4081/jphr.2020.1842. PMID: 32728581; PMCID: PMC7376486. https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7376486/

[^7]: Betrán, A.P., Wojdyla, D., Posner, S.F. et al. National estimates for maternal mortality: an analysis based on the WHO systematic review of maternal mortality and morbidity. BMC Public Health 5, 131 (2005). https://doi.org/10.1186/1471-2458-5-131 https://bmcpublichealth.biomedcentral.com/articles/10.1186/1471-2458-5-131

[^8]: Girum T, Wasie A. Correlates of maternal mortality in developing countries: an ecological study in 82 countries. Matern Health Neonatol Perinatol. 2017 Nov 7;3:19. doi: 10.1186/s40748-017-0059-8. PMID: 29142757; PMCID: PMC5674830. <https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5674830/>

[^9]: Karlsen S, Say L, Souza JP, Hogue CJ, Calles DL, Gülmezoglu AM, Raine R. The relationship between maternal education and mortality among women giving birth in health care institutions: analysis of the cross sectional WHO Global Survey on Maternal and Perinatal Health. BMC Public Health. 2011 Jul 29;11:606. doi: 10.1186/1471-2458-11-606. PMID: 21801399; PMCID: PMC3162526. <https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3162526/>

[^10]: Girum T, Wasie A. Correlates of maternal mortality in developing countries: an ecological study in 82 countries. Matern Health Neonatol Perinatol. 2017 Nov 7;3:19. doi: 10.1186/s40748-017-0059-8. PMID: 29142757; PMCID: PMC5674830. <https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5674830/>

[^11]: Maternal mortality, World Health Organization (2023) <https://www.who.int/news-room/fact-sheets/detail/maternal-mortality>

This research intends to explore the relationship between women's empowerment and maternal mortality. In particular, this analysis will focus on variables related to women's political empowerment, civil liberties, and participation in civil society and will control for confounding variables related to economic development.

Past research has found a link between higher levels of gender equality and women's empowerment and reduced maternal mortality, although analysis has primarily focused on individual countries or a subset of countries (typically lower-income).[^12] [^13] [^14] This research will investigate whether these trends hold true when looking at over 165 countries from all levels of economic development.

[^12]: Chirowa F, Atwood S, Van der Putten M. Gender inequality, health expenditure and maternal mortality in sub-Saharan Africa: A secondary data analysis. Afr J Prim Health Care Fam Med. 2013 Aug 13;5(1):471. doi: 10.4102/phcfm.v5i1.471. PMCID: PMC4709496. https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4709496/

[^13]: Ahmed F, Oni FA, Hossen SS. Does gender inequality matter for access to and utilization of maternal healthcare services in Bangladesh? PLoS One. 2021 Sep 16;16(9):e0257388. doi: 10.1371/journal.pone.0257388. PMID: 34529701; PMCID: PMC8445442. https://pubmed.ncbi.nlm.nih.gov/34529701/

[^14]: MATERNAL MORTALITY AND WOMEN\'S POLITICAL POWER, Sonia R. Bhalotra, Damian Clarke, Joseph F. Gomes, Atheendar, Venkataramani. NATIONAL BUREAU OF ECONOMIC RESEARCH, (June 2022) https://www.nber.org/system/files/working_papers/w30103/w30103.pdf

In short, this analysis seeks to answer the question: How do factors related to women's empowerment impact maternal mortality?

## Hypothesis

**Hypothesis 1:** Where women are politically empowered, have civil liberties, and are able to freely participate in civil society, health outcomes are better and the maternal mortality ratio is lower. This will hold true even when controlling for variables which are known to have an effect on both women's empowerment and maternal mortality.

*Note for professor: Another way that this analysis could address the research question above is to explore the mechanisms through which women's empowerment affects maternal mortality. Specifically, since much of the literature discusses immediate healthcare concerns this analysis could examine (in addition to the link between women's empowerment and MMR) whether women's empowerment is associated with the two variables: skilled birth attendance and antenatal visits. In this case, the second hypothesis would be:*

**Hypothesis 2:** Women's political empowerment and their participation in civil society has a positive effect on skilled birth attendance and antenatal visits, resulting in a reduced maternal mortality ratio.

*In this way the analysis will both **establish the link** between women's empowerment and reduced maternal mortality while also **providing insight into how** women's empowerment affects maternal mortality.*

## Introduction to the Data

This research will utilize several datasets in order to capture all of the relevant variables.

The dependent variable, **Maternal Mortality Ratio**, is drawn from the *Maternal and newborn health* table from *UNICEF*. This data is from 2017.[^15]

[^15]: The State of the World's Children 2021: Statistical tables, UNICEF (2021) https://data.unicef.org/resources/dataset/the-state-of-the-worlds-children-2021-statistical-tables/

::: callout-tip
## Definitions

According to the **World Health Organization**, the **maternal mortality ratio (MMR)** is defined as: "the number of maternal deaths during a given time period per 100,000 live births during the same time period."

**Maternal deaths** are defined as: "deaths from any cause related to or aggravated by pregnancy or its management (excluding accidental or incidental causes) during pregnancy and childbirth or within 42 days of termination of pregnancy..."[^16]
:::

[^16]: Indicator Metadata Registry List, https://www.who.int/data/gho/indicator-metadata-registry/imr-details/26

The **independent variables** featured in this analysis will be drawn from the *Varieties of Democracy V-Dem* dataset.[^17] From there, the following three variables that relate to gender equality and women's rights will be used:

[^17]: Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, David Altman, Michael Bernhard, Agnes Cornell, M. Steven Fish, Lisa Gastaldi, Haakon Gjerløw, Adam Glynn, Ana Good God, Sandra Grahn, Allen Hicken, Katrin Kinzelbach, Joshua Krusell, Kyle L. Marquardt, Kelly McMann, Valeriya Mechkova, Juraj Medzihorsky, Natalia Natsika, Anja Neundorf, Pamela Paxton, Daniel Pemstein, Josefine Pernes, Oskar Rydén, Johannes von Römer, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Aksel Sundström, Eitan Tzelgov, Yi-ting Wang, Tore Wig, Steven Wilson and Daniel Ziblatt. 2023. "V-Dem \[Country-Year/Country-Date\] Dataset v13" Varieties of Democracy (V-Dem) Project. https://doi.org/10.23696/vdemds23.

-   ***Women civil society participation index***[^18]

    As defined by the V-Dem codebook: "Women's civil society participation is understood to include open discussion of political issues, participation in civil society organizations, and representation in the ranks of journalists."[^19]

-   ***Women civil liberties index***[^20]

    As defined by the V-Dem codebook: "Women's civil liberties are understood to include freedom of domestic movement, the right to private property, freedom from forced labor, and access to justice."[^21]

-   ***Women political empowerment index***[^22]

    As defined by the V-Dem codebook: "Women's political empowerment is defined as a process of increasing capacity for women, leading to greater choice, agency, and participation in societal decision-making. It is understood to incorporate three equally-weighted dimensions: fundamental civil liberties, women's open discussion of political issues and participation in civil society organizations, and the descriptive representation of women in formal political positions."[^23]

[^18]: Sundström et al. (2017, V-Dem Working Paper Series 2017:19); Pemstein et al. (2023, V-Dem Working Paper Series 2023:21); V-Dem Codebook (see suggested citation at the top of this document).

[^19]: Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, David Altman, Michael Bernhard, Agnes Cornell, M. Steven Fish, Lisa Gastaldi, Haakon Gjerløw, Adam Glynn, Sandra Grahn, Allen Hicken, Katrin Kinzelbach, Kyle L. Marquardt, Kelly McMann, Valeriya Mechkova, Anja Neundorf, Pamela Paxton, Daniel Pemstein, Oskar Rydén, Johannes von Römer, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Aksel Sundström, Eitan Tzelgov, Luca Uberti, Yi-ting Wang, Tore Wig, and Daniel Ziblatt. 2023. "V-Dem Codebook v13" Varieties of Democracy (V-Dem) Project.

[^20]: Sundström et al. (2017, V-Dem Working Paper Series 2017:19); Pemstein et al. (2023, V-Dem Working Paper Series 2023:21); V-Dem Codebook (see suggested citation at the top of this document).

[^21]: Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, David Altman, Michael Bernhard, Agnes Cornell, M. Steven Fish, Lisa Gastaldi, Haakon Gjerløw, Adam Glynn, Sandra Grahn, Allen Hicken, Katrin Kinzelbach, Kyle L. Marquardt, Kelly McMann, Valeriya Mechkova, Anja Neundorf, Pamela Paxton, Daniel Pemstein, Oskar Rydén, Johannes von Römer, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Aksel Sundström, Eitan Tzelgov, Luca Uberti, Yi-ting Wang, Tore Wig, and Daniel Ziblatt. 2023. "V-Dem Codebook v13" Varieties of Democracy (V-Dem) Project.

[^22]: Sundström et al. (2017, V-Dem Working Paper Series 2017:19); V-Dem Codebook (see suggested citation at the top of this document).

[^23]: Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, David Altman, Michael Bernhard, Agnes Cornell, M. Steven Fish, Lisa Gastaldi, Haakon Gjerløw, Adam Glynn, Sandra Grahn, Allen Hicken, Katrin Kinzelbach, Kyle L. Marquardt, Kelly McMann, Valeriya Mechkova, Anja Neundorf, Pamela Paxton, Daniel Pemstein, Oskar Rydén, Johannes von Römer, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Aksel Sundström, Eitan Tzelgov, Luca Uberti, Yi-ting Wang, Tore Wig, and Daniel Ziblatt. 2023. "V-Dem Codebook v13" Varieties of Democracy (V-Dem) Project.

All three of these variables are on an interval scale, from low to high (0-1). In order to match the timeframe of the dependent variable, the 2017 value for all of these variables will be used.

In addition to these variables, a **confounding variable** that seeks to account for a nation's wealth and development will be considered in order to control for its effect on both maternal mortality and women's empowerment.

**GDP per capita** will be used as a way to control for a nation's economic development. This variable is drawn from the *International Monetary Fund (IMF)'s* data.[^24] GDP per capita from 2017 will be used in order to be consistent with the data in the rest of the analysis dataset.

[^24]: https://www.imf.org/external/datamapper/NGDPDPC\@WEO/OEMDC/ADVEC/WEOWORLD

The resulting database will be a cross-sectional, cross-national sample of data from 169 countries (some observations were removed because of missing values).

## Loading and Cleaning the Data

First let's load and set up the data. Because we are drawing from multiple datasets for this analysis, this will be a fairly involved process. First we'll load and clean up the Maternal and Newborn Health dataset from UNICEF.

```{r}
library(tidyverse)
library(readxl)
library(summarytools)

# Load the dataset, skip the first five rows, rename all of the variables we 
# want and delete the ones we don't
MaternalNewbornHealth <- read_excel("UNICEF_mat_newborn.xlsx", skip = 6, 
                                    col_names= c("Countries_and_areas", 
                                                 rep("delete", 28), 
                                                 "Maternal_Mortality_Ratio_2017", 
                                                 "delete", "delete")) %>% 
  select(!contains("delete"))

# remove the last 57 rows, which include summaries and endnotes, not data
MNH <- slice(MaternalNewbornHealth, 1:(nrow(MaternalNewbornHealth) - 57)) 

# select just the variables we need for this analysis
MNH_trim <- MNH %>% 
  select(Countries_and_areas, Maternal_Mortality_Ratio_2017) 

# instead of NAs, this dataset has " - " where there are no values. We'll 
# remove those here.
MNH_trim <- MNH_trim %>% 
  filter(!str_detect(`Maternal_Mortality_Ratio_2017`, "-"))

# change the Maternal Mortality variable to numeric.
MNH_trim$`Maternal_Mortality_Ratio_2017` <- as.numeric(MNH_trim$`Maternal_Mortality_Ratio_2017`)

# take a quick look at our UNICEF data
head(MNH_trim)
```

Next let's add in the variables related to women's empowerment. We'll first need to load in the Varieties of Democracy dataset, then we'll select the appropriate variables and clean it up a bit. 

```{r}
V_DEM <- readRDS("V-Dem-CY-Full+Others-v13.rds")

# Select just the variables we need for this analysis and filter to year 2017.
V_DEM_trim <- V_DEM %>% 
  filter(year == 2017) %>% 
  select(country_name, v2x_gender, v2x_gencl, 
       v2x_gencs)

# rename the country_name variable so we can join it with the UNICEF data later. 
V_DEM_trim <- rename(V_DEM_trim, "Countries_and_areas" = country_name)

# Change some of the country names where the V-Dem dataset differs from the 
# UNICEF data. 
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Moldova"] <- "Republic of Moldova"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Laos"] <- "Lao People's Democratic Republic"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "The Gambia"] <- "Gambia"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Republic of the Congo"] <- "Congo"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Syria"] <- "Syrian Arab Republic"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Iran"] <- "Iran (Islamic Republic of)"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Cape Verde"] <- "Cabo Verde"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Ivory Coast"] <- "Côte d'Ivoire"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Venezuela"] <- "Venezuela (Bolivarian Republic of)"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Tanzania"] <- "United Republic of Tanzania"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "South Korea"] <- "Republic of Korea"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "North Korea"] <- "Democratic People's Republic of Korea"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Vietnam"] <- "Viet Nam"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Bolivia"] <- "Bolivia (Plurinational State of)"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "United States of America"] <- "Gambia"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Russia"] <- "Russian Federation"
V_DEM_trim$Countries_and_areas[V_DEM_trim$Countries_and_areas == "Burma/Myanmar"] <- "Myanmar"

# take a quick look at our V-Dem data
head(V_DEM_trim)
```

Next, we'll load anc clean the IMF dataset with the data on GDP per capita.

```{r}
imf_GDP <- read_excel("imf-dm-export-20230322.xls")

# Select just the variables we need for this analysis.
imf_GDP_2017 <- select(imf_GDP, `GDP per capita, current prices\n (U.S. dollars per capita)`, `2017`)

# rename the country_name variable so we can join it with the UNICEF data later.
imf_GDP_2017 <- rename(imf_GDP_2017, "Countries_and_areas" = `GDP per capita, current prices\n (U.S. dollars per capita)`)

# change the GDP per capita variable to numeric.
imf_GDP_2017$`2017` <- as.numeric(imf_GDP_2017$`2017`)

# Change some of the country names where the IMF dataset differs from the 
# UNICEF data. 
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Bahamas, The"] <- "Bahamas"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Bolivia"] <- "Bolivia (Plurinational State of)"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "China, People's Republic of"] <- "China"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Congo, Dem. Rep. of the"] <- "Democratic Republic of the Congo"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Congo, Republic of"] <- "Congo"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Czech Republic"] <- "Czechia"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Gambia, The"] <- "Gambia"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Iran"] <- "Iran (Islamic Republic of)"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Korea, Republic of"] <- "Republic of Korea"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Kyrgyz Republic"] <- "Kyrgyzstan"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Lao P.D.R."] <- "Lao People's Democratic Republic"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Micronesia, Fed. States of"] <- "Micronesia (Federated States of)"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Moldova"] <- "Republic of Moldova"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Slovak Republic"] <- "Slovakia"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "South Sudan, Republic of"] <- "South Sudan"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Syria"] <- "Syrian Arab Republic"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "São Tomé and Príncipe"] <- "Sao Tome and Principe"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Tanzania"] <- "United Republic of Tanzania"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Türkiye, Republic of"] <- "Turkey"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Venezuela"] <- "Venezuela (Bolivarian Republic of)"
imf_GDP_2017$Countries_and_areas[imf_GDP_2017$Countries_and_areas == "Vietnam"] <- "Viet Nam"

# take a quick look at our IMF data
head(imf_GDP_2017)
```

Finally, let's join together all three datasets!
```{r}
# first we'll join the UNICEF and V-DEM data.
JOIN_UNICEF_VDEM <- full_join(MNH_trim, V_DEM_trim, by = "Countries_and_areas") 

# then we'll add in the IMF data.
JOIN_UNICEF_VDEM_IMF <- full_join(JOIN_UNICEF_VDEM, imf_GDP_2017, by = "Countries_and_areas") 

# remove the last 56 rows, which include notes, countries, and regions not in 
# the original UN dataset
MMR_Combined <- slice(JOIN_UNICEF_VDEM_IMF, 1:(nrow(JOIN_UNICEF_VDEM_IMF) - 56)) 

# rename our variables for ease of use
MMR_Analysis <-rename(MMR_Combined, "Country" = Countries_and_areas, 
                      "PolEmp" = v2x_gender, "CivLib" = v2x_gencl, 
                      "CivSoc" = v2x_gencs, "GDP_2017" = `2017`)

# remove NAs for anlaysis
MMR_Analysis <- na.omit(MMR_Analysis)

# Let's take a look!
head(MMR_Analysis, limit = 10)
glimpse(MMR_Analysis)
```
## Descriptive Statistics

Let's take a closer look at the dependent variable, the Maternal Mortality Ratio. We can check out the mean, standard deviation, the full range (the lowest and highest values), values in the 25th, and 75th quantiles, as well as the median. This will give us a sense of the full spread of this variable.

```{r}
MMR_spread <- MMR_Analysis %>% 
  summarize("Mean" = mean(Maternal_Mortality_Ratio_2017, na.rm = TRUE),
            "Standard_Deviation" = sd(Maternal_Mortality_Ratio_2017, na.rm = TRUE),
            "Lowest" = min(Maternal_Mortality_Ratio_2017, na.rm = TRUE), 
            "25th quantile" = quantile(x = Maternal_Mortality_Ratio_2017, probs = .25),
            "Median" = median(Maternal_Mortality_Ratio_2017, na.rm = TRUE),
            "75th quantile" = quantile(x = Maternal_Mortality_Ratio_2017, probs = .75),
            "Highest" = max(Maternal_Mortality_Ratio_2017, na.rm = TRUE))
MMR_spread
```

We can see that there is quite a range, from 2 to 1,150. The relatively large standard deviation indicates there is a fair amount of variability as well. We can see from the median and 75th quantile numbers that most of the countries in this data have MMRs lower than about 250. But the highest value here shows us that some countries have an MMR of well over 4 times that amount.

Let's take a look at which countries have the highest MMRs (the most deaths per 100,000 live births).

```{r}
MMR_high <- MMR_Analysis %>% 
  select("Country", "Maternal_Mortality_Ratio_2017") %>% 
  arrange(desc(as.numeric(Maternal_Mortality_Ratio_2017)))

print(MMR_high, limit = 10)
```

Let's see which countries have the lowest MMRs.

```{r}
MMR_low <- MMR_Analysis %>% 
  select("Country", "Maternal_Mortality_Ratio_2017") %>% 
  arrange(as.numeric(Maternal_Mortality_Ratio_2017))

print(MMR_low, limit = 10)
```

Finally, we'll take a look at some of the explanatory variables we're interested in exploring.

```{r}
summary(MMR_Analysis[, c("PolEmp", "CivLib", "CivSoc", "GDP_2017")])
dfSummary(MMR_Analysis[, c("PolEmp", "CivLib", "CivSoc", "GDP_2017")])
```








