---
title: "Kristin Abijaoude_HW4"
editor: visual
date: "04/22/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Hw4
  - kristin abijaoude
---

# Question 1

ŷ = −10,536 + 53.8x1 + 2.84x2

y = selling price of home (in dollars) x1 = size of home (in square feet) x2 = lot size (in square feet)

A. particular home of 1240 square feet on a lot of 18,000 square feet sold for \$145,000. Find the predicted selling price and the residual, and interpret.

```{r}
price = -10536 + (53.8 * 1240) + (2.84 * 18000)
print(price)

residual = price - 145000
print(residual)
```

The predicted price is `$107,296`, which is much under the `$145,000` price sold by `-$37,704`.

B. For fixed lot size, how much is the house selling price predicted to increase for each square- foot increase in home size? Why?

It's `$53.80` for every 1x increase in square feet of the selling house, as shown in the equation as a coefficient of 1x.

C. According to this prediction equation, for fixed home size, how much would lot size need to increase to have the same impact as a one-square-foot increase in home size?

You need to increase the lot size by 18.94 sq feet to have the same impact as a one square foot increase.

```{r}
# x1 / x2

lot = 53.80 / 2.84
print(lot)
```

# Question 2

```{r}
library(alr4)
data(salary, package = "alr4")
head(salary)
```

A. Test the hypothesis that the mean salary for men and women is the same, without regard to any other variable but sex.

With a p-value of 0.09, we can conclude that the mean salary for both men and women is the same, thus accepting the null hypothesis.

```{r}
sex <- select(salary, c(sex))
salary1 <-  select(salary, c(salary))

sex <- as.numeric(unlist(sex))
salary1 <- as.numeric(unlist(salary1))

mean <- t.test(salary1 ~ sex, var.equal = FALSE, alternative = "two.sided")
print(mean)
```

B. Run a multiple linear regression with salary as the outcome variable and everything else as predictors, including sex.

```{r}
model <- lm(formula = salary ~ ., data = salary)
summary(model)
```

```{r}
confint(model, 'sexFemale', level=0.95)
```

```{r}
confint(model, level=0.95)
```

C. Interpret your finding for each predictor variable; discuss (a) statistical significance, (b) interpretation of the coefficient / slope in relation to the outcome variable and other variables

`degreePD` - This means that, all else being equal, PhD holders can expect to earn an average of `$1,388.60` more in salary.

`rankAssoc` - Associate professors can expect to earn an average of `$5,292.40` more in salary.

`rankProf` - Professors would earn an average of `$11,118.80` more in salary.

`sexFemale` - Female faculty workers would earn an average of `$1,166.4` more in salary.

`year` - The longer a faculty member works at a college, the more they earn in salary, with an average of `$476.30` in increase.

`ysdeg` - However, if it's been several years since you earned your last degree, expect a decrease of `-$124.60` in salary on average.

D. Change the baseline category for the rank variable. Interpret the coefficients related to rank again.

```{r}
salary$rank <- relevel(salary$rank, ref = "Prof")
model2 <- lm(formula = salary ~ ., data = salary)
summary(model2)
```

In the new model, we can see that assistant professors and associate professors lose an average of `-$11,118.80` and `-$5,826.40` in salary, respectively.

E Removing `rank` from the model

```{r}
model3 <- lm(formula = salary ~ degree + sex + year + ysdeg, data = salary)
summary(model3)
```

When removing `rank` from the equation, we can see a decrease in the average salary for PhD holders and female faculty members by `-$3,299.30` and `-$1,286.50`, respectively. In addition, the salary average for year is smaller than before, but it's still an increase. The biggest shift is `ysdeg` with `$339.40` increase in salary on average.

F New variable, new hypothesis

Some people have argued that the new Dean has been making offers that are a lot more generous to newly hired faculty than the previous one and that this might explain some of the variation in Salary.

```{r}
salary <- salary %>%
  mutate(ysdeg15 = ifelse(ysdeg <= 15, 1,0))
salary

model4 <- lm(formula = salary ~ degree + sex + year + ysdeg15, data = salary)
print(model4)

cor.test(salary$ysdeg, salary$ysdeg15)
```

I took out the `ysdeg` variable as they're too similar to the new variable I created `ysdeg15` to avoid multicollinearity.

The correlation is `-0.8434239`, in which we can reject that alternative hypothesis. In other words, there are no changes in salary average.

# Question 3

```{r}
library(smss)
data("house.selling.price", package = "smss")
head(house.selling.price)
```

A. Using the house.selling.price data, run and report regression results modeling y = selling price (in dollars) in terms of size of home (in square feet) and whether the home is new (1 = yes; 0 = no).

```{r}
selling <- lm(formula = Price ~ Size + New, data = house.selling.price)
summary(selling)
```

A new house sells at `$57,736.30` more, while for every one square foot is sold at `$116.10`.

B. Report and interpret the prediction equation, and form separate equations relating selling price to size for new and for not new homes.

y = `-40,230` + `116.10`x1 + `57,736.30`x2

C. Find the predicted selling price for a home of 3000 square feet that is (i) new, (ii) not new.

A not new house would be sold at `$308,070`

```{r}
price1 = -40230 + (116.10 * 3000) + (57736.30 * 0)
print(price1)
```

D. Fit another model, this time with an interaction term allowing interaction between size and new, and report the regression results

```{r}
selling2 <- lm(formula = Price ~ Size*New, data = house.selling.price)
summary(selling2)
```

E. Report the lines relating the predicted selling price to the size for homes that are (i) new, (ii) not new.

```{r}
# new
price2 = -22227.808 + (104.44 * 3000 ) - (78527.50 * 1) + (61.92 * 1)
price2

# not new 
price3 = -22227.808 + (104.44 * 3000 ) + 61.92
price3
```

F. Find the predicted selling price for a home of 3000 square feet that is (i) new, (ii) not new.

A new house would sell at `$212,626.60`, while a not new house would sell at `$291,154.10`.

G. Find the predicted selling price for a home of 1500 square feet that is (i) new, (ii) not new. Comparing to (F), explain how the difference in predicted selling prices changes as the size of home increases.

```{r}
# new
price4 = -40230 + (116.10 * 1500) + (57736.30 * 1)
price4

# not new 
price5 = -40230 + (116.10 * 1500) 
price5
```

A new house would sell at `$191,656.30`, less than the `$212,626.60` in Question F, and a not new house would sell at `$133,920`, also less than the `$291,154.10` in Question F as well.

H. Do you think the model with interaction or the one without it represents the relationship of size and new to the outcome price? What makes you prefer one model over another?

I think the model without the interaction represents the relationship of size and newness to the outcome price because it's simpler, easier to interpret, and there is more statistical significance as opposed to the model with the interaction. Despite that, the RSS, R-square, and adjusted R-square are similar.
