---
title: "Hw 2 by Kristin Abijaoude"
author: "Kristin Abijaoude"
description: "HW2"
date: "03/16/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Hw2
  - kristin abijaoude
  - distribution 
  - probability
---

```{r}
# load packages
packages <- c("readr", "ggplot2", "caret", "summarytools", "tidyverse", "dplyr", "stats", "pwr")
lapply(packages, require, character.only = TRUE)
```

# Question 1

```{r}
surgical_procedures <- c("bypass","angiography")
sample_size <- c(589, 847)
mean_wait_time <- c(19, 18)
standard_deviation <- c(10,9)

surgery_data <- data.frame(surgical_procedures, sample_size, mean_wait_time, standard_deviation)
surgery_data
```
Wait times are in days. 

```{r}
# confidence level for bypass and angiography
conf_level <- 0.9

# standard error for bypass
bypass_se <- 10 / sqrt(589)

# confidence interval for bypass
bypassCI <- 19 + qt(c(0.05, 0.95), 589-1) * bypass_se
bypassCI
```

```{r}
# standard error for angiography
angio_se <- 9 / sqrt(847)

# confidence interval for angiography
angioCI <- 18 + qt(c(0.05, 0.95), 847-1) * angio_se
angioCI
```

# Question 2 

```{r}
# out of 1031 Americans surveyed
p <- 567 / 1031 
# 54% of Americans believe college education is essential for success

# 95% significant level
conf<- 0.95 

# standard error
college_se <- sqrt(p*(1-p)/1031) 

# confidence interval
collegeCI <- p + qnorm(c(0.025, 0.975)) * college_se
collegeCI 
```

# Question 3

```{r}
# $5 or less within the estimate
est <- 5

# money spent on textbooks varies widely, mostly between $30 and $200
sigma <- (200 - 30) / 4

# significant level 
alpha <- 0.5

# z-alpha
z_alpha <- qnorm(1 - alpha / 2)

# sample size of books
n <- ceiling((z_alpha * sigma / est) ^ 2)
n
```

# Question 4
## A
```{r}
# t test
f_emp <- 410
income <- 500
s <- 90
t <- (f_emp - income) / (s / sqrt(9))
t
```

## B
```{r}
# degree of freedom
df <- 9 - 1 
# df = 8

# p-value
p_value <- pt(t, df)
p_value

# significant level
alpha <- 0.05

# to reject or not to reject
if (p_value < alpha/2 || p_value > 1-alpha/2) {
  cat("Reject the null hypothesis")
} else {
  cat("Fail to reject the null hypothesis")
}
```

## C
```{r}
1-p_value

# fail to reject null hypothesis
```

# Question 5
## A

```{r}
#  T values
t_jones <- (519.5 - 500) / 10 # sample mean = 519.5 - 500 for population mean / sample error of 10.0
t_jones

t_smith <- (519.7 - 500) / 10 # sample mean = 519.7 - 500 for population mean / sample error of 10.0
t_smith
```

```{r}
# p values
p_jones <- 2 * pt(-abs(t_jones), df = 999)
p_jones

p_smith <- 2 * pt(-abs(t_smith), df = 999)
p_smith
```

## B

Smith's result is statistically significant, while Jones' is not.

## C

While the two results are close in variables, one of them is significant while the other is not. That's why P-value is important in determining whether to reject or fail to reject the hypothesis. 

# Question 6

```{r}
healthy <- c(31, 43, 51)
unhealthy <- c(69, 57, 49)

snack <- rbind(healthy, unhealthy)

colnames(snack) <- c("6th grade", "7th grade", "8th grade")
rownames(snack) <- c("healthy", "unhealthy")

snack
# α = 0.05
```
```{r}
chisq.test(snack, correct = FALSE)
```

Since the p-value is smaller than the 0.05 threshold, we can conclude that there is an association between grade and snack choices.

# Question 7

```{r}
tuition <- data.frame(
  area = c(rep("Area_1", 6), rep("Area_2", 6), rep("Area_3", 6)),
  cost = c(6.2, 9.3, 6.8, 6.1, 6.7, 7.5, 7.5, 8.2, 8.5, 8.2, 7.0, 9.3, 5.8, 6.4, 5.6, 7.1, 3.0, 3.5))

tuition
```

```{r}
tuition_model <- aov(cost ~ area, data = tuition)
summary(tuition_model)
```

From these results, we can conclude that there is not much statistical significance between the area of a given charter school and the cost of the tuition in said charter schools.