---
title: "Final Project Checkin-2"
author: "Xiaoyan"
description: "Template of course blog qmd file"
date: "04/17/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart1
---
```{r}
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
```
# {.tabset}

## Introduction and background

 The Chinese government implemented the one-child policy in 1979, which resulted in the increasing proportion of one-child families and the "four-two-one" family structure consisting of four grandparents, two parents, and one child. Despite being blessed with relatively more family and social resources, only children may face physical and socio-psychological problems during development, including an elevated risk for overweight and obesity and negative psychosocial consequences. Previous studies have shown that only children had a higher likelihood of overweight or obesity, compared with children who had one or more siblings. Over obesity, mental healthy is also interesting to explore that how it is related to overweight/obesity, as well as sib-size, in young adolescents affects mental health.。
 

## research questions
1. Does obesity positively related to mental health?
2. what are factors that affects mental healthy?
3. does sibling or obeisty directily related to mental health?

## key predictors
1. mental health
2. sibling number
3. obisity rate
4. Family location, finance and education


## hypothesis
1. Higher obesity rate increase the risk of depression
2. higher family income increase the rate of obesity
3. More sibling reduce the risk of both depression and anxiety. 

In these hypothesis, the response variables are depression rate, axiety rate and BMI index. The explanatory variables can be factors listed below. Analysis is needed to identify the control variables. For exapmle, in hypothesis 2, family income is the explanatory varible and rate of obsity(BMI) is response varible, the control varible may also be family financial situation. 

## data description
### overlook of data
```{r}
data<-read_excel("/Users/cassie199/Desktop/23spring/603_Spring_2023-1/posts/_data/mentalhealth_data.xlsx")
head(data)
sum(is.na(data))
plot(data$T0depression~data$BMI)
```
This dataset including 1348 variables and 29 columns. there are 728 NA in this data set. all variables was presented as numberic data. descriptive data was also presented as degrees such as education level, family financial situation and depression rate. By pre-plotting depression rate vs BMI, we can see that some ouliers may need to deal with and there is no siginifcant disrtibution on graph. More data processing is needed in future process.

```{r}
variables <- c("Internalizing problem - Depression (SDS)", "Internalizing problem - Anxiety (SAS)", 
               "Obesity parameters - BMI", "Obesity parameters - WC", "Obesity parameters - WHR",
               "Obesity parameters - WHtR", "Biochemical parameters - TG", "Biochemical parameters - FBG",
               "Biochemical parameters - TC", "Biochemical parameters - HDL-C", "Biochemical parameters - LDL-C",
               "Blood pressure - SBP", "Blood pressure - DBP","Family location", "Number of siblings", " time  spend with father in elementary school?", 
          " time spend with mother in elementary school?", "Father’s education level", 
          "Mother’s education level", "Family financial situation", "Sleeping hours", "Skipping breakfast", 
          "Vigorous", "Moderate")
abreviations <- c("Depression", "Anxiety", "BMI", "WC", "WHR", "WHtR", "TG", "FBG", "TC", "HDL-C", "LDL-C", "SBP", "DBP","FL", "NS", "TFE", "TME", "FEL", "MEL", "FS", "SL", "SB", "VG", "MD")


cat("varible table\n")
variable_table <- data.frame(variables, abreviations)
variable_table

```
### parameter explaination
```{r}
# Create the data frame for SAS and SDS scales
sas_levels <- c("Normal", "Mild to Moderate", "Marked to Severe", "Extreme")
sas_scores <- c("<45", "45-59", "60-74", ">=75")
sas_table <- data.frame(Level = sas_levels, Score = sas_scores)

sds_levels <- c("Normal", "Mild", "Moderate to Marked Major", "Severe or Extreme Major")
sds_scores <- c("<50", "50-59", "60-69", ">=70")
sds_table <- data.frame(Level = sds_levels, Score = sds_scores)

# Create the data frame for BMI categories
bmi_levels <- c("Underweight", "Normal Weight", "Overweight", "Obesity")
bmi_values <- c("<18.5", "18.5-24.9", "25-29.9", ">=30")
bmi_table <- data.frame(Category = bmi_levels, BMI = bmi_values)

# Print the SAS scale table
cat("Self-rating Anxiety Scale (SAS)\n")
print(sas_table)

# Print the SDS scale table
cat("\nSDS scores (SDS)\n")
print(sds_table)

# Print the BMI category table
cat("\nBMI Categories\n")
print(bmi_table)
```
## hypothesis test

### 1. Higher obesity rate increase the risk of depression
```{r}
colnames(data)<-c("T0depression","T0anxiety","T1depression","T1anxiety","Height","Weight","WC","HC","SBP","DBP","FBG","TC","TG","HDL-C","LDL-C","BMI","WHR","WtHR","FL", "NS", "TFE", "TME", "FEL", "MEL", "FS", "SL", "SB","Vigorous","Moderate")

ggplot(data, aes(x = T1depression, y = BMI)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
cor.test(data$T1depression, data$BMI,method = c("spearman"))
cor.test(data$T1depression, data$BMI,method = c("pearson"))
```
The Pearson correlation test is a statistical test used to measure the linear relationship between two continuous variables.The Spearman's rank correlation coefficient (rho) measures the strength and direction of the association between two variables which don't have to be both continuous or have a linear relationship. It ranges between -1 and 1, where -1 indicates a perfect negative correlation, 1 indicates a perfect positive correlation, and 0 indicates no correlation.
Due to the High p value of both test, the depression rate is less likely related to BMI. 

### 2. higher family income increase the rate of obesity
```{r}
ggplot(data, aes(x = FS, y = BMI)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
cor.test(data$FS, data$BMI,method = c("spearman"))
cor.test(data$FS, data$BMI,method = c("pearson"))
```
No significant corelationship on family financial status and obesity. 

### 3. More sibling reduce the risk of both depression and anxiety. 
```{r}
ggplot(data, aes(x = NS, y = T1depression)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
cor.test(data$T1depression,data$NS, method = c("spearman"))
cor.test(data$T1depression,data$NS, method = c("pearson"))
data$NS <- factor(data$NS)
t.test(T1depression ~ NS, data = data)
```


In this case, the sample estimate of the correlation coefficient (rho) is 0.1184734, indicating a positive correlation between T0depression and NS. However, the p-value of the test is 0.008575, which is less than 0.05, suggesting that the correlation is statistically significant at a 5% level of significance.

Therefore, we can conclude that there is a significant positive correlation between the number of siblings (NS) and the degree of depression  in this dataset.

By carrying out a Welch t-test, the group with more siblings have higher depression index and p value <0.05 indicates the result is siginifcant. (Not sure why the confident interval is negtive and none of the data was negative. )

```{r}
ggplot(data, aes(x = NS, y = T1anxiety)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
data$NS <- as.numeric(data$NS)
cor.test(data$T1anxiety,data$NS, method = c("spearman"))

```
### others
```{r}

# Create sample data

plot(data$BMI~data$T1depression)
plot(data$T1depression~data$FL)
data$BMI_category <- cut(data$BMI, 
                       breaks = c(-Inf, 18.5, 24.9, 29.9, Inf),
                       labels = c("Underweight", "Normal weight", "Overweight", "Obesity"))
data$Depression_category <- cut(data$T1depression, 
                       breaks = c(0,45,	59,74,75),
                       labels = c("Normal", "Mild", "Moderate to Marked Major", "Severe or Extreme Major"))
# Plot the bar chart



```
 


## Answers to the feedbacks on check in 1

Here are a few things you may want to work on in future steps:
1. Please provide more information of the dataset: what each variable means (e.g. WC, HC, SBP etc) and how it is measured. This is to make sure audiences understand your confounders.
  a table with explaination of abrivation is updated in data description

2. Since gender is one of your key predictors, you may consider using the interaction between gender and other key variables in the model to see whether gender influences the impact of other predictors. Also, seem I didn't find the gender variable in the dataset you provided?
  Thanks for pointing out. Since gender is missiong, I will not use gender as a key predicor. 
  
3. As you mentioned, there are some outliers in the data, especially the one on the top-right corner. This outlier can change the slope of the regression. Also, the relationship between BMI and depression is not very clear in the graph, as you mentioned, more data processing is needed. You can also try plotting different groups (e.g. gender, family location) in different colors to see if there's any pattern.

Thanks for the comments, I will try to process the data this time  and plot more patterns. 

## Questions need to be addressed
1. the varibles such as family locations or education level can be expressed either as rank or ordinal, as drawed below,it is hard to find a correlationship with this kind of varibles. How can i explore the relationship between an ordinal varible and a continuous varible?
```{r}
pairs(data[c("T1depression","T1anxiety","BMI","FL", "NS", "TFE", "TME", "FEL")])
pairs(data[c("T1depression","T1anxiety", "MEL", "FS", "SL", "SB","Vigorous","Moderate")])
```
2. as some of the continuous varibles can also converted to ordinal varibles, what would be some method or test good to find the relationship between them?
```{r}
#convert continuous varibles into categorical varibles
data$FL1 <- factor(sample(1:5, 1348, replace = TRUE), levels = 1:5, 
                            labels = c("Rural", "Suburban", "Urban", "City", "Metropolis"))
data$BMI_category <- cut(data$BMI, 
                       breaks = c(-Inf, 18.5, 24.9, 29.9, Inf),
                       labels = c("Underweight", "Normal weight", "Overweight", "Obesity"))
data$Depression_category <- cut(data$T1depression, 
                       breaks = c(0,45,	59,74,75),
                       labels = c("Normal", "Mild", "Moderate to Marked Major", "Severe or Extreme Major"))
#plot
ggplot(data, aes(x = FL1, y = BMI, fill = BMI_category)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("#1b9e77", "#d95f02", "#7570b3", "#e7298a")) +
  xlab("Family location") + ylab("BMI") + 
  ggtitle("BMI category and family location") +
  theme_bw()
ggplot(data, aes(x = FL1, y = T1depression, fill = Depression_category)) + 
  geom_bar(stat = "identity", position = "stack")
```