---
title: "Homework 1"
author: "Zhiyuan Zhou"
description: "Homework 1"
date: "02/28/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---
```{r}
library(dplyr)
```
# Question 1

## a

First, let's read in the data from the Excel file:

```{r, echo=T}
library(readxl)
df <- read_excel("_data/LungCapData.xls")
```


The distribution of LungCap looks as follows:

```{r, echo=T}
hist(df$LungCap)
```

The histogram suggests that the distribution is close to a normal distribution. Most of the observations are close to the mean. Very few observations are close to the margins (0 and 15).

##b
```{r, echo=T}
boxplot(df$LungCap~df$Gender,
main = "Lung Capacity by Gender",
xlab = "Gender",
ylab = "Lung Capacity",
)
```
##c
```{r, echo=T}
df %>%
  group_by(Smoke) %>%
  summarize(mean = mean(LungCap))
```
This result surprised me that smokers have more lung capacity than non-smokers.

##d
```{r}
df["AgeGroup"] = 
  cut(df$Age,
      c(0, 13, 15, 17, Inf),
      c("<=13", "14-15","16-17", ">=18"),
      right = T
  )

df%>%
  group_by(AgeGroup, Smoke)%>%
  summarize(meanLungCap = mean(LungCap), meanAge = mean(Age), count = n())
```

##e
In age group "0-13", smokers have higher lung capacity than non-smokers. In all other groups, smokers have less lung capacity than non-smokers. The number of samples under 13 gave it a clue about the interesting finding in 1c. And the mean age difference among smokers and non-smokers pointed out that the age difference is more likely to be the reason of higher lung capacity instead of smoking.

#Question 2

##a
```{r}
prob_2 <- (160 / 810)
prob_2
```
##b
```{r}
prob_fewer2 <- (128 + 434) / 810
prob_fewer2
```

##c
```{r}
prob_2OrFewer <- (128 + 434 + 160) / 810
prob_2OrFewer
```

##d
```{r}
prob_more2 <- (64 + 24) / 810
prob_more2
```

##e
```{r}
expectation <- (0 * 128 + 1 * 434 + 2 * 160 + 3 * 64 + 4 * 24) / 810
expectation
```

##f
```{r}
variance <- sum(128 * (0 - expectation) ^ 2,
                434 * (1 - expectation) ^ 2,
                160 * (2 - expectation) ^ 2,
                64 * (3 - expectation) ^ 2,
                24 * (4 - expectation) ^ 2) / 810
variance

sd <- sqrt(variance)
sd
```
